---
title: "083_Get_LIMS_samples"
author: "DHJ"
date: "2024-09-22"
output: 
  html_document:
    keep_md: true
    toc: true  
    toc_float: true
params:
  oracleuser: 
    label: "NIVA initials (3 capital letters)"
    value: "DHJ"
    input: text
  oraclepass: 
    label: "Oracle password"
    value: ""
    input: password
---

**Get last year's LIMS sample data from Nivabasen**    

* Just the sample data,not the concentrations - one row per sample  

* Filtering and joining is done using lazy table loading    
    - "Lazy" means that the data are not actually downloaded until you use 'collect()'  
    - So, if you wait using collect() until you have done the filtering, you avoid downloading unnecessary data  

* This script accesses Nivabasen on Oracle and therefore needs your Oracle user name (your initials) and password. It can be run in two ways:  
    - Interactively - run chunk by chunk (by clicking the little arrow in the corner of the chunk, or Shift+Ctrl+Enter and Shift-Alt-N) - enter username and password when you are asked for it 
    - Or by using "Knit with Parameters..." from the menu (pull-down menu from the Knit button) - enter username and password, then click 'Knit'  

## 1. Libraries

```{r}

# For loading libraries without startup messages
library2 <- function(...) suppressPackageStartupMessages(library(...))

library2(dbplyr)
library2(dplyr)
library2(lubridate)
library2(ggplot2)
library2(stringr)
library2(knitr)

# help(package = "dbplyr")

```

## 2. Settings  
```{r}

measurement_year <- 2023

# Set to TRUE if you want to download the data (ca. 6 minutes)
download_and_save <- TRUE

```



## 3. Set up Oracle connection   

- This part of the script will ask you about your user name and password to Nivabasen  

```{r}

# if the script is run interactively
if (interactive()){
  
  con <- DBI::dbConnect(odbc::odbc(),
                      Driver = "/opt/conda/orahome/libsqora.so.12.1",
                      DBQ = "dbora-niva-prod01.niva.corp:1555/NIVABPRD",
                      UID = rstudioapi::askForPassword("Database username (NIVA initials, 3 capital letters)"),
                      PWD = rstudioapi::askForPassword("Nivabasen password")
  )
  
} else {
# if the script is knitted
  
  # If params$oraclepass is just an empty string, you have probably just clicked the Knit button  
  if (params$oraclepass == ""){
    stop("Password = ''. You must use 'Knit with Parameters...' from the menu (pull-down menu from the Knit button), and enter you password in the dialog")
  }
  
  con <- DBI::dbConnect(odbc::odbc(),
                        Driver = "/opt/conda/orahome/libsqora.so.12.1",
                        DBQ = "dbora-niva-prod01.niva.corp:1555/NIVABPRD",
                        UID = params$oracleuser,
                        PWD = params$oraclepass
  )
  
}

test_connection <- DBI::dbGetQuery(con, "select * from NIVADATABASE.PROJECTS where rownum < 4")
if ("test_connection" %in% ls()){
  if (nrow(test_connection) == 3)
    cat("Connection to Nivadatabase set up and tested")
}

rm(test_connection)

```

## 4. Set up queries  

* These define how to get data from Nivabasen, but don't actually get them  


### Get pointers to tables  

```{r}

projects <- tbl(con, in_schema("NIVADATABASE", "LABWARE_CHECK_SAMPLE")) %>%
  count(ACCOUNT_NUMBER, PROSJEKT, CUSTOMER)

samples <- tbl(con, in_schema("NIVADATABASE", "LABWARE_CHECK_SAMPLE"))

```

## 5. Get samples  

### List projects 

```{r}

# All Merete's projects  
dat_projects <- projects %>%
  filter(upper(CUSTOMER) %in% 'MSC') %>%
  collect()

# View(dat_projects)

```

### Get samples for all years    

```{r}

project <- "O 210200.ANAIN"
cat("Name of project:", project, " \n\n")

labware_sample_all <- samples %>%
  filter(PROSJEKT %in% project) %>%
  collect()

# labware_sample_all_back <- labware_sample_all
```


### b2. Check samples with conflicting year info    

* In some cases, year in SAMPLED_DATE is different from year in TEXT_ID

#### Data for checking  

```{r}
# Check those with conflicting year info  
# All years
df_check_all <- labware_sample_all %>%
  mutate(
    year_sample = year(SAMPLED_DATE),
    month_sample = month(SAMPLED_DATE),
    year_textid = as.numeric(substr(TEXT_ID, 4, 7))) %>%
  # Select those where year in SAMPLED_DATE is different from year in TEXT_ID  
  filter(year_sample != year_textid) %>%
  select(SAMPLED_DATE, TEXT_ID, year_sample, month_sample, year_textid,
         DESCRIPTION, AQUAMONITOR_CODE, TISSUE) %>%
  collect()

```

#### Check  
```{r}

cat("-----------------------------------------------\n")
cat("All years\n-----------------------------------------------\n")
xtabs(~year_sample + year_textid, df_check_all)

cat("-----------------------------------------------\n")
cat("One of the year info is 2023\n-----------------------------------------------\n")
df_check1 <- df_check_all %>%
  filter(
    year_sample == 2023 | year_textid == 2023) %>%
  # Does DESCRIPTION contain 'blank'?  
  mutate(
    blank = grepl("blank", DESCRIPTION, ignore.case = TRUE))

xtabs(~year_sample + year_textid, df_check1)

xtabs(~AQUAMONITOR_CODE + paste0(year_sample, ",", year_textid), df_check1)

cat("-----------------------------------------------\n")
cat("2023 - exclude blanks\n-----------------------------------------------\n")

df_check2 <- df_check1 %>%
  filter(!blank)

xtabs(~AQUAMONITOR_CODE + paste0(year_sample, ",", year_textid), df_check2)

df_check2 %>%
  count(AQUAMONITOR_CODE, SAMPLED_DATE, TISSUE, year_textid, year_sample) %>%
  kable()


```

### b3. Get 'labware_sample'   

* From code above, we found that we can download all samples where either year in SAMPLED_DATE or year in TEXT_ID is 2023 (or both)   
* Download should take around 20 seconds  

```{r}

labware_samples <- labware_sample_all %>%
  filter(
    year(SAMPLED_DATE) == 2023 | as.numeric(substr(TEXT_ID, 4, 7)) == 2023) %>%
  collect()

# Exlude blanks
labware_samples <- labware_samples %>%
  filter(
    !grepl("blank", DESCRIPTION, ignore.case = TRUE)) 

# make timestamp to use in file name (NOTE: UTC = Greenwich mean time used, not local time)
now <- lubridate::now()
timestring <- paste0(substr(now, 1, 10), "_", substr(now, 12, 13), "h", substr(now, 15, 16), "m")  

cat("-----------------------------------------------\n")
cat("Number of samples - 2023 (excluding blanks)\n-----------------------------------------------\n")
nrow(labware_samples)

```

## 6. Save   

```{r}

# Needs only be save if different from previously downloaded data

fn_rds <- paste0("83_Labware_samples_", measurement_year, "_", timestring, ".rds")  
fn_csv <- sub(".rds", ".csv", fn_rds)  

# fn_rds; fn_csv
saveRDS(labware_samples, paste0("Input_data/", fn_rds))
readr::write_csv(labware_samples, paste0("Input_data/", fn_csv))
  
message("Data saved to 'Input_data' as ", sQuote(fn_csv), " (and as a corresponding .rds file)")
  

```
