---
title: "Script 101 appendix"
output: html_document
---

**Various code snippets that were used in the development of script 101; might be useful later** 

## 1. Load libraries and functions   
```{r, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(readxl)
library(readr)


```



## Appendix

### Check ReplicateNO vs SAMPLE_NO2 (which equals LATIN_NAME for data from Nivabasen)
```{r}

dat_new1 %>%
  filter(StationCode == "30B" & year(CatchDateFirst) == 2017 & TissueName == "Lever" & Substance == "As") %>%
  select(StationCode, CatchDateFirst, ReplicateNo, Value) %>%
  arrange(ReplicateNo)

data_through_2018 %>%
  filter(STATION_CODE == "30B" & MYEAR == 2017 & TISSUE_NAME == "Lever" & PARAM == "AS") %>%
  select(MYEAR, STATION_CODE, SAMPLE_NO2, VALUE) %>%
  arrange(SAMPLE_NO2)

```

### Check parameters in fish liver  
Sort by name
```{r}

dat_new1 %>%
  filter(StationCode == "30B" & year(CatchDateFirst) == 2017 & TissueName == "Lever" & ReplicateNo == 1) %>%
  select(Substance, Value) %>%
  arrange(Substance)

data_through_2018 %>%
  filter(STATION_CODE == "30B" & MYEAR == 2017 & TISSUE_NAME == "Lever" & SAMPLE_NO == 1) %>%
  select(PARAM, VALUE) %>%
  arrange(PARAM)

```

Sort by value
```{r}

a <- dat_new1 %>%
  filter(StationCode == "30B" & year(CatchDateFirst) == 2017 & TissueName == "Lever" & ReplicateNo == 1) %>%
  select(Substance, Value) %>%
  arrange(Value)

b <- data_through_2018 %>%
  filter(STATION_CODE == "30B" & MYEAR == 2017 & TISSUE_NAME == "Lever" & SAMPLE_NO == 1) %>%
  select(PARAM, VALUE) %>%
  arrange(VALUE)

left_join(b, a, by = c("VALUE" = "Value"))

# not included in  AqM:
#   WHO dioxins, SN (?!), Aldrin, DDEOP, DDTOP, Heptaklor, HCHA, HCHB, HCHD, HCHG 
# 4-t- -> 4-T-
# 4-n- -> 4-N-

```

Sort by name
```{r}

dat_new1 %>%
  filter(StationCode == "30B" & year(CatchDateFirst) == 2017 & TissueName == "Lever" & ReplicateNo == 1) %>%
  select(Substance, Value) %>%
  arrange(Substance)

data_through_2018 %>%
  filter(STATION_CODE == "30B" & MYEAR == 2017 & TISSUE_NAME == "Lever" & SAMPLE_NO == 1) %>%
  select(PARAM, VALUE) %>%
  arrange(PARAM)

```

### Check parameters in fish muscle    
```{r}

dat_new1 %>%
  filter(StationCode == "30B" & year(CatchDateFirst) == 2017 & TissueName == "Muskel" & ReplicateNo == 1) %>%
  select(Substance, Value) %>%
  arrange(Substance)

data_through_2018 %>%
  filter(STATION_CODE == "30B" & MYEAR == 2017 & TISSUE_NAME == "Muskel" & SAMPLE_NO == 1) %>%
  select(PARAM, VALUE) %>%
  arrange(PARAM)

# TTS -> DRYWT%

```

### Check PBDEs  
```{r}

dat_new1 %>%
  filter(StationCode == "30B" & year(CatchDateFirst) == 2017 & substr(Substance,1,2) == "BD") %>%
  xtabs(~Substance, .)

data_through_2018 %>%
  filter(STATION_CODE == "30B" & MYEAR == 2017 & substr(PARAM,1,2) == "BD") %>%
  xtabs(~PARAM, .)

# "BD183" - > "BDE183"

dat_new1 %>%
  filter(StationCode == "30B" & year(CatchDateFirst) == 2017 & grepl("HBC", Substance)) %>%
  xtabs(~Substance, .) %>% names()

data_through_2018 %>%
  filter(STATION_CODE == "30B" & MYEAR == 2017 & grepl("HBC", PARAM)) %>%
  xtabs(~PARAM, .) %>% names()

# "a-HBCD"   "b-HBCD"   "g-HBCD"   "Sum HBCD"
# "HBCDA"    "HBCDB"    "HBCDG"    "Sum HBCD"

dat_new1 %>%
  filter(StationCode == "30B" & year(CatchDateFirst) == 2017 & substr(Substance,1,2) == "PF") %>%
  xtabs(~Substance, .)

data_through_2018 %>%
  filter(STATION_CODE == "30B" & MYEAR == 2017 & substr(PARAM,1,2) == "PF") %>%
  xtabs(~PARAM, .)

# All ok

dat_new1 %>%
  filter(StationCode == "30B" & year(CatchDateFirst) == 2017 & nchar(Substance) == 2) %>%
  xtabs(~Substance, .)

data_through_2018 %>%
  filter(STATION_CODE == "30B" & MYEAR == 2017 & nchar(PARAM) == 2) %>%
  xtabs(~PARAM, .)

# Metals (all Substance with 2 characters) -> all capitals

```

### Check nonylphenols  
In particular, "4-nonylfenol" used in 2018  
- seems like it is 4-n-NP
```{r}

dat_new1 %>%
  mutate(Year = year(CatchDateFirst)) %>%
  filter(TissueName == "Lever" & Substance %in% c("4-n-NP", "4-t-NP", "4-nonylfenol")) %>%
  mutate(Value = case_when(
    Unit == "mg/kg" ~ Value*1000,
    Unit == "µg/kg" ~ Value,
    TRUE ~ -999)
    ) %>%
  group_by(StationCode, Year, TissueName, Substance) %>%
  summarise(Value = median(Value)) %>%
  ggplot(aes(Year, Value, color = Substance)) +
  geom_point() + geom_line() + 
  facet_wrap(vars(StationCode))

           
```


### Check ReplicateNO vs SAMPLE_NO2 (which equals LATIN_NAME for data from Nivabasen)
```{r}

dat_new1 %>%
  filter(StationCode == "30B" & year(CatchDateFirst) == 2017 & TissueName == "Lever" & Substance == "TTS") %>%
  select(StationCode, CatchDateFirst, ReplicateNo, Value) %>%
  arrange(ReplicateNo)

data_through_2018 %>%
  filter(STATION_CODE == "30B" & MYEAR == 2017 & TISSUE_NAME == "Lever" & PARAM == "DRYWT%") %>%
  select(MYEAR, STATION_CODE, SAMPLE_NO2, VALUE) %>%
  arrange(SAMPLE_NO2)

```

### Check TissueName vs TISSUE_NAME 
Note that the Aquamonitor data doesn't include fish biological effect data   
```{r}

dat_new1 %>%
  filter(year(CatchDateFirst) >= 2017) %>%
  xtabs(~TissueName, .)  %>% sort() %>% names() %>% dput()

data_through_2018 %>%
  filter(MYEAR == 2017) %>%
  xtabs(~TISSUE_NAME, .) %>% sort() %>% names() %>% dput()


```

### Check TaxonName vs LATIN_NAME 
Used to make lookup_latin_name
```{r}

dat_new1 %>%
  filter(year(CatchDateFirst) >= 2017) %>%
  xtabs(~TaxonName, .)  %>% sort() %>% names() %>% dput()

data_through_2018 %>%
  filter(MYEAR == 2017) %>%
  xtabs(~LATIN_NAME, .) %>% sort() %>% names() %>% dput()

```

### Check Unit vs UNIT 
Used to make lookup_latin_name
```{r}

dat_new1 %>%
  filter(year(CatchDateFirst) >= 2017) %>%
  xtabs(~Unit, .)  %>% sort() %>% names() %>% rev() %>% dput()

data_through_2018 %>%
  filter(MYEAR == 2017) %>%
  xtabs(~UNIT, .) %>% sort() %>% names() %>% rev() %>% dput()

dat_new1 %>%
  filter(year(CatchDateFirst) >= 2017 & !Unit %in% c("mg/kg", "µg/kg")) %>%
  xtabs(~Substance + Unit, .)

data_through_2018 %>%
  filter(MYEAR == 2017 & !UNIT %in% c("MG_P_KG", "UG_P_KG")) %>%
  xtabs(~PARAM + UNIT, .)

# "Delta13C", "Delta15N"
# "TTS"

```


### Make table for preferred parameter units   
"Lookup table - preferred parameter units.xlsx"
```{r}

df_units <- data_through_2018 %>%
  count(PARAM, UNIT)

# openxlsx::write.xlsx(df_units, "Input_data/Units2.xlsx")
# Table was corrected manually afterwards

param_multiple_units <- df_units %>%
  count(PARAM) %>%
  filter(n > 1) %>%
  pull(PARAM)

df_units %>%
  filter(PARAM %in% param_multiple_units)


```




### Check TBTs  
Checking in the original data (df1a), downloaded directly from Nivabasen (Biota_chemistry_values)   
Example: 227G:  
- Tributyltinn (TBT) = 2.30  
- Tributyltinn (TBT)-Sn	= 0.96  
- so "Tributyltinn (TBT)" is the ion weigth and "Tributyltinn (TBT)-Sn" is the tim atom weight  
Aquamonitor returns TBT = 0.96, so it returns the tin atom weight  
  
We have had troube mixing up these before - see  
- `\\niva-of5\osl-userdata$\DHJ\Documents\seksjon 212\MIME 2019\TODO intersessionaly.txt`
- `C:\Data\seksjon 212\Milkys_2018\ICES\Check TBT data from Rob.R`   
- Analysis reports:  
    - `K:\Prosjekter\Sjøvann\JAMP\2015\analyser\bestillinger i LIMS\analyserapporter\snegl`   
    - `K:\Prosjekter\Sjøvann\JAMP\2016\analyser\bestillinger i LIMS\analyserapporter` - mangler snegl    
    - `K:\Prosjekter\Sjøvann\JAMP\2017\analyser\Analyserapporter\snegl`  
    - `K:\Prosjekter\Sjøvann\JAMP\2018\analyser\Analyserapporter\snegler`     
    - `K:\Prosjekter\Sjøvann\JAMP\2019\analyser\Analyserapporter\snegler`  

### Data from this script, TBT  
```{r}
df1 <- data_updated %>%
  filter(grepl("TBT", PARAM) | grepl("tributyl", PARAM, ignore.case = TRUE)) %>%
  filter(!grepl("TTBT", PARAM))
```

### Check occurence of TBT/TBTIN  
```{r}
# Some years, TBTIN is also given:
df1 %>%
  xtabs(~PARAM + MYEAR, .)

# Some years, TBTIN is also given:
df1 %>%
  filter(MYEAR == 2007) %>%
  xtabs(~PARAM + STATION_CODE, .)

# But no stations/years have TBTIN and not TBT: 
number_tntin_only <- df1 %>%
  count(STATION_CODE, MYEAR, PARAM) %>%
  tidyr::pivot_wider(names_from = PARAM, values_from = n) %>%
  mutate(TBTIN_only = (TBT > 0) & (TBTIN == 0)) %>%
  filter(TBTIN_only) %>%
  nrow()
number_tntin_only  # 0
```

### Check TBT for one station   
- Up through 2015, data are from the Access database and seems to be correct there 
- Check years 2014-2018 (using pdf lab reports in `K:\Prosjekter\Sjøvann\JAMP\`):
    - 2015: Value given is TBT cation, given correctly
    - 2016: Could not find lab report
    - 2017: Value given is TBT cation, given correctly 
    - 2018: Value given is TBT cation, given correctly 
    - 2019: Value given is TBT tin atom, not given correctly
```{r}
station <- "36G"
station <- "11G"

df1 %>%
  filter(PARAM == "TBT" & STATION_CODE == station) %>%
  select(MYEAR, PARAM, VALUE_WW, VALUE_DW, FLAG1, UNIT)  

```



```{r}


# View(df1b)

# Aquamonitor
df2 <- df %>%
  filter(grepl("TBT", PARAM) | grepl("tributyl", PARAM, ignore.case = TRUE)) %>%
  filter(!grepl("TTBT", PARAM))
df2 %>%
  filter(MYEAR == 2019) %>%
  select(PARAM, VALUE_WW, FLAG1, UNIT)
# View(df2)

```
### 227G etc.  
```{r}
df1b <- df1a %>% filter(grepl("227G", STATION_CODE))
df1b %>%
  filter(grepl("Tributyltinn", NAME)) %>%
  select(NAME, VALUE, FLAG1, UNIT)
```

### Check and save biological effect data     
```{r}

# Get stations
biol_effect_stations <- data_updated %>%
  filter(PARAM %in% c("PYR1O", "PYR1OH", "VDSI") & MYEAR >= 2010) %>%
  count(STATION_CODE) %>%
  pull(STATION_CODE)

# Get data
data_biol_effects <- data_updated %>%
  filter(MYEAR >= 2010 & STATION_CODE %in% biol_effect_stations)
  
df <- data_biol_effects %>%
  filter(grepl("227G", STATION_CODE) & MYEAR >= 2010) %>% # View()
  filter(!PARAM %in% c("HTMEA", "IMPS", "LNFPE", "LNMPE"))
df %>% xtabs(~STATION_CODE + MYEAR, .)
df %>%
  filter(PARAM %in% "VDSI") %>%
  xtabs(~STATION_CODE + MYEAR, .)
View(df)

df2 <- df %>%
  filter(grepl("TBT", PARAM) | grepl("tributyl", PARAM, ignore.case = TRUE)) %>%
  filter(!grepl("TTBT", PARAM))
View(df2)
  xtabs(~PARAM + MYEAR, df2)

if (FALSE){
  
  # Check snail parameters
  data_updated %>%
    filter(grepl("227G", STATION_CODE) & MYEAR >= 2010) %>%
    xtabs(~PARAM + MYEAR, .)
}

# individual-based snail measurements (2010.2013 only)- we skip these:
c("HTMEA", "IMPS", "LNFPE", "LNMPE")  

# "Tributyltinn (TBT)-Sn" 
  
```


### Fix erronous tin data   
Checks ratio ion wight / tin weight in data - if the ratio is close to one, it is assumed that the tim weight is given and a new ion weight is calculated and used to replace old data. Deprecated as we found the source of the confusion.  
```{r}

#
# Ratios for ion wight / tin weight
# For pairs 5 -8 only (for lazyness)
# Can be checked agains plots above
#
ratio <- c(NA,NA,NA,NA, 233.95/118.710, 345.2/118.710, 350/118.710, 736.3/(2*118.710))
cat("Monooktyltinn (MOT) ratio:", ratio[5], "\n")
cat("Dioktyltinn (DOT) ratio:", ratio[6], "\n")
cat("Trifenyltinn (TPhT) ratio:", ratio[7], "\n")
cat("Trisykloheksyltinn (TCHT) ratio:", ratio[8], "\n")

for (i in 5:8){
  i <- 5
  pair <- pairs[[i]]
  cat("ion weight data: \n")
  sel1 <- data_updated$PARAM %in% pair[1]; sum(sel1)  # select ion weight data
  cat("tin weight data: \n")
  sel2 <- data_updated$PARAM %in% pair[2]; sum(sel2)  # select tin weight data
  # Add 'VALUE_WW_tin' and 'Ratio' to the data supposed to be ion weights
  df_for_join <- data_updated[sel2,] %>% 
    select(MYEAR, STATION_CODE, SAMPLE_NO2, VALUE_WW) %>%
    mutate(PARAM = pair[1]) %>%
    rename(VALUE_WW_tin = VALUE_WW)
  # df_for_join
  data_updated2 <- data_updated %>% 
    left_join(df_for_join, by = c("MYEAR", "STATION_CODE", "SAMPLE_NO2", "PARAM")) %>%
    mutate(Ratio = VALUE_WW/VALUE_WW_tin)
  # Select the values that needs correction
  cat("ion weight data that needs correction: \n")
  needs_correction <- data_updated2$Ratio < 1.05 & !is.na(data_updated2$Ratio); sum(needs_correction)
  # Correc this vales by multiplying the values by the ratio of ion weight to tin weight
  data_updated2$VALUE_WW[needs_correction] <- data_updated2$VALUE_WW[needs_correction]*ratio[i]
  
  & MYEAR >= 2014) %>% 
    mutate(PARAM = factor(PARAM, levels = pair)) %>%
    select(MYEAR, STATION_CODE, SAMPLE_NO2, PARAM, VALUE_WW) %>%
    pivot_wider(names_from = "PARAM", values_from = "VALUE_WW", names_sort = TRUE)
  df2$Rat1 = df2[[5]]/df2[[4]]
  df2$Rat2 = df2[[4]]/df2[[5]]  
  dubious_stations <- df2 %>% filter(Rat2 > 0.95 & Rat2 < 1.05) %>% pull(STATION_CODE) %>% unique()
  cat(paste(pair, collapse = " + "), ": ", paste(dubious_stations, collapse = ", "), "\n")
  gg <- ggplot(df2, aes(MYEAR, Rat2, color = STATION_CODE)) + 
    geom_point() +
    labs(title = paste(pair, collapse = " + "))
  print(gg)  
}

```
