---
title: "105_Add_uncertainty"
author: "DHJ"
date: '2022-09-07'
output: html_document
---

For now, we assume that uncertainty is the same for all years (sounds worse than it is?)

## 0. Settings  
```{r}

selected_year <- 2021

```


## 1. Packages  
```{r}

library(dplyr)
library(purrr)
library(readxl)
library(ggplot2)
source("002_Utility_functions.R")

```


## 2. Get data  

### a. Main data  
Read and reformat the most recent data (by default)  
```{r}

# Files 
files <- list_files("Data", pattern = "101_data_updated")

# Pick file
filename <- files[1]
cat("\nLast file:", filename, "\n")
cat("Time since this file was modified: \n")
Sys.time() - file.info(paste0("Data/", filename))$mtime 

# Info
cat("\nIf you want to read a different file, input a different name for 'filename' \n")

# Read data
dat_all <- readRDS(paste0("Data/", filename))

# We save the date part of the text (e.g., '2020-04-23')
# This will be used in part 10, when we save the resulting file
file_date <- substr(filename, 18, 27)    # pick out the part of the text from character no. 17 to no. 26

```

### b. Add parameter (substance) groups  
```{r}

df_paramgroups <- read.csv("Input_data/Lookup_tables/Lookup table - substance groups.csv")  

sel <- !subset(dat_all, MYEAR == selected_year)$PARAM %in% df_paramgroups$PARAM

if (sum(sel) > 0){
  # Should be 'stop'? set this as warning for now 
  warning("Not found:", 
      unique(dat_all$PARAM[sel]) %>% paste(collapse = "; "), "\n")
} else {
  cat("All parameters (PARAM) found in file for parameter (substancroups \n")
}

n1 <- nrow(dat_all)
dat_all <- dat_all %>%
  left_join(df_paramgroups %>% select(PARAM, Substance.Group))
n2 <- nrow(dat_all)

if (n2 > n1){
  stop("PARAM in 'df_paramgroups' not unique! Fix, and recreate dat_all.")
}

cat("\n\n")

check <- dat_all %>%
  filter(is.na(Substance.Group)) %>%
  count(PARAM)

if (nrow(check) > 0){
  # Should be 'stop'? set this as warning for now 
  warning("Some values of 'Substance.Group' in 'df_paramgroups' lacks data! Fix, and recreate dat_all.")
}

cat("\n\n")
cat("'Substance.Group' added to data\n")

# For copying to clipboard -> paste to excel
# dat_all[!sel,] %>% count(PARAM) %>% write.table("clipboard", sep= "\t")

# Table example
# dat_allyears %>% filter(grepl("OP", PARAM)) %>% xtabs(~MYEAR + PARAM, .)

```
### c. Fix PFDcA and PFUdA in cod liver    
- Set values under 0.5 to "< 0.5"  
- selected_year only
```{r}

sel <- with(dat_all, 
            LATIN_NAME == "Gadus morhua" & PARAM == "PFDcA" & MYEAR == selected_year & VALUE_WW < 0.5)
sum(sel)
table(addNA(dat_all$FLAG1[sel]))  

sel2 <- with(dat_all, is.na(FLAG1))
sum(sel & sel2)

sel3 <- sel & sel2
dat_all[sel3,]

# We let this one be
# message("Change ", sum(sel), " PFDcA records")
# dat_all$VALUE_WW[sel] <- 0.5 
# dat_all$FLAG1[sel] <- "<" 
# dat_all$VALUE_DW[sel] <- NA 
# dat_all$VALUE_FB[sel] <- NA 

# PLot again
# plot_raw_data_liver_highlight("PFDcA") # strange values  

```



### d. Fix concentrations below zero  
```{r}

if (FALSE){
  
  dat_all %>%
    filter() %>%
    xtabs(~LATIN_NAME + PARAM + MYEAR, .)
  
}

sel <- with(
  dat_all, 
  VALUE_WW < 0 & !PARAM %in% c("Delta13C") & LATIN_NAME == "Somateria mollissima")
message("Fix ", sum(sel, na.rm = TRUE), " eider duck PCBs with conc. < 0, these are below LOQ")

if (sum(sel, na.rm = TRUE) > 0){
  dat_all$VALUE_WW[sel] <- -dat_all$VALUE_WW[sel]
  dat_all$VALUE_DW[sel] <- -dat_all$VALUE_DW[sel]
  dat_all$VALUE_FB[sel] <- -dat_all$VALUE_WW[sel]
  dat_all$FLAG1[sel] <- "<"
}

```




## 3a. Uncertainty data   

### Check existing uncertainty by year      
```{r}

for (yr in c(2021, 2020, 2019, 2018)){
cat("==========================================\n", yr, "\n==========================================\n")

dat_all %>%
  filter(MYEAR == yr) %>%
  xtabs(~ is.na(UNCERTAINTY), .) %>% print()

dat_all %>%
  filter(MYEAR == yr) %>%
  xtabs(~ (UNCERTAINTY>0), .) %>% print()
}


```
### Check 2020-2021 data  (1)
```{r}

df1 <- dat_all %>%
  mutate(
    Uncert_group = case_when(
      is.na(UNCERTAINTY) ~ "NA",
      UNCERTAINTY == 0 ~ "=0",
      UNCERTAINTY > 0 ~ ">0")
    )

df2020 <- df1 %>%
  filter(MYEAR == 2020 & LATIN_NAME %in% c("Gadus morhua", "Mytilus edulis"))
df2021 <- df1 %>%
  filter(MYEAR == 2021 & LATIN_NAME %in% c("Gadus morhua", "Mytilus edulis"))

df2020 %>%
  count(Substance.Group, Uncert_group) %>%
  tidyr::pivot_wider(c(Substance.Group), values_from = n, names_from = Uncert_group)

df2021 %>%
  count(Substance.Group, Uncert_group) %>%
  tidyr::pivot_wider(c(Substance.Group), values_from = n, names_from = Uncert_group)
```

### Check 2020 data (2)   
- Compare with data set made in 'Uncertainty, combined', below   
- Largely comparable  
```{r}
if (FALSE){

df2020 %>%
  filter(Uncert_group == ">0") %>%
  group_by(Substance.Group) %>%
  summarize(
    Uncertainty = range(UNCERTAINTY) %>% unique() %>% as.character() %>% paste(collapse = "-") 
  ) %>% View("Uncert 2020")

}

```


### Add 'Lab¨' to data  

* Source: Tender. Table 7 in '3-3. Dokumentasjon for Løsningsforslag (MILKYS 2021-2025).docx', see folder 'Input_files_2021\Uncertainty'

* Blåskjell; blandprøver av bløtdelene	
    - NIVA:	PFAS, fettprosent
    - Eurofins:	Metaller, Hg, PAH, PCB7, DDT, klororganiske forbindelser, PBDE, HBCDD, klorerte parafiner (SCCP/MCCP), Organotinn, tørrvekt
    - IFE:	SIA  
* Fisk; Lever, muskel, blod, galle 	
    - NIVA:	PFAS, EROD/CYP1A (lever); ALA-D (blod); OH-pyren (galle)   
    - Eurofins	Hg (muskel); Metaller, PCB7, DDT, klororganiske forbindelser, PBDE, HBCDD, klorerte parafiner (SCCP/MCCP), fettprosent (lever)
    - NILU:	Siloksaner (lever)  
    - IFE:	SIA (muskel)  
* Purpursnegl (strandsnegl); bløtdel	 
    - NIVA:	VDSI (utregnes fra enkeltindivider) (ISI utregnes fra enkeltindivider for strandsnegl)   
    - Eurofins:	Organotinn forbindelser (blandprøver)
* Ærfugl; egg og blod	  
    - NIVA:	PFAS   
    - NILU:	Metaller, Hg, PCB7, PBDE, HBCDD, klorerte parafiner (SCCP/MCCP), siloksaner, fettprosent   
    - IFE:	SIA

```{r}

# dat_all %>%
#   count(Substance.Group, PARAM)  

dat_all %>%
  count(LATIN_NAME)  

dat_all <- dat_all %>%
  mutate(Lab = case_when(
    Substance.Group %in% "Organofluorines" ~ "NIVA",
    PARAM %in% "Fett" & LATIN_NAME %in% "Mytilus edulis" ~ "NIVA",
    Substance.Group %in% c("Biological effects: molecular/biochemical/cellular/assays", "Biomarkers") ~ "NIVA",
    Substance.Group %in% "Siloxans" ~ "NILU",
    Substance.Group %in% "Isotopes" ~ "IFE",  
    LATIN_NAME %in% "Somateria mollissima" ~ "NILU",  
    TRUE ~ "Eurofins")
  )

```

### Uncertainty, Eurofins   
- NOTE: Uncertainty given in data sheet is 2*SE ("expanded uncertainty")   
- Use uncertainty by substance group  
```{r}

# Uncertainty data  
fn <- "Input_data/Uncertainty/til2021-data/2021-2022_MILKYS_QAdata_Eurofins.xlsx"
# excel_sheets(fn)
df_uncert_eurofins1 <- read_excel(fn, sheet = "Uncertainty by group")

table(df_uncert_eurofins1$Uncertainty)

xtabs(~Substance.Group + addNA(Uncertainty), df_uncert_eurofins1)

df_uncert_eurofins2 <- df_uncert_eurofins1 %>%
  filter(!is.na(Uncertainty)) %>%
  filter(!Uncertainty %in% "-")

xtabs(~Substance.Group + addNA(Uncertainty), df_uncert_eurofins2)

df_uncert_eurofins <- df_uncert_eurofins2 %>%
  distinct(Substance.Group, Uncertainty) %>%
  mutate(Uncertainty = as.numeric(Uncertainty))

```

### Uncertainty, NILU     
- NOTE: Uncertainty given in data sheet is 2*SE ("expanded uncertainty")   
- Use uncertainty by substance group  
```{r}

# Uncertainty data  
fn <- "Input_data/Uncertainty/til2021-data/NILU_2021.xlsx"
# excel_sheets(fn)
df_uncert_nilu <- read_excel(fn, sheet = "Uncertainty by group") %>%
  select(Substance.Group, Uncertainty) %>%
  # Percentage formatting in Excel file, so most values are 0-1 - but some are in the 0-100 range 
  mutate(Uncertainty = case_when(
    Uncertainty < 2 ~ Uncertainty,
    Uncertainty > 2 ~ Uncertainty/100)
    )

```

### Uncertainty, NIVA       
- See table 18 in the tender  
    - '3-3. Dokumentasjon for Løsningsforslag (MILKYS 2021-2025).docx', in folder 'Input_files_2021\Uncertainty'
```{r}

df_uncert_niva <- tibble::tribble(
  ~Substance.Group, ~Uncertainty,
  "Organofluorines", 0.25,
  "Fett", 0.15)

```

### Uncertainty, combined  
- Metals at NILU - assumed the same as Eurofins  
- 'Organo-metallic compounds' + Pesticides at Eurofins - from tender  
- The rest - guessed  
```{r}

df_uncert <- bind_rows(
  df_uncert_eurofins %>% mutate(Lab = "Eurofins"),
  df_uncert_nilu %>% mutate(Lab = "NILU"),
  df_uncert_niva %>% mutate(Lab = "NIVA"),
  tibble::tribble(
    ~Substance.Group, ~Uncertainty, ~Lab,
    "Metals and metalloids", 0.20, "NILU",
    "Organo-metallic compounds", 0.30, "Eurofins",
    "Pesticides", 0.30, "Eurofins",
    "Others", 0.30, "Eurofins",    # also largely pesticides/herbicides
    "Biological effects: molecular/biochemical/cellular/assays", 0.40, "NIVA",
    "Biomarkers", 0.15, "NIVA",
    "Isotopes", 0.35, "IFE"
    )
)

```

### Uncertainty, add to data    
```{r}

# Split data set

uncert_group <- dat_all %>%
  mutate(
    Uncert_group = case_when(
      is.na(UNCERTAINTY) ~ "NA",
      UNCERTAINTY == 0 ~ "=0",
      UNCERTAINTY > 0 ~ ">0")
    ) %>%
  pull(Uncert_group)

# 1. Data which already have uncertainty given (2020 only)  
dat_updated1 <- dat_all[uncert_group %in% ">0",] %>%
  mutate(Uncertainty = UNCERTAINTY/100) %>%
  mutate(UNCRT = UNCERTAINTY) %>%
  mutate(METCU = "%")
  
# 2. Data which not not have uncertainty given   
dat_updated2 <- dat_all[uncert_group %in% c("NA", "=0"),] %>%
  left_join(df_uncert, by = c("Substance.Group", "Lab")) %>%
  mutate(
    Uncertainty = case_when(
      Substance.Group %in% "Fat and dry weight" ~ 0.15,
      Substance.Group %in% "Percentage C and N" ~ 0.20,
      TRUE ~ Uncertainty)
  ) %>%
  mutate(UNCRT = 100*Uncertainty) %>%
  mutate(METCU = "%")

dat_updated <- bind_rows(dat_updated1, dat_updated2)

if (nrow(dat_updated) != nrow(dat_all)){
  stop("Number of rows increased - make sure Substance.Group and Lab are unique")
} else {
  cat("Uncertainty added to data \n")
}

```
### Check 1

```{r}

df1 <- dat_updated %>%
  mutate(
    Uncert_group = case_when(
      is.na(Uncertainty) ~ "NA",
      Uncertainty == 0 ~ "=0",
      Uncertainty > 0 ~ ">0")
    )

xtabs(~MYEAR + Uncert_group, df1 %>% filter(MYEAR >= 2000))

```

### Check  
```{r}

check <- dat_updated %>%
  filter(is.na(Uncertainty))

if (nrow(check) == 0){
  cat("Uncertainty added to all data")
} else {
  warning("Still some substances / labs lacking uncertainty - check tables below")
}

if (FALSE){
  
  check %>% count(Substance.Group, Lab)
  check %>% count(Substance.Group, PARAM, Lab)
  
}

```



### Show raw data  
- jjust picks selected columns 
```{r}
#
# data set with selected columns, just for View
#
dat_updated_view <- dat_updated %>% 
  select(
    STATION_CODE, STATION_NAME, SAMPLE_DATE, LATIN_NAME, TISSUE_NAME, PARAM, MYEAR, 
    SAMPLE_NO2, VALUE_WW, FLAG1, Lab,   
    UNCRT, METCU
  )

if (FALSE){
  
  View(dat_updated_view %>% 
         filter(LATIN_NAME == "Gadus morhua" & PARAM == "CB180") %>% 
         arrange(VALUE_WW), title = "CB180")
  View(dat_updated_view %>% 
         filter(LATIN_NAME == "Gadus morhua" & PARAM == "PFUdA") %>% 
         arrange(VALUE_WW), title = "PFUdA")
  View(dat_updated_view %>% 
         filter(LATIN_NAME == "Gadus morhua" & PARAM == "PFDcA") %>% 
         arrange(VALUE_WW), title = "PFDcA")
  View(dat_updated_view %>% 
         filter(LATIN_NAME == "Gadus morhua" & PARAM == "PFOS") %>% 
         arrange(VALUE_WW), title = "PFOS")
}

```


## 4. Save  

### a. R data file    
```{r}

filename <- paste0("Data/105_data_with_uncertainty_", file_date, ".rds")

saveRDS(dat_updated, filename)

cat("Data on sample level, with uncertainty, saved as: \n")
cat(" ", filename, "\n")

if (FALSE){
  # For reading the file instead 
  dat_all <- readRDS(filename)
}

```

