---
title: "105 Download Labware sample data"
author: "DHJ"
date: "5 5 2020"
output: html_document
---

This script reads sample data (including info on pooling of samples etc.) from Nivabasen   

- __NOTE: This must be done from your personal PC, it doesn't work n Jupyterhub__  
- Also, you need to prepare by installing an Oracle client and ge a password to Nivabasen (see the readme file for niRvana)  
  

## 1. Libraries  
```{r, message=FALSE}

library(dplyr)
library(niRvana)
library(lubridate)

```


## 2. Set username and pasword for Nivabasen  
```{r}

set_credentials()

```

## 3. Get samples  
```{r}

# Get projects, but doesn't really make sense since the projects names fiffer from those in PROJECT in table LABWARE_CHECK_SAMPLE
if (FALSE){
  df_proj <- get_projects()
  grep("CEMP", df_proj$PROJECT_NAME, value = TRUE)
}

dat <- get_nivabase_data("select * from NIVADATABASE.LABWARE_CHECK_SAMPLE where lower(PROSJEKT) like '%milkys%'")
nrow(dat) # 3568

sel <- grepl("2019", dat$TEXT_ID)
sum(sel)   # 656
mean(sel)  # 0.18

```

## 4. Read previously downloaded data and compare 
```{r}

# NOT WRItTEN YET


```


## 5. Save  
```{r}

# Needs only be save if different from previously downloaded data

file_date <- Sys.time() %>% substr(1, 10)
file_date

filename <- paste0("Labware_samples_2019_", file_date, ".rds")
# filename

saveRDS(dat, paste0("Input_data/", filename))

```


