---
title: "109_Adjust_for_fish_length"
output: 
  html_document:
    toc: true
    toc_float: true
---

## 1. Packages
```{r}

# install safejoin, if that has not already been done
# It may also want to update some packages; if so,  go down to the console 
if (!"safejoin" %in% installed.packages()){
  devtools::install_github(repo = "moodymudskipper/safejoin", upgrade = "always")
}

library(dplyr)
library(purrr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(tidyr)
library(safejoin)  # from https://github.com/moodymudskipper/safejoin 

source("002_Utility_functions.R")
source("109_Adjust_for_fish_length_functions.R")

```

### Year  
```{r}

selected_year <- 2020

```


## 2. Data

### a. Main data  
Read and reformat the most recent data (by default)  
```{r}

# Files 
files <- list_files("Data", pattern = "101_data_updated")

# Pick file
filename <- files[1]
cat("\nLast file:", filename, "\n")
cat("Time since this file was modified: \n")
Sys.time() - file.info(paste0("Data/", filename))$mtime 

# Info
cat("\nIf you want to read a different file, input a different name for 'filename' \n")

# Read data
data_all <- readRDS(paste0("Data/", filename))

# We save the date part of the text (e.g., '2020-04-23')
# This will be used in part 10, when we save the resulting file
file_date <- substr(filename, 18, 27)    # pick out the part of the text from character no. 17 to no. 26

```


### b. Length data for most recent year    
```{r}

# Data on length  
filename <- paste0("Data/150_Fish_individual_data_", selected_year, ".rds")
data_ind_fish <- readRDS(filename)

cat("\nStations:")
table(addNA(data_ind_fish$STATION_CODE))
cat("\n")

#
# Check if data have STATION_CODE
#
sel <- is.na(data_ind_fish$STATION_CODE)  


if (sum(sel) > 0){
  stop("WARNING: Some data lack STATION_CODE")
  cat("\nStations:")
  table(addNA(data_ind_fish$STATION_CODE[sel]))
} else {
  message("Test 1 OK (all data have STATION_CODE)")
}

#
# Check if data have length
#
sel <- is.na(data_ind_fish$Length)  

if (sum(sel) > 0){
  warning("WARNING: Some data lack length")
  cat("\nStations:")
  table(addNA(data_ind_fish$STATION_CODE[sel]))
} else {
  message("Test 2 OK (all data have length)")
}

```


### c1. Sample data for most recent year   
Read and reformat the most recent data (by default)  
```{r}

files <- dir("Input_data", "Labware_samples_") %>% rev()
filename <- files[1] 
df_samples <- readRDS(paste0("Input_data/", filename))
# xtabs(~TISSUE, df_samples)

# 
# Prepare sample data    
#
df_fishsamples <- df_samples %>%
  select(AQUAMONITOR_CODE, BIOTA_SAMPLENO, TISSUE, X_BULK_BIO, DESCRIPTION) %>%
  distinct(AQUAMONITOR_CODE, TISSUE, BIOTA_SAMPLENO, X_BULK_BIO, DESCRIPTION) %>%
  mutate(
    TISSUE_NAME = str_sub(TISSUE, start = 4),
    # Extract SPECIMEN_NO by getting the last number of DESCRIPTION (using regex):
    SPECIMEN_NO = str_extract_all(DESCRIPTION, "[0-9]+") %>% map(rev) %>% map_chr(~.[1])) %>% 
  select(-TISSUE)

# Check sample number of muscle samples
check <- df_fishsamples %>% 
  filter(TISSUE_NAME == "Muskel") %>%
  filter(!is.na(X_BULK_BIO) & SPECIMEN_NO != X_BULK_BIO)
cat("Number of muscle samples where fish number isn't equal to sample number:", nrow(check), "\n")  
if (nrow(check) > 0){
  cat("- stations of those: \n")
  cat(check %>% pull(AQUAMONITOR_CODE) %>% unique())  
}

if (FALSE){
  View((df_fishsamples %>% filter(TISSUE_NAME == "Muskel"))[sel,])
}
```

### c2. Sample data, special case: galle (bile) at 30B   
```{r}

# Special case: galle (bile) at 30B
sel_1 <- grepl("30B", df_samples$DESCRIPTION)
message(sum(sel_1), " samples marked 30B")

sel <- is.na(df_samples$AQUAMONITOR_CODE) & sel_1
message(sum(sel), " samples from 30B that lack AQUAMONITOR_CODE")

# Check: 
#  View(df_samples[sel,])

# Add galle (bile) at 30B, if needed
if (sum(sel) > 0) {
  
  df_fishsamples_extra <- df_samples[sel,] %>%
    mutate(AQUAMONITOR_CODE = "30B",
           BIOTA_SAMPLENO = as.numeric(X_BULK_BIO),
           X_BULK_BIO = NA,
           TISSUE_NAME = TISSUE) %>%
    select(AQUAMONITOR_CODE, TISSUE_NAME, BIOTA_SAMPLENO, X_BULK_BIO)
  
  df_fishsamples <- bind_rows(
    df_fishsamples %>% filter(!is.na(AQUAMONITOR_CODE)), 
    df_fishsamples_extra
    )
}

# Check lines with AQUAMONITOR_CODE = NA, if any
if (FALSE){

    sel <- is.na(df_samples$AQUAMONITOR_CODE); sum(sel)
  View(df_samples[sel,])
  
  sel <- is.na(df_fishsamples$AQUAMONITOR_CODE); sum(sel)
  View(df_fishsamples[sel,])

  }

```

## 3. Prepare 'df_lnmea'  
Data with mean length per sample  

### Add rows for biol. effects in cod
```{r}

# Only a check of tissue
# test <- data_ind2 %>%
#   filter(MYEAR == selected_year & PARAM %in% c("PA1OH", "PYR1OH", "BAP3OH", "ALAD", "EROD")) %>%
#   count(PARAM, TISSUE_NAME)

# We must specify Lever - microsome (single specimen-sample) for EROD, otherwise it is 
# mixed up with Lever (often pooled sample) when Length is added
df_fishsamples_bioleff <- data_all %>%
  filter(MYEAR == selected_year & PARAM %in% c("PA1OH", "PYR1OH", "BAP3OH", "ALAD", "EROD")) %>%
  distinct(STATION_CODE, SAMPLE_NO2, TISSUE_NAME) %>%
  rename(AQUAMONITOR_CODE = STATION_CODE,
         BIOTA_SAMPLENO = SAMPLE_NO2) %>%
  mutate(TISSUE_NAME = ifelse(TISSUE_NAME == "Lever", "Lever - microsome", TISSUE_NAME))

# Check if these data have already been added
check <- df_fishsamples %>% filter(AQUAMONITOR_CODE == "15B" & TISSUE_NAME == "Galle" & BIOTA_SAMPLENO > 0) %>% nrow()

# If not already added
if (nrow(df_fishsamples_bioleff) > 0 & check == 0){
  df_fishsamples <- bind_rows(
    df_fishsamples,
    df_fishsamples_bioleff
  )
  cat("Rows for biol. effects in cod added\n")
}

```

### df_fishsamples2 - split X_BULK_BIO
Split up X_BULK_BIO (the fish each samples consists of) so each fish gets a separate row  
* E.g., a pooled liver sample from 3 fish becomes 3 rows
```{r}

df_fishsamples_temp <- tidyr::separate_rows(df_fishsamples, X_BULK_BIO)

#
# Check that all 'X_BULK_BIO' are numeric
#
check <- check_variable_is_numeric(df_fishsamples_temp, "X_BULK_BIO")

#
# Special case for 2020 data (after check above) 
# Remove "G" from someeider duck X_BULK_BIO    
#
sel <- df_fishsamples_temp$AQUAMONITOR_CODE == "19N"
df_fishsamples_temp$X_BULK_BIO[sel] <- sub("G", "", df_fishsamples_temp$X_BULK_BIO[sel])

# Repeat check - should be OK now
check <- check_variable_is_numeric(df_fishsamples_temp, "X_BULK_BIO")

#
# Special case, end
#

df_fishsamples2 <- df_fishsamples_temp %>%
  mutate(X_BULK_BIO = as.numeric(X_BULK_BIO))

cat("\n")
message("Number of rows in 'df_fishsamples': ", nrow(df_fishsamples))   # 1710
message("Number of rows in 'df_fishsamples2': ", nrow(df_fishsamples2))   # 1710

if (FALSE){
 
  # Show example of a pooled fish sample 
  sel <- grepl(",", df_fishsamples$X_BULK_BIO, fixed = TRUE)
  df_example <- df_fishsamples[which(sel)[1],]
  
  df_fishsamples %>%
    filter(AQUAMONITOR_CODE == df_example$AQUAMONITOR_CODE & 
             TISSUE_NAME == df_example$TISSUE_NAME & BIOTA_SAMPLENO == df_example$BIOTA_SAMPLENO)
  df_fishsamples2 %>%
    filter(AQUAMONITOR_CODE == df_example$AQUAMONITOR_CODE & 
             TISSUE_NAME == df_example$TISSUE_NAME & BIOTA_SAMPLENO == df_example$BIOTA_SAMPLENO)
  
}

```


### Add Length to df_fishsamples2
```{r}

df_fishsamples3 <- df_fishsamples2 %>%
  rename(STATION_CODE = AQUAMONITOR_CODE,
         SAMPLE_NO2 = BIOTA_SAMPLENO,
         Fish_no = X_BULK_BIO) %>%
  mutate(Fish_no = 
           case_when(is.na(Fish_no) ~ SAMPLE_NO2,
                     !is.na(Fish_no) ~ Fish_no)
         )

if (FALSE){
  df_fishsamples3 %>% xtabs(~addNA(STATION_CODE), .)
  df_fishsamples3 %>% xtabs(~addNA(Fish_no), .)
  data_ind_fish %>% xtabs(~addNA(STATION_CODE), .)
  data_ind_fish %>% xtabs(~addNA(Fish_no), .)
}

df_fishsamples3 <- df_fishsamples3 %>%
  safe_left_join(data_ind_fish %>% 
                   select(STATION_CODE, Fish_no, Length, Weight),
                 by = c("STATION_CODE", "Fish_no"),
                 na_matches = "never",
                 check = "v")

```

### LNMEA (mean length per sample)
```{r}
df_lnmea <- df_fishsamples3 %>%
  mutate(MYEAR = selected_year) %>%
  ungroup() %>%
  group_by(MYEAR, STATION_CODE, TISSUE_NAME, SAMPLE_NO2) %>%
  summarise(LNMEA = 10*mean(Length, na.rm = TRUE),             # multiply by 10 to get millimeters
            WTMEAN = mean(Weight, na.rm = TRUE),
            .groups = "drop"
            ) %>%
  ungroup()

message("nrow(df_lnmea): ", nrow(df_lnmea))  # 752

```

## 4. Add mean length (LNMEA) to last year's data  

### Pick last year's data and add LNMEA  
```{r}

data_lastyear <- data_all %>%
  filter(MYEAR == selected_year) %>%
  # remove LNMEA for last year (empty anyway)
  select(-LNMEA) %>%
  # Also here we must specify Lever - microsome for EROD
  mutate(TISSUE_NAME = case_when(
    TISSUE_NAME == "Lever" & PARAM %in% "EROD" ~ "Lever - microsome",
    TRUE ~ TISSUE_NAME)
  ) %>%
  # add LNMEA by joining that column from df_lnmea 
  safe_left_join(
    df_lnmea, 
    by = c("MYEAR", "STATION_CODE", "TISSUE_NAME", "SAMPLE_NO2"),
    na_matches = "never",
    check = "bCV") %>%
  # Change "Lever - microsome" back (in order to fit with the rest of the data)
  mutate(TISSUE_NAME = case_when(
    TISSUE_NAME == "Lever - microsome" ~ "Lever", 
    TRUE ~ TISSUE_NAME)
  )

message("nrow(data_all): ", nrow(data_all))  # 752
message("nrow(data_all), selected year: ", nrow(data_all %>% filter(MYEAR == selected_year)))  # 752
message("nrow(data_lastyear): ", nrow(data_lastyear))  # 752

```

### Checks
```{r}

if (FALSE){
  
  data_lastyear %>% 
    filter(STATION_CODE == "30B" & TISSUE_NAME == "Galle") %>% 
    select(MYEAR, STATION_CODE, TISSUE_NAME, SAMPLE_NO2, PARAM, LNMEA)
  
  data_lastyear %>% 
    filter(STATION_CODE == "15B" & TISSUE_NAME == "Lever" & SAMPLE_NO2 == 8) %>% 
    select(MYEAR, STATION_CODE, TISSUE_NAME, SAMPLE_NO2, PARAM, LNMEA)
  
  # data_ind2_old
  
}

```

### Check uniqueness
```{r}
check <- data_lastyear %>%
  ungroup() %>%
  group_by(STATION_CODE, LATIN_NAME, TISSUE_NAME, MYEAR, SAMPLE_NO2, UNIT, PARAM) %>%
  mutate(n = n()) %>%
  filter(n > 1)
nrow(check)  # should be 0
```

### Create data_all2 by adding last years' data to the rest  
Creates data_all2  

```{r}

# names(data_all)
# names(data_lastyear)

data_all2 <- bind_rows(
  data_all %>% filter(MYEAR < selected_year),
  data_lastyear %>% select(-WTMEAN)
) %>%
  mutate(VALUE_WWa = NA,
         VALUE_DWa = NA,
         VALUE_FBa = NA
         )

```



## 5. Length adjustment - preparations 
### a. Make data set 'data_lengthreg' to save regression results
```{r}

data_lengthreg <- subset(data_all2, MYEAR %in% 1980:2017 & is.finite(VALUE_WW) &
                           LATIN_NAME %in% "Gadus morhua") %>%
    group_by(STATION_CODE, PARAM, LATIN_NAME, TISSUE_NAME) %>%
    summarise(N = n(), yr_start = min(MYEAR), yr_end = max(MYEAR)) %>%
    filter(yr_end >= 2012 & N >= 50)  # need at least 50 observations, and at least until 2012
nrow(data_lengthreg) # 865

#
# If present, delete HG and C/N ++ (but not DRYWT%)  in fish liver
#
cat("\n")
cat("If present, delete HG and C/N ++ (but not DRYWT%) in fish liver \n")
sel <- with(data_lengthreg, LATIN_NAME %in% 'Gadus morhua' & PARAM %in% c("HG","C/N","Delta13C","Delta15N") & TISSUE_NAME %in% "Lever")
sum(sel)  # 0
data_lengthreg <- data_lengthreg[!sel,]  

#
# Delete everyting "not HG" in fish muscle (some PCB data)
#
cat("\n\n")
cat("Delete non-fitting parameters from fish muscle (some PCB data) \n")
sel <- with(data_lengthreg, LATIN_NAME %in% 'Gadus morhua' & !PARAM %in% c("HG","C/N","Delta13C","Delta15N", "DRYWT%") & TISSUE_NAME %in% "Muskel")
sum(sel)  # 19
table(data_lengthreg$PARAM[sel])
data_lengthreg <- data_lengthreg[!sel,]

cat("\n")
cat("Resulting time series: \n")
nrow(data_lengthreg) # 846
head(data_lengthreg) 

ggplot(data_lengthreg, aes(N)) + geom_histogram(binwidth = 1)

```


### b. Add variables to be filled in data_lengthreg
```{r}

# names(data_lengthreg)

k <- ncol(data_lengthreg)   # for setting col. numbers (col_ww etc.)
data_lengthreg <- data.frame(
  data_lengthreg,
  WW_Est = NA, WW_SE = NA, WW_P = NA, WW_N = NA, WW_Nreg = NA, 
  DW_Est = NA, DW_SE = NA, DW_P = NA, DW_N = NA, DW_Nreg = NA, 
  FB_Est = NA, FB_SE = NA, FB_P = NA, FB_N = NA, FB_Nreg = NA)

col_ww <- k + 1:5
col_dw <- k + 1:5 + 5
col_fb <- k + 1:5 + 10
 
```


### c. Find parameters that includes many zeros (so 'a' must be >0)  
If <1% is at/below zero, we just delete the zeros instead  
```{r}

tab <- xtabs(~(VALUE_WW <= 0) + PARAM, subset(data_all2, LATIN_NAME %in% "Gadus morhua"))
tab2 <- tab[,tab[2,] > 0]
tab2

# Percent at or below zero
percent_zero <- 100*tab2[2,]/tab2[1,]

# If <1% is at/below zero, we delete the zeros
param_delete_zeros <- names(percent_zero)[percent_zero < 1]

# Show them
cat("\n")
cat("param_delete_zeros: \n")
param_delete_zeros

# Delete those
sel <- with(data_all2, !is.na(VALUE_WW) & VALUE_WW <= 0 & PARAM %in% param_delete_zeros); sum(sel)  # 0
data_all2 <- data_all2[!sel,]

# For the rest, we don't log values before regression
param_incl_zero <- names(percent_zero)[percent_zero >= 1]

# Show them
cat("\n")
cat("param_incl_zero: \n")
param_incl_zero

```


### d. Check uniqueness
```{r}

check1 <- data_all2 %>%
  ungroup() %>%
  group_by(STATION_CODE, LATIN_NAME, TISSUE_NAME, MYEAR, SAMPLE_NO2, UNIT, PARAM) %>%
  mutate(n = n()) %>%
  filter(n > 1)
nrow(check1)  # should be zero!

check2 <-data_lengthreg %>%
  ungroup() %>%
  group_by(STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM) %>%
  mutate(n = n()) %>%
  filter(n > 1)
nrow(check2)  # should be zero!
  
```

## 6a. Test adjustment   

### Select data  
```{r}

station <- "13B"
par <- "CB118"

cat("Chosen", par, "at", station, "\n\n")

# Chose data in 'data_lengthreg'
i <- with(data_lengthreg, STATION_CODE %in% station & PARAM %in% par) %>% which()
cat("Chose line", i, "in 'data_lengthreg' \n\n")

# Select data
sel <- with(data_all2, 
            is.finite(VALUE_WW) & is.finite(LNMEA) & 
              PARAM %in% data_lengthreg$PARAM[i] & 
              LATIN_NAME %in% data_lengthreg$LATIN_NAME[i] &
              TISSUE_NAME %in% data_lengthreg$TISSUE_NAME[i] & 
              STATION_CODE %in% data_lengthreg$STATION_CODE[i])
df <- data_all2[sel, ]

cat(nrow(df), "rows selected from data_all \n")


```

### Show data byyear and length    
```{r, fig.width = 9, fig.height=3}

gg1 <- df %>%
  mutate(Above_LOQ = is.na(FLAG1)) %>%
  ggplot(aes(MYEAR, VALUE_WW, color = Above_LOQ)) +
  geom_point() +
  scale_y_log10()

gg2 <- df %>%
  mutate(Above_LOQ = is.na(FLAG1)) %>%
  ggplot(aes(LNMEA, VALUE_WW, color = Above_LOQ)) +
  geom_point() +
  geom_smooth(method = "lm", formula = 'y ~ x') +
  scale_y_log10()

cowplot::plot_grid(gg1, gg2, nrow = 1)

```


### Set `a`
```{r}

if (data_lengthreg$PARAM[i] %in% param_incl_zero){
  a <- 1                        # if the parameter includes zero, we add 1 before log
} else {
  a <- 0
}

cat("'a' (constant that will be used in log-transformation) set to", a)

```

### Perform test adjustment
```{r, fig.width = 9, fig.height=3}

result <- try(adjust_selected(data_all2, sel, var = "VALUE_WW", a = a))        # adjust_selected()
if (class(result) != "try-error"){
  data_all2$VALUE_WWa[sel] <- result$value
  data_lengthreg[i, col_ww] <- result$summary[1:5]
} 

# Printout
cat("\n\n")
if (class(result) != "try-error"){
  cat("Regression succeeded \n")
  cat(length(result$value), "values of 'VALUE_WWa' calculated and substituted in 'data_all2' \n")
  cat("\nRegression summary: \n")
  result$summary %>% print()
  cat("\n")

  gg1 <- data_all2[sel,] %>%
    mutate(Above_LOQ = is.na(FLAG1)) %>%
    ggplot(aes(MYEAR, VALUE_WWa, color = Above_LOQ)) +
    geom_point() +
    scale_y_log10()
  
  gg2 <- data_all2[sel,]  %>%
    mutate(Above_LOQ = is.na(FLAG1)) %>%
    ggplot(aes(LNMEA, VALUE_WWa, color = Above_LOQ)) +
    geom_point() +
    geom_smooth(method = "lm", formula = 'y ~ x') +
    scale_y_log10()
  
  cowplot::plot_grid(gg1, gg2, nrow = 1)
  
} else {
  cat("Regression failed \n")
}



```



## 6b. Performing the adjustment  
Takes 2-3 minutes on DHJ's PC    
```{r, error=FALSE, warning=FALSE, message=FALSE, results='hide'}

if (FALSE){
  par <- "CB118"
  data_all2 %>% 
    filter(STATION_CODE %in% "15B" & PARAM %in% par & MYEAR == "selected_year" & SAMPLE_NO2 == 1) 
  data_all2 %>% 
    filter(STATION_CODE %in% "15B" & PARAM %in% par & MYEAR == "selected_year" & SAMPLE_NO2 == 1) 
}

i_max <- nrow(data_lengthreg)
# i_max <- 700

# Doesn't work on Linux
# pb <- winProgressBar(max=i_max)

# A minute or two
for (i in 1:i_max){

  #
  # Set 'a' (constant that will be used in log-transformation) 
  #
  if (data_lengthreg$PARAM[i] %in% param_incl_zero){
    a <- 1                        # if the parameter includes zero, we add 1 before log
  } else {
    a <- 0
  }

  #
  # Select data, wet-weight
  #
  sel <- with(data_all2, 
    is.finite(VALUE_WW) & is.finite(LNMEA) & 
    PARAM %in% data_lengthreg$PARAM[i] & 
    LATIN_NAME %in% data_lengthreg$LATIN_NAME[i] &
    TISSUE_NAME %in% data_lengthreg$TISSUE_NAME[i] & 
    STATION_CODE %in% data_lengthreg$STATION_CODE[i])
  df <- data_all2[sel, ]
  nrow(df)
  # debugonce(adjust_selected2)
  # print(subset(data_lengthreg[i,], select = c(PARAM, LATIN_NAME, TISSUE_NAME, STATION_CODE)))
  # cat("WW\n")
  # result <- try(adjust_selected2(data_all2, sel, var = "VALUE_WW"))     # adjust_selected2()
  
  #
  # Perform regression, wet-weight data
  #
  result <- try(adjust_selected(data_all2, sel, var = "VALUE_WW", a = a))        # adjust_selected()
  if (class(result) != "try-error"){
    data_all2$VALUE_WWa[sel] <- result$value
    data_lengthreg[i, col_ww] <- result$summary[1:5]   # number 6 is "a"
    }
  #cat("DW\n")

  #
  # Select data, dry-weight
  #
  sel <- with(data_all2, 
    is.finite(VALUE_DW) & is.finite(LNMEA) & 
    PARAM %in% data_lengthreg$PARAM[i] & 
    LATIN_NAME %in% data_lengthreg$LATIN_NAME[i] &
    TISSUE_NAME %in% data_lengthreg$TISSUE_NAME[i] & 
    STATION_CODE %in% data_lengthreg$STATION_CODE[i])
  #
  # Perform regression
  #
  result <- try(adjust_selected(data_all2, sel, var = "VALUE_DW", a = a))        # adjust_selected()
  if (class(result) != "try-error"){
    data_all2$VALUE_DWa[sel] <- result$value
    data_lengthreg[i, col_dw] <- result$summary[1:5]   # number 6 is "a"
    }
  #
  # Select data, fat-weight
  #
  sel <- with(data_all2, 
    is.finite(VALUE_FB) & is.finite(LNMEA) & 
    PARAM %in% data_lengthreg$PARAM[i] & 
    LATIN_NAME %in% data_lengthreg$LATIN_NAME[i] &
    TISSUE_NAME %in% data_lengthreg$TISSUE_NAME[i] & 
    STATION_CODE %in% data_lengthreg$STATION_CODE[i])
  #
  # Perform regression
  #
  result <- try(adjust_selected(data_all2, sel, var = "VALUE_FB", a = a))        # adjust_selected()
  if (class(result) != "try-error"){
    data_all2$VALUE_FBa[sel] <- result$value
    data_lengthreg[i, col_fb] <- result$summary[1:5]   # number 6 is "a"
    }
  # setWinProgressBar(pb, i)
  }
# close(pb)

data_all2 <- as_tibble(data_all2)

```

## 7. Tables/statistics  

### a. Results for example parameter  

```{r}

param <- "CB118"

# How many have length?
data_all2 %>%
  filter(PARAM == param & TISSUE_NAME == "Lever" & MYEAR >= 2013)  %>%
  group_by(STATION_CODE, MYEAR) %>%
  summarize(Prop_length = mean(!is.na(LNMEA)), .groups = "drop") %>%
  tidyr::pivot_wider(names_from = "MYEAR", values_from = Prop_length)

# Of those data that have length, how many have length-addjusted CB118?
df %>%
  filter(PARAM == param & TISSUE_NAME == "Lever" & MYEAR >= 2013 & !is.na(LNMEA))  %>%
  group_by(STATION_CODE, MYEAR) %>%
  summarize(Prop_adjusted = sum(!is.na(VALUE_WWa))/sum(!is.na(VALUE_WW)), .groups = "drop") %>%
  tidyr::pivot_wider(names_from = "MYEAR", values_from = Prop_adjusted)

```

### b. Proportion of data that has length, per cod time series (since 2010)  
```{r}

data_all2 %>%
  filter(LATIN_NAME %in% "Gadus morhua" & MYEAR >= 2010)  %>%
  group_by(PARAM, STATION_CODE) %>%
  summarize(Percent_length = 100*mean(!is.na(LNMEA)), .groups = "drop") %>%
  ggplot(aes(Percent_length)) +
  geom_histogram()

```
### c. Proportion of length-adjusted time series per parameter, for cod data that has length (since 2010)  
```{r}

# 
data_all2 %>%
  filter(LATIN_NAME %in% "Gadus morhua" & MYEAR >= 2010 & !is.na(LNMEA))  %>%
  group_by(PARAM, STATION_CODE) %>%
  summarize(Prop_adjusted = sum(!is.na(VALUE_WWa))/sum(!is.na(VALUE_WW)), .groups = "drop") %>%
  group_by(PARAM) %>%
  summarize(Percent_adjusted = 100*mean(Prop_adjusted), .groups = "drop") %>%
  ggplot(aes(Percent_adjusted)) +
  geom_histogram()

```


### d. Check uniqueness
```{r}

check <- data_all2 %>%
  ungroup() %>%
  group_by(STATION_CODE, LATIN_NAME, TISSUE_NAME, MYEAR, SAMPLE_NO2, UNIT, PARAM) %>%
  mutate(n = n()) %>%
  filter(n > 1)

if (nrow(check) > 0){
  cat("WARNING!", nrow(check), "records in the data seem to be duplicates. Use 'View(check)' to see them. \n")
} else {
  cat("No duplicates found in the data. \n")
}

```

### e. Test graphs
```{r, fig.width=9, fig.height=6}
par <- "CB_S7"
par <- "HG"

if (FALSE){
  # Not length-adjusted 
  data_all2 %>% 
    filter(LATIN_NAME %in% "Gadus morhua" & PARAM %in% par) %>%
    group_by(STATION_CODE) %>%
    mutate(n_year = length(unique(MYEAR))) %>%
    filter(n_year >= 7) %>%
    ggplot(aes(MYEAR, VALUE_WW, color = TISSUE_NAME)) +
    geom_point() +
    scale_y_log10() +
    facet_wrap(vars(STATION_CODE)) +
    labs(title = paste(par, "in cod"))
}

# Length-adjusted 
data_all2 %>% 
  filter(LATIN_NAME %in% "Gadus morhua" & PARAM %in% par) %>%
  group_by(STATION_CODE) %>%
  mutate(n_year = length(unique(MYEAR))) %>%
  filter(n_year >= 7) %>%
  ggplot(aes(MYEAR, VALUE_WWa, color = TISSUE_NAME)) +
    geom_point() +
    scale_y_log10() +
    facet_wrap(vars(STATION_CODE)) +
  labs(title = paste(par, "in cod"))

# Length-adjusted vs unajusted
data_all2 %>% 
  filter(LATIN_NAME %in% "Gadus morhua" & PARAM %in% par) %>%
  group_by(STATION_CODE) %>%
  mutate(n_year = length(unique(MYEAR))) %>%
  filter(n_year >= 7) %>%
  ggplot(aes(VALUE_WW, VALUE_WWa, color = TISSUE_NAME)) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0)  +
    scale_x_log10() +
    scale_y_log10() +
    facet_wrap(vars(STATION_CODE)) +
    labs(title = paste(par, "in cod"))

# Time series for one tissue
tissue <- case_when(
  par == "HG" ~ "Muskel",
  TRUE ~ "Lever"
)

data_all2 %>% 
  filter(LATIN_NAME %in% "Gadus morhua" & PARAM %in% par & 
           TISSUE_NAME %in% tissue) %>%
  tidyr::pivot_longer(c(VALUE_WW, VALUE_WWa), names_to = "Type", values_to = "Concentration", 
                      names_prefix = "VALUE_") %>%
  group_by(STATION_CODE, MYEAR, Type) %>%
  summarise(Median = median(Concentration), .groups = "drop") %>%
  group_by(STATION_CODE) %>%
  mutate(n_year = length(unique(MYEAR))) %>%
  filter(n_year >= 7) %>%
  ggplot(aes(MYEAR, Median, color = Type)) +
    geom_line() +
    geom_point() +
    scale_y_log10() +
    facet_wrap(vars(STATION_CODE)) +
  labs(title = paste0(par, " in cod (tissue = ", tissue))

```



## 8. Save data   
The data should include one column for whether data is pooled, and fish number of the unpooled ones

### a. Make data to add to the raw data    
```{r}
# Summarise to one line per sample, including 'Fish_number'
df_fishsamples3_summ <- df_fishsamples3 %>%
  group_by(STATION_CODE, TISSUE_NAME, SAMPLE_NO2) %>%
  summarise(
    Pool_size = n(),
    Fish_number = first(Fish_no),
    .groups = "drop"
  )

# Set 'Fish_number' to NA if it's a pooled sample
#   (only meaningful if the sample is from a single fish)
sel <- df_fishsamples3_summ$Pool_size > 1
df_fishsamples3_summ$Fish_number[sel] <- NA

# Pooling of samples
cat("Number of fish pooled per sample \n-----------------------------------\n")
xtabs(~TISSUE_NAME + Pool_size, df_fishsamples3_summ)

```
### b. Add columns for pooling and fish number  
```{r}

data_all2 <- data_all2 %>%
  safe_left_join(
    df_fishsamples3_summ,
    by = c("STATION_CODE", "TISSUE_NAME", "SAMPLE_NO2"),
    check = "bCV"
  )

if (FALSE){

  # For manual check  
  data_all2 %>%
    filter(MYEAR == selected_year & STATION_CODE == "13B" & 
             TISSUE_NAME == "Lever" & PARAM == "CD") %>%
    select(MYEAR, STATION_CODE, TISSUE_NAME, SAMPLE_NO2, 
           Pool_size, Fish_number, PARAM, VALUE_WW, FLAG1) %>%
    View()

}
```

### c. R data file    
```{r}

filename <- paste0("Data/109_adjusted_data_", file_date, ".rds")

saveRDS(data_all2, filename)

cat("Data of medians saved as: \n")
cat(" ", filename, "\n")

if (FALSE){
  # For reading the file instead 
  data_all2 <- readRDS(filename)
}

```

### d. csv file
```{r}

fn <- paste0("Data/10_data_all2_", file_date, ".csv")
readr::write_excel_csv(data_all2, fn, na = "")

```

### c. Save last years as Excel file  
Needed by Norman and Anders   

#### c1. Pick years  
```{r}
from_year <- max(data_all2$MYEAR) - 4
to_year <- max(data_all2$MYEAR)

data_export <- data_all2 %>%
  filter(MYEAR %in% from_year:to_year)

cat("'data_export':", nrow(data_export), "rows")

```


#### c2. Add trophic level data  
```{r}
data_trophiclevel <- readRDS("Input_data/104_dat3.rds") 

cat("'data_trophiclevel' contains muscle only: \n")
xtabs(~is.na(TL) + TISSUE_NAME, data_trophiclevel)

data_export <- data_export %>%
  safe_left_join(data_trophiclevel %>%
                   rename(Fish_number = SAMPLE_NO2) %>%
                   select(MYEAR, STATION_CODE, Fish_number, Delta15N, TL),
                 by = c("MYEAR", "STATION_CODE", "Fish_number"),
                 na_matches = "never",
                 check = "bCV")

cat("\n\n")
cat("Final data: trophic level (TL) for most of muscle data and some of liver data: \n")
xtabs(~is.na(TL) + TISSUE_NAME, data_export %>% filter(MYEAR >= from_year))

```

#### c3. Change tissue names  
```{r}

if (FALSE){
  
  # Check SH and WO parameters

  table(data_export$TISSUE_NAME)
  
  data_export %>% filter(TISSUE_NAME == "SH") %>% xtabs(~MYEAR + PARAM, .)
  data_export %>% filter(TISSUE_NAME == "SH") %>% xtabs(~STATION_CODE + PARAM, .)

  data_export %>% filter(TISSUE_NAME == "WO") %>% xtabs(~MYEAR + PARAM, .)
  data_export %>% filter(TISSUE_NAME == "WO") %>% xtabs(~STATION_CODE + PARAM, .)
  
  data_export %>% 
    filter(TISSUE_NAME == "WO" & PARAM %in% "IMPS") %>%    # imposex on an individual level
    ggplot(aes(MYEAR, VALUE_WW)) +
    geom_jitter(width = 0.2, height = 0.1) +
    facet_wrap(vars(STATION_CODE))

  data_export %>% 
    filter(PARAM %in% "VDSI") %>%
    ggplot(aes(MYEAR, VALUE_WW)) +
    geom_point() +
    facet_wrap(vars(STATION_CODE))

  data_export %>% 
    filter(TISSUE_NAME == "WO" & PARAM %in% "RSPI") %>%
    ggplot(aes(MYEAR, VALUE_WW)) +
    geom_point() +
    facet_wrap(vars(STATION_CODE))

  }

# - delete "sample" and "TISSUE_NAME == 'WO' and 'SH'"
# - change TISSUE_NAME into English versions for the rest
data_export <- data_export %>%
  filter(!PARAM %in% "Sample") %>%
  filter(!TISSUE_NAME %in% c('SH', 'WO')) %>%
  mutate(TISSUE_NAME = case_when(
    TISSUE_NAME %in% "Lever" ~ "Liver",
    TISSUE_NAME %in% "Muskel" ~ "Muscle",
    TISSUE_NAME %in% "Blod" ~ "Blood",
    TISSUE_NAME %in% "Galle" ~ "Bile",
    TRUE ~ TISSUE_NAME)
  )

cat("TISSUE_NAME values after change: \n")
table(data_export$TISSUE_NAME)

```

#### c4. Pick data and save  
```{r}

# 2015-selected_year   
data_last_years <- data_export %>%
  select(-c(VALUE_DW, VALUE_FB, VALUE_WWa, VALUE_DWa, VALUE_FBa)) 

fn <- paste0(
  "Data/Milkys_raw_data_", 
  from_year, "-", substr(to_year, 3, 4),
  " (script 109).xlsx")

data_last_years %>%
  writexl::write_xlsx(fn)

cat("\nData written to \n")
cat(" ", fn, "\n")

```


