---
title: "102_Medians_and PROREF"
output: html_document
---

## 1. Packages
```{r, message=FALSE, warning=FALSE, results='hide'}

install.packages("dtplyr")

library(readxl)
library(tidyr)
library(data.table)
library(dtplyr)
library(dplyr, warn.conflicts = FALSE)
library(purrr)


```



## 2. Data

### Main data  
Read and reformat the most recent data (by default)  
```{r}

files <- dir("Data", pattern = "101_data_updated") %>% rev()

cat("Reading the last file downloaded:")
cat("\n", files[1])
cat("\n")
cat("If you want to read a different file, replace 'files[1]' with the file you want")
cat("\n")

filename <- paste0("Data/", files[1]) 
data_all <- readRDS(filename)

# We save the date part of the text (e.g., '2020-04-23')
# This will be used in part 10, when we save the resulting file
file_date <- substr(filename, 18, 27)    # pick out the part of the text from character no. 17 to no. 26

```

### Fixing 
Fixing station 227G2 (shall count as 227G) 
```{r}

sel <- data_all$STATION_CODE %in% "227G2"; sum(sel)
# xtabs(~MYEAR, data_all[sel,])
data_all$STATION_CODE[sel] <- "227G"

```


### Proref values  
We do not update Proref values, as we don't want to have a "moving target". However, a selection of wet-weight Proref values were re-estimated for the Proref paper, so we need to use these. Thus we use a combination of   
- proref_old, which are the Proref values used in the report for the 2017 data - is used for all parameter/basis combinattions that are not in proref_paper  
- proref_paper, which are the Proref values used in the Proref paper (a selection of parameters, and wet weigth only)  
These are combined to proref_updated02, which added to the data in part 4  
```{r}

# Old ones 
proref_old <- read_excel("Input_data/Proref_report_2017.xlsx")
proref_paper <- read_excel("Input_data/Proref_paper.xlsx")

# proref_updated01 = old data set, with selected columns,
#   and removing the rows that were produced for the paper (proref_ww)
proref_updated01 <- proref_old %>%
  # Select columns to keep
  select(PARAM, LATIN_NAME, TISSUE_NAME, Basis, Stations, N_stations, N, Median, Q95) %>%
  # Pick rows that are *not* in proref_paper
  anti_join(proref_paper %>% select(PARAM, LATIN_NAME, TISSUE_NAME, Basis),
            by = c("PARAM", "LATIN_NAME", "TISSUE_NAME", "Basis")
            )

# proref_updated02 - adding the rows from proref_paper
proref_updated02 <- proref_updated01 %>%
  bind_rows(
    proref_paper %>%
      select(PARAM, LATIN_NAME, TISSUE_NAME, Basis, Stations, N_stations, N, Median, Q95)
  )

```

### Check Proref  
```{r}

check <- proref_updated02 %>%
  count(PARAM, LATIN_NAME, TISSUE_NAME, Basis)

if (sum(check$n > 1) == 0){
  cat("All OK - only one Proref value per PARAM, LATIN_NAME, TISSUE_NAME, Basis combination  \n")
} else {
  cat("Warning: more than one Proref value per PARAM, LATIN_NAME, TISSUE_NAME, Basis combination. Must be fixed!  \n")
}

```

```{r}
data_all[1:20,]$STATION_CODE
```


## 3. Summarise data for each year/station/parameter  

### Calculate medians using dtplyr  
```{r}

data_all_dt <- lazy_dt(data_all)

t0 <- Sys.time()

data_med_1a <- data_all_dt %>% 
  group_by(MYEAR, STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM, UNIT) %>% 
  summarise_at(vars(DRYWT, FAT_PERC, VALUE_WW, VALUE_DW, VALUE_FB, VALUE_WWa, VALUE_DWa, VALUE_FBa), 
               median, 
               na.rm = TRUE) %>%
  as_tibble()

Sys.time() - t0  # 12 secs (took 0.19 sec with data.table)

```


### Calculate N, Det_limit and Over_LOQ
```{r}

t0 <- Sys.time()

data_med_1b <- data_all_dt %>%
  group_by(MYEAR, STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM, UNIT) %>%
  summarise(N = n(), Det_limit = median(VALUE_WW[!is.na(FLAG1)]), Over_LOQ = sum(is.na(FLAG1))) %>%
  as_tibble()

Sys.time() - t0  # 1.4 secs


```

### Combine data_med_1 and data_med_2    
Add N, Det_limit and Over_LOQ to the data set with medians  
```{r}

# Make sure the data sets are sorted the same way
data_med_1a <- data_med_1a %>% arrange(MYEAR, STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM, UNIT)
data_med_1b <- data_med_1b %>% arrange(MYEAR, STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM, UNIT)

# Make sure the data sets have the same number of rows
if (nrow(data_med_1a) == nrow(data_med_1b)){
  # Join data by adding the N, Det_limit, Over_LOQ columns 
  data_med_1 <- bind_cols(
    data_med_1a,
    data_med_1b %>% select(N, Det_limit, Over_LOQ)
  )
} else {
  cat("data_med_1a and data_med_1b have different number of rows! Must be fixed, or use left_join(). \n")
}

```

## 4. Add Proref  

### Add "VDSI/Intersex" to PROREF data 
```{r}

# Pick one VDSI line and change it
sel <- proref_updated02$PARAM == "VDSI" & proref_updated02$Basis == "WW"; sum(sel)
proref_to_add <- proref_updated02[sel,]
proref_to_add$PARAM <- "VDSI/Intersex"

# Add 
proref_updated03 <- bind_rows(
  proref_updated02,
  proref_to_add
)

```

### Add PROREF data to medians ----
```{r}

data_med_2 <- data_med_1 %>%
  pivot_longer(cols = c(VALUE_WW, VALUE_DW, VALUE_FB, VALUE_WWa, VALUE_DWa, VALUE_FBa), names_to = "Basis", values_to = "Value") %>%
  mutate(Basis = stringr::str_sub(Basis, start = 7))
nrow(data_med_2)   # 429432

data_med_3 <- data_med_2 %>%
  left_join(proref_updated03 %>% select(-N), by = c("LATIN_NAME", "TISSUE_NAME", "PARAM", "Basis"))

nrow(data_med_3)  # 448680

```

## 5. Save  
```{r}

filename <- paste0("Data/110_mediandata_updated_", file_date)

saveRDS(data_med_3, filename)

```




