---
title: "118_Compare_files"
output: 
  html_document:
    keep_md: true
    
---

Comparing median files of different versions    
  
The purpose of this script is to cut down on time used to run time trend analyses (script 120) 
by (re)analysing only the data series that have changed since the last run.  
  
The script
* reads both the newest median data as well as the median data used in the last run of script 120 (by default, the second last median data); see section 'Read files: Medians (3a)'. Both of these data files are the result of script 110.  
* filters the data, so as to only keep data that are used in time series analysis  (3c)
* finds data series that are either new or longer (4a), or that have changes in the data (4b)  
  
This script produces `data_series_reanalyse`and `data_series_remove` which must be **in memory** for script 120 to work.  


## 1. Set 'series_lastyear'   
```{r}

series_lastyear <- 2018   # The series must last at least until this year    
                   # Series without data in any of the three last years will be excluded

```


## 2. Packages   
```{r}

library(dplyr)
library(ggplot2)
library(readxl)
source("001_Add_trends_functions.R")  # 

# Tried this:
# library(compareDF )
# - not needed. And it's quite slow.

```



## 3. Read files  

### a. Medians  

* The two versions we want to compare  

```{r, collapse=TRUE}

filepattern <- "110_mediandata_updated_"         # entire file name except date and extension

files <- dir("Data", pattern = filepattern) %>% rev()

#
# The following file should be the last existing file with medians (product of script 110):
#
# (typically filenumber = 1, i.e. the last file made)
#
filenumber <- 1 
data_list <- read_rds_file("Data",
                     files, 
                     filenumber = filenumber,   # "1" gets the newest file   
                     get_date = TRUE, time_since_modified = TRUE)
data_med_all_1 <- data_list$data%>%
  rename(Proref_median = Median,
         Median = Value) 


#
# The following file should be the file that the last trends (product of script 120) 
#    are based on. I.e., the file used the last time you ran script 120.
#
# (typically filenumber = 2, i.e. the second last file made)
#
filenumber <- 2 
data_list <- read_rds_file("Data",
                     files, 
                     filenumber = filenumber,   # "1" gets the newest file   
                     get_date = TRUE, time_since_modified = TRUE)
data_med_all_2 <- data_list$data%>%
  rename(Proref_median = Median,
         Median = Value) 

```
### b. Data series   

* One row per parameter 

```{r}

data_series1 <- get_trend_series(data_med_all_1, 
                                 parameters = get_trend_parameters(), 
                                 series_lasting_to = series_lastyear)

data_series2 <- get_trend_series(data_med_all_2, 
                                 parameters = get_trend_parameters(), 
                                 series_lasting_to = series_lastyear)  


```

### c. Filter median data  

* Keep only data found in 'data_series1', 'data_series2'
* I.e., keep only data in series long enough to be time series, and lasting until 'lastyear'    

```{r}

data_med1 <- data_med_all_1 %>% 
    semi_join(data_series1,
              by = c("STATION_CODE", "LATIN_NAME", "TISSUE_NAME", "PARAM"))

data_med2 <- data_med_all_2 %>% 
    semi_join(data_series2,
              by = c("STATION_CODE", "LATIN_NAME", "TISSUE_NAME", "PARAM"))


```



## 4. Compare data  

* Note: `data_med2`is the oldest one  

### a. Find data series of different length  

* Includes data series that are added or removed  

```{r}

data_series_length <- full_join(
  data_med1 %>%
    count(STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM, name = "n1"), 
  data_med2 %>%
    count(STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM, name = "n2"),
  by = c("STATION_CODE", "LATIN_NAME", "TISSUE_NAME", "PARAM")
) %>%
  mutate(
    n1 = ifelse(is.na(n1), 0, n1),
    n2 = ifelse(is.na(n2), 0, n2)
  )

data_series_different_length <- data_series_length %>%
  filter(n1 != n2)

data_series_different_length

```

###  b. Compare data series of same length   

#### Prepare data sets for comparison  

* Pick only data series of same length  
* Order rows identically  

```{r}

data_med1_samelength <- data_med1 %>%
  left_join(data_series_different_length, 
            by = c( "STATION_CODE", "LATIN_NAME", "TISSUE_NAME", "PARAM")) %>%
  filter(is.na(n1)) %>%
  select(STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM, MYEAR, Median, N_median, Over_LOQ) %>%
  arrange(STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM, MYEAR)

data_med2_samelength <- data_med2 %>%
  left_join(data_series_different_length, 
            by = c( "STATION_CODE", "LATIN_NAME", "TISSUE_NAME", "PARAM")) %>%
  filter(is.na(n1)) %>%
  select(STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM, MYEAR, Median, N_median, Over_LOQ) %>%
  arrange(STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM, MYEAR)

# Check length
if (nrow(data_med1_samelength) != nrow(data_med2_samelength) ){
  stop("Data should have the same number of rows, but don't !")
}

#
# Make sure all key variables (STATION_CODE, LATIN_NAME, etc.) are the same
#

# For column 1-5, check taht all valeus are identical
key_variables_differences <- 1:5 %>% 
  purrr::map_int(
    ~sum(data_med1_samelength[,.x] != data_med2_samelength[,.x])
  )

if (sum(key_variables_differences) > 0){
  cat("Variable number", 
      which(key_variables_differences > 0), "are different")
  stop("Some of the key variables are different!")
} else {
  cat("All key variables are equal")
}

```


#### Combine series   

```{r}

# Combine using bind_cols
data_med_compare <- bind_cols(
  data_med1_samelength %>%
    select(STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM, MYEAR, Median, N_median, Over_LOQ) %>%
    rename(Median_1 = Median, N_median_1 = N_median, Over_LOQ_1 = Over_LOQ),
  data_med2_samelength %>%
    select(Median, N_median, Over_LOQ) %>%
    rename(Median_2 = Median, N_median_2 = N_median, Over_LOQ_2 = Over_LOQ)
) %>%
  select(STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM, MYEAR, 
         Median_1, Median_2, N_median_1, N_median_2, Over_LOQ_1, Over_LOQ_2)

sel_a1 <- with(data_med_compare, abs(Median_1 - Median_2) > 0.0001 & !is.na(Median_1) & !is.na(Median_2))
sel_a2 <- with(data_med_compare, !is.na(Median_1) & is.na(Median_2))
sel_a3 <- with(data_med_compare, is.na(Median_1) & !is.na(Median_2))
sel_b <- with(data_med_compare, N_median_1 != N_median_2)
sel_c <- with(data_med_compare, Over_LOQ_1 != Over_LOQ_2 )

cat("Different 'Median'", sum(sel_a1), "\n")
cat("Data for Median_1, lacks Median_2:", sum(sel_a2), "\n")
cat("Data for Median_1, lacks Median_1:", sum(sel_a3), "\n")
cat("Different 'N_median'", sum(sel_b), "\n")
cat("Different 'Over_LOQ'", sum(sel_c), "\n")

sel_combined <- sel_a1 | sel_a2 | sel_a3 | sel_b | sel_c
cat("Difference in one or more columns: ", sum(sel_combined), "\n")

```

#### Get data series with different values  
```{r}

# View differences
if (FALSE)
  View(data_med_compare[sel_combined,])

data_series_different_values <- data_med_compare[sel_combined,] %>%
  distinct(STATION_CODE, LATIN_NAME, TISSUE_NAME, PARAM) 

xtabs(~PARAM + STATION_CODE, data_series_different_values)
xtabs(~MYEAR + STATION_CODE, data_med_compare[sel_combined,])


```

### c. Final series to use in script 120    

#### Data series to reanalyse  

```{r}

data_series_reanalyse <- bind_rows(
  data_series_different_length %>% filter(n1 > 0),
  data_series_different_values)

```

#### Data series to remove  

```{r}

data_series_remove <- data_series_different_length %>% filter(n1 == 0)

```


## 5. Remove biggish data from memory  
```{r}

if (FALSE)
  rm(data_list,
     data_med_all_1, data_med_all_2,
     data_med1, data_med2,
     data_med1_samelength, data_med2_samelength,
     data_med_compare)

```

