---
title: "121_Check_trends"
output: 
  html_document:
    keep_md: true
    toc: true  
    toc_float: true
    
---

## 0. Year  
```{r}

last_year <- 2019

```


## 1. Libraries and functions  
```{r, results='hide', message=FALSE, warning=FALSE, echo=FALSE}

# install.packages("lubridate")

library(dplyr)
library(purrr)
library(lubridate)
library(readxl)
library(mgcv)
library(AICcmodavg)   # AICc()
library(ggplot2)
library(safejoin)     # installed from https://github.com/moodymudskipper/safejoin   
source("001_Add_trends_functions.R")  # Copied from '16_Trend_functions.R' in Milkys_2018

#
# Define a couple of extra functions for shortening code
#
get_stats <- function(df){
  calc_models_one_station2(df, gam = TRUE, log = TRUE)$statistics_for_file
}

# model_from_medians - get statistic summary only
model_from_medians_stat <- function(...)
  model_from_medians(...)$statistics

# PLotting graph only
model_from_medians_graph <- function(...){
  invisible(model_from_medians(...))
}


```

## 2. Get data    

### a. Medians per year   
```{r, echo = FALSE, results='hold'}
# Medians
# Read and reformat the most recent data (by default)  
files <- dir("Data", pattern = "110_mediandata_updated") %>% rev()

cat("Reading the last file downloaded:")
cat("\n", files[1])
cat("\n")
cat("If you want to read a different file, replace 'files[1]' with the file you want")
cat("\n\n")

filename <- files[1]   
cat("Time since this file was modified: \n")
Sys.time() - file.info(paste0("Data/", filename))$mtime

data_med <- readRDS(paste0("Data/", filename)) %>%
  rename(Proref_median = Median,
         Median = Value) 

# We save the date part of the text (e.g., '2020-04-23')
# This will be used in part 10, when we save the resulting file
file_date <- substr(filename, 24, 33)    # pick out the part of the text from character no. 17 to no. 26

```


### b. Calculated time trend results        
```{r, echo=FALSE, results='hold'}

date_pattern <- "[0-9][0-9][0-9][0-9]-[0-9][0-9]-[0-9][0-9]"
version_pattern <- "run[0-9][0-9]"
pattern_10yr <- paste0("^120_result_10yr_", last_year, "_", 
                       date_pattern, "_", version_pattern, ".rds$")
pattern_long <- paste0("^120_result_long_", last_year, "_", 
                       date_pattern, "_", version_pattern, ".rds$")

# Find files and sort with the newest first
files1 <- dir("Data", pattern = pattern_10yr) %>% rev()
files2 <- dir("Data", pattern = pattern_long) %>% rev()

if (FALSE){
  cat("Latest files: \n")
  cat(files1[1], "\n")
  cat(files2[1], "\n")
}

# Print for user
cat("Trend estimates - reading the last files produced:")
cat("\n ", files1[1])
cat("\n ", files2[1])

# Read files
result_10yr <- readRDS(paste0("Data/", files1[1]))
result_long <- readRDS(paste0("Data/", files2[1]))

# Set as factor  
levels_status <- c("GAM OK", "GAM failed", "Linear regr. and GAM failed", "No variation in data")
result_10yr <- result_10yr %>% mutate(Status = factor(Status, levels = levels_status))
result_long <- result_long %>% mutate(Status = factor(Status, levels = levels_status))

```


## 3. Time trends calculated - tables  

### Calculation of time trends   
```{r, echo = FALSE}

left_join(result_long %>% count(Status) %>% rename(`Long-term trend` = n),
          result_10yr %>% count(Status) %>% rename(`10-year trend` = n),
          by = c("Status")
) %>%
  knitr::kable()
```


### Time trend results  
```{r, echo = FALSE}

left_join(result_long %>% 
            filter(Status == "GAM OK") %>%
            mutate(Dir_change = ifelse(Dir_change=="", "No trend", Dir_change)) %>%
            count(Model_used, Dir_change) %>% rename(`Long-term trend` = n),
          result_10yr %>% 
            filter(Status == "GAM OK") %>%
            mutate(Dir_change = ifelse(Dir_change=="", "No trend", Dir_change)) %>%
            count(Model_used, Dir_change) %>% rename(`10-year trend` = n),
          by = c("Model_used", "Dir_change")) %>%
  knitr::kable()

```

## 4. Results where linear and non-linear results differs  

**Checking p-values**   

* E.g., 'linear p > 0.20, non-lin p < 0.05' indicates that the *non-linear* model a quite signinficant, while in contrast the *linear* model was quite 'non-signinficant'     
  
```{r, echo = FALSE, results='hide'}

result_deviance_nonlinear_long <- result_long %>% 
  filter(Model_used %in% c("Linear","Nonlinear") & Basis %in% c("WW","WWa") & Year1 < 2010) %>%
  filter(p_nonlinear < 0.05 & p_linear > 0.2) 
# cat("Long-term trend: linear p > 0.20, non-lin p < 0.05:",
#     nrow(result_deviance_nonlinear_long), "cases \n")

result_deviance_linear_long <- result_long %>% 
  filter(Model_used %in% c("Linear","Nonlinear") & Basis %in% c("WW","WWa") & Year1 < 2010) %>%
  filter(p_linear < 0.05 & p_nonlinear > 0.2) 
# cat("Long-term trend: linear p < 0.05, non-lin p > 0.20:",
#     nrow(result_deviance_linear_long), "cases \n")

result_deviance_nonlinear_10yr <- result_10yr %>% 
  filter(Model_used %in% c("Linear","Nonlinear") & Basis %in% c("WW","WWa")) %>%
  filter(p_nonlinear < 0.05 & p_linear > 0.2) 
# cat("10-year trend: linear p > 0.20, non-lin p < 0.05:",
#     nrow(result_deviance_nonlinear_10yr), "cases \n")

result_deviance_linear_10yr <- result_10yr %>% 
  filter(Model_used %in% c("Linear","Nonlinear") & Basis %in% c("WW","WWa")) %>%
  filter(p_linear < 0.05 & p_nonlinear > 0.2) 
# cat("10-year trend: linear p < 0.05, non-lin p > 0.20:",
#     nrow(result_deviance_linear_10yr), "cases \n")

```

### Long-term trend: linear p > 0.20, non-linear p < 0.05  
```{r, echo = FALSE, warning = FALSE, fig.width=6, fig.height=4}

for (i in 1:nrow(result_deviance_nonlinear_long)){
  
  stat <- with(result_deviance_nonlinear_long,
       model_from_medians_stat(
         PARAM[i], LATIN_NAME[i], TISSUE_NAME[i], 
         STATION_CODE[i], Basis[i], 1980:2019, data_med)
  )
  
  if (stat$p_linear > 0.2 & stat$p_nonlinear < 0.05){
    
    with(result_deviance_nonlinear_long,
         model_from_medians_graph(
           PARAM[i], LATIN_NAME[i], TISSUE_NAME[i], 
           STATION_CODE[i], Basis[i], 1980:2019, data_med, 
           plotname = "window", ggplot = TRUE)
    )
  }

}

```

### Long-term trend: linear p < 0.05, non-linear p > 0.20    
```{r, echo = FALSE, warning = FALSE, fig.width=6, fig.height=4}

for (i in 1:nrow(result_deviance_linear_long)){
  
  stat <- with(result_deviance_linear_long,
       model_from_medians_stat(
         PARAM[i], LATIN_NAME[i], TISSUE_NAME[i], 
         STATION_CODE[i], Basis[i], 1980:2019, data_med)
  )
  
  if (stat$p_linear < 0.05 & stat$p_nonlinear > 0.2){
    
    with(result_deviance_linear_long,
         model_from_medians_graph(
           PARAM[i], LATIN_NAME[i], TISSUE_NAME[i], 
           STATION_CODE[i], Basis[i], 1980:2019, data_med, 
           plotname = "window", ggplot = TRUE)
    )
  }
  
}

```

### 10-year trend: linear p > 0.20, non-linear p < 0.05  
```{r, echo = FALSE, warning = FALSE, fig.width=6, fig.height=4}

for (i in 1:nrow(result_deviance_nonlinear_10yr)){
  
  stat <- with(result_deviance_nonlinear_10yr,
       model_from_medians_stat(
         PARAM[i], LATIN_NAME[i], TISSUE_NAME[i], 
         STATION_CODE[i], Basis[i], 2010:2019, data_med)
  )
  
  if (stat$p_linear > 0.2 & stat$p_nonlinear < 0.05){
    
    with(result_deviance_nonlinear_10yr,
         model_from_medians_graph(
           PARAM[i], LATIN_NAME[i], TISSUE_NAME[i], 
           STATION_CODE[i], Basis[i], 2010:2019, data_med, 
           plotname = "window", ggplot = TRUE)
    )
  }
  
}

```

### 10-year trend: linear p < 0.05, non-linear p > 0.20    
```{r, echo = FALSE, warning = FALSE, fig.width=6, fig.height=4}

for (i in 1:nrow(result_deviance_linear_10yr)){
  
  stat <- with(result_deviance_linear_10yr,
       model_from_medians_stat(
         PARAM[i], LATIN_NAME[i], TISSUE_NAME[i], 
         STATION_CODE[i], Basis[i], 2010:2019, data_med)
  )
  
  if (stat$p_linear < 0.05 & stat$p_nonlinear > 0.20){
    
    with(result_deviance_linear_10yr,
         model_from_medians_graph(
           PARAM[i], LATIN_NAME[i], TISSUE_NAME[i], 
           STATION_CODE[i], Basis[i], 2010:2019, data_med, 
           plotname = "window", ggplot = TRUE)
    )
  }
  
}

```




