---
title: "125 Calculate time trends using pakage leftadjusted"
output: html_document
---

Note: Ca 35 minutes run time for running all trends    

This procedure uses - at this point - data from script 105.  
- Should finally use data from script 109  


## 0. Constants (update each year)     
```{r}

last_year <- 2021

series_lastyear <- 2018   # The series must last at least until this year    
                          # Series without data in any of the three last years will be excluded


```
  
## 1. Libraries and functions  
```{r, results='hide', message=FALSE, warning=FALSE}

# install.packages("lubridate")

# General purpose
library(dplyr)
library(tidyr)
library(purrr)
library(mgcv)
library(ggplot2)

# Specific for the analysis
library(rjags)
library(runjags)
devtools::install_github("DagHjermann/leftcensored")
library(leftcensored)

# For parallel computing
install.packages("doParallel")
library(doParallel)

# Load functions defined in other scripts  
source("110_Medians_and_PROREF_functions.R")  # for homogenize_series
source("002_Utility_functions.R")

```

## 2. Read data  

### a. Main data  
Read and reformat the most recent data (by default)  
```{r, collapse=TRUE}

# If we have NOT length-adjusted the last year's data:
filepattern <- "105_data_with_uncertainty_"         # entire file name except date and extension

# Normally, if we have length-adjusted the last year's data:
# filepattern <- "109_adjusted_data_"       # entire file name except date and extension

filenumber <- 1                           # filenumber = 1 means "read the newest file"

files <- dir("Data", pattern = filepattern) %>% rev()

data_list <- read_rds_file("Data",
                     files, 
                     filenumber = filenumber,   # "1" gets the newest file   
                     get_date = TRUE, time_since_modified = TRUE)

dat_all <- data_list$data

# The file_date text will be used in part 10, when we save the resulting file
cat("File date text:", data_list$file_date, "\n")

```


### b. Homogenize time series  
Change STATION_CODE, in order to make a single time series for data with different STATION_CODE that in practice should count as the same station   
* Fixing station 227G1 and 227G2 (shall count as 227G)  
* Fixing station 36A and 36A1 (36A1 shall count as 36A)  
  
Also combines PARAM = VDSI and PARAM = Intersex to PARAM = "VDSI/Intersex" for station 71G  

```{r}

dat_all <- homogenize_series(dat_all)

```

### c. Add N_years, N_years_plus, Trend_model and k_max    
```{r}

dat_all <- dat_all %>%
  group_by(PARAM, STATION_CODE, TISSUE_NAME, MYEAR) %>%
  mutate(
    N = n(),
    N_over_LOQ = sum(is.na(FLAG1)),
    P_over_LOQ = N_over_LOQ/N) %>%
  group_by(PARAM, STATION_CODE, TISSUE_NAME) %>%
  mutate(
    N_years = length(unique(MYEAR)), 
    N_years_plus = length(unique(MYEAR[N_over_LOQ >= 1])),
    Last_year = max(MYEAR)
    )  %>%
  ungroup() %>%
  mutate(
    Trend_model = case_when(
      N_years_plus <= 1 ~ "No model",
      N_years_plus %in% 2 & N_years %in% 2 ~ "No model",
      N_years_plus %in% 2:4 & N_years >= 3 ~ "Mean",
      N_years_plus %in% 5:6 ~ "Linear",
      N_years_plus %in% 7:9 ~ "Smooth, k_max=3",
      N_years_plus %in% 10:14 ~ "Smooth, k_max=4",
      N_years_plus >= 15 ~ "Smooth, k_max=5"),
    Trend_model = factor(
      Trend_model,
      levels = c("No model", "Mean", "Linear", 
                 "Smooth, k_max=3", "Smooth, k_max=4", "Smooth, k_max=5")),
    k_max = case_when(
      Trend_model %in% "No model" ~ as.numeric(NA),
      Trend_model %in% "Mean" ~ 1,
      Trend_model %in% "Linear" ~ 2,
      Trend_model %in% "Smooth, k_max=3" ~ 3,
      Trend_model %in% "Smooth, k_max=4" ~ 4,
      Trend_model %in% "Smooth, k_max=5" ~ 5)
    )


```


### d. Statistics per group  
```{r}

dat_all %>%
  filter(!is.na(UNCERTAINTY) & Last_year %in% last_year) %>%
  mutate(
    Substance.Group = substr(Substance.Group, 1,14)) %>%
  distinct(Substance.Group, PARAM, STATION_CODE, TISSUE_NAME, Trend_model) %>%
  count(Substance.Group, Trend_model) %>%
  pivot_wider(Substance.Group, names_from = Trend_model, values_from = n,
              names_sort= TRUE)

dat_all %>%
  filter(!is.na(UNCERTAINTY) & Last_year %in% last_year) %>%
  mutate(
    Substance.Group = substr(Substance.Group, 1,14),
    P_over_LOQ = cut(P_over_LOQ, breaks = seq(0,1,0.2), 
                     include.lowest = TRUE)) %>%
  distinct(Substance.Group, PARAM, STATION_CODE, TISSUE_NAME, P_over_LOQ) %>%
  count(Substance.Group, P_over_LOQ) %>%
  pivot_wider(Substance.Group, names_from = P_over_LOQ, values_from = n,
              names_sort= TRUE)

if (FALSE){
  
  dat_all %>% xtabs(~ addNA(UNCERTAINTY) + MYEAR, .)
  dat_all %>% xtabs(~ is.na(UNCERTAINTY) + MYEAR, .)
  
  dat_all %>%
  filter(!is.na(UNCERTAINTY) & UNCERTAINTY > 0 & Last_year %in% last_year) %>%
  mutate(
    Substance.Group = substr(Substance.Group, 1,14),
    PARAM = substr(PARAM, 1,14),
    P_over_LOQ = cut(P_over_LOQ, breaks = seq(0,1,0.2), 
                     include.lowest = TRUE)) %>%
  distinct(Substance.Group, PARAM, STATION_CODE, TISSUE_NAME, P_over_LOQ) %>%
  count(Substance.Group, PARAM, P_over_LOQ) %>%
  pivot_wider(c(Substance.Group, PARAM), names_from = P_over_LOQ, values_from = n,
              names_sort= TRUE)

}
```

## 3. Trend analysis - tests  

### One parameter, several stations - plot  
```{r, fig.width=9, fig.height=7}

param <- "BDE99"
param <- "CB28"
dat_all %>%
  filter(PARAM == param & Last_year %in% last_year) %>%
  ggplot(aes(MYEAR, VALUE_WW, color = is.na(FLAG1))) +
  geom_point() + 
  scale_y_log10() +
  facet_wrap(vars(STATION_CODE))

df <- dat_all %>%
  filter(PARAM == param & Last_year %in% last_year) %>%
  group_by(STATION_CODE, TISSUE_NAME) %>%
  summarise(Trend_model = first(Trend_model))

df
table(df)

```

### One parameter, one station    
```{r, fig.width=7, fig.height=4}

param <- "CB28"
station <- "36A"
data_test_prep <- lc_prepare(subset(dat_all, PARAM %in% param & STATION_CODE %in% station),
                             x = "MYEAR",
                             y = "VALUE_WW", 
                             censored = "FLAG1",
                             log = TRUE,
                             keep_original_columns = TRUE)

lc_plot(data_test_prep)
```


### Test a single model (k = 3)
```{r, fig.width=7, fig.height=4}

# Entire series
model_test1 <- leftcensored::lc_fixedsplines_tp(
  data = data_test_prep, 
  normalize = FALSE, k = 3, raftery = FALSE, measurement_error = "Uncertainty", 
  predict_x = seq(min(data_test_prep$x), max(data_test_prep$x)), 
  compare_with_last = TRUE)

lc_plot(data_test_prep, results = model_test1)

# Last 12 years
model_test2 <- leftcensored::lc_fixedsplines_tp(
  data = data_test_prep %>% filter(MYEAR >= 2009), 
  normalize = FALSE, k = 3, raftery = FALSE, measurement_error = "Uncertainty", 
  predict_x = seq(2009, 2021), 
  compare_with_last = TRUE)

lc_plot(data_test_prep, results = model_test2)

ggplot(model_test2$diff_data, aes(x, y)) +
  geom_ribbon(aes(ymin = y_lo, ymax = y_hi), fill = "lightblue") + 
  geom_line()

```


## 4. Test function `get_splines_results`  

```{r, fig.width=7, fig.height=4}

# data must contain variables
#   x, y, threshold, uncensored
#   k_max

param <- "CB28"
station <- "36A"; tissue = "Whole soft body"
station <- "10A2"; tissue = "Whole soft body"   # mean
station <- "19N"; tissue = "Egg"                 # linear

# One time only: create new folder for the results  
# dir.create("Data/124_results")

# debugonce(get_splines_results)
test <- get_splines_results(
  subset(dat_all, PARAM %in% param & STATION_CODE %in% station & TISSUE_NAME %in% tissue), 
  raftery = FALSE, save_path = "Data/124_results/test1.rda")

```

## 5. Test analysing and saving results    

- We save results as R files as we go, in case something goes wrong  

### Data series to run  
```{r}

dat_series <- dat_all %>%
  filter(
    !Trend_model %in% c("No model", "Mean"),
    Last_year %in% last_year) %>%
  distinct(PARAM, STATION_CODE, TISSUE_NAME, Trend_model, k_max)

```


### Example data series    
```{r}

set.seed(21)
dat_series <- sample_n(dat_series, 2)

```

### Test  `get_splines_results_seriesno_s`  

- Function that works on series number i  
- Also saves results as a separate rda file   

```{r, results='hide'}

# test
test <- get_splines_results_seriesno_s(2, dat_all, dat_series)

```


### Run analyses using doParallel  

```{r, results='hide'}

if (FALSE){
  
  # Do only once
  future::availableCores()
  # 4
  cl <- makeCluster(4)
  registerDoParallel(cores = 4)
  
}


#
# Parallel (for n series)
#

t0 <- Sys.time()
result <- foreach(i = 1:2, 
                  .export = c("dat_all","dat_series")) %dopar%
  get_splines_results_seriesno_s(i, dat_all, dat_series, raftery = FALSE)
t1 <- Sys.time()
t1-t0
# 52.43795 secs  


```

