---
title: "126_add_lengthadjusted"
author: "DHJ"
date: "2022-09-28"
output: html_document
---


```{r}

last_year <- 2021

# 04 is the run for all WW data, used in the big excel, before JAGS model code was updated  
# 05 is a later run for all WW data, after JAGS model code was updated
# 06 is also later, but for WWa (length-adjusted) HG (12 series only)


indexvars <- c("PARAM", "STATION_CODE", "TISSUE_NAME", "LATIN_NAME")

```

  
## 1. Libraries and functions  

```{r, results='hide', message=FALSE, warning=FALSE}

# install.packages("lubridate")

# General purpose
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)

source("125_Calculate_trends_leftadjusted_functions.R")

```


## Read results  

```{r}

df_trend_wide_04 <- readRDS("Data/125_results_2021_04_output/126_df_trend_wide_2021.rds")
df_trend_wide_05 <- readRDS("Data/125_results_2021_05_output/126_df_trend_wide_2021.rds")
df_trend_wide_06 <- readRDS("Data/125_results_2021_06_output/126_df_trend_wide_2021.rds") %>%
  mutate(Basis = "WWa")

```

## Combine results,  check changes from 04 to 05  

```{r}

# Combine (for comparison only)
df_trend_wide_04_compare <- df_trend_wide_04 %>%
  rename(Trend_string_long_04 = Trend_string_long, Trend_string_short_04 = Trend_string_short, Trend_symbol_04 = Trend_symbol) %>%
  left_join(
    df_trend_wide_05 %>% 
      select(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis, Trend_string_long, Trend_string_short, Trend_symbol) %>%
      rename(Trend_string_long_05 = Trend_string_long, Trend_string_short_05 = Trend_string_short, Trend_symbol_05 = Trend_symbol),
    by = c("PARAM", "STATION_CODE", "TISSUE_NAME", "LATIN_NAME", "Basis")  
  )

xtabs(~Trend_string_long_04 + Trend_string_long_05, df_trend_wide_04_compare)
xtabs(~Trend_string_short_04 + Trend_string_short_05, df_trend_wide_04_compare)


```

### Check "Estimation failed"  

```{r}

# df_index <- readRDS("Data/125_results_2021_04_input/125_dat_series_trend.rds")  

xtabs(~N_years_10yr +Trend_string_short_05, df_trend_wide_04_compare)

df_trend_wide_04_compare %>%
  mutate(N_years = cut(N_years, breaks = c(1:9, seq(10,40,5)), include.lowest = TRUE)) %>%
  xtabs(~N_years + Trend_string_long_05, .)



```

### Check 'Estimation failed' series for the longer time series   
```{r}

df_trend_wide_06 %>%
  filter(Trend_string_long == "Estimation failed" & N_years > 5)
  
df_trend_wide_04_compare %>%
  filter(Trend_string_long_05 == "Estimation failed" & N_years)


```


## Make 'df_trend_wide_04_updated'  

*-* We only update the cases where 04 = "Estimation failed"   

* 'df_trend_wide_05_to_replace' - new rows with results from 05  
* 'df_trend_wide_04_original' - old rows from 04  
* df_trend_wide_06 - Hg lengthadjusted  


```{r}

df_trend_wide_05_to_replace <- df_trend_wide_04 %>%
  filter(Trend_string_long %in% "Estimation failed" | Trend_string_short %in% "Estimation failed") %>%
  select(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis) %>%
  left_join(df_trend_wide_05)

df_trend_wide_04_original <- df_trend_wide_04 %>%
  anti_join(df_trend_wide_05_to_replace, 
            by = c("PARAM", "STATION_CODE", "TISSUE_NAME", "LATIN_NAME", "Basis"))

df_trend_wide_04_updated <- bind_rows(
  df_trend_wide_04_original,
  df_trend_wide_05_to_replace,
  df_trend_wide_06
)

```


### Check Trend_symbol  

```{r}

xtabs(~addNA(Trend_symbol), df_trend_wide_04_updated)

```
## Save  
```{r}

saveRDS(df_trend_wide_04_updated, "Data/125_results_2021_04_output/126_df_trend_wide_2021_updated_01.rds")


```

