---
title: "126_Extract_trends_leftadjusted"
author: "DHJ"
date: '2022-09-13'
output: html_document
---

- This is sensitive to changes in the saved sereise files ('dat_series' and 'dat_series_trend') on disk (created in script 125)  
- They should be built up from the results files (in the folder given by 'folder_results')  


## 0. Constants (update each year)     

```{r}

last_year <- 2021

#
# REMEMBER TO UPDATE!
#

# 04 is the run for all WW data, used in the big excel, before JAGS model code was updated  
# 05 is a later run for all WW data, after JAGS model code was updated
# 06 is also later, but for WWa (length-adjusted) HG (17 series only)
# 07 combines 05 and 06, and also adds 'Basis' to all relevant files ('series' data)
#   and to trend result files (see Appendix 2, script 125)

folder_results <- "Data/125_results_2021_07"         
folder_input <- paste0(folder_results, "_input")
folder_output <- paste0(folder_results, "_output")

if (!dir.exists(folder_output)){
  dir.create(folder_output)
}

indexvars <- c("PARAM", "STATION_CODE", "TISSUE_NAME", "LATIN_NAME", "Basis")

```

### Basis   

Not recorded in any way during the procedure of 125 - must be set manually!   
```{r}

# given_basis <- "WW"   # for folder 01-05 
# given_basis <- "WWa"  # for folder 06

```


  
## 1. Libraries and functions  

```{r, results='hide', message=FALSE, warning=FALSE}

# install.packages("lubridate")

# General purpose
library(dplyr)
library(tidyr)
library(purrr)
library(mgcv)    #  mgcv_1.8-39
library(ggplot2)

source("125_Calculate_trends_leftadjusted_functions.R")

```


## 2. Data  

### a. Main data and data series  

```{r, collapse=TRUE}

dat_series <- readRDS(paste0(folder_input, "/125_dat_series.rds"))
dat_series_trend <- readRDS(paste0(folder_input, "/125_dat_series_trend.rds"))
# dat_series_trend <- dat_series     # special case for folder 06
dat_all_prep3 <- readRDS(paste0(folder_input, "/125_dat_all_prep3.rds"))

```

### b. Make Rule1 / Rule2    

- Rule 1. Time series should be truncated from the left until Nplus/N >= 0.5     
- Rule 2. If a linear/smooth trend is fitted, the first year must be non-censored   
- Get the first value for both of them 
    - If Rule 1 or 2 is kicking in, it always effects the first value
    - if TRUE, the filtering has no effect  
- Are added to 'dat_success' below  

```{r}

df_rule12 <- dat_all_prep3 %>%
  group_by(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis) %>%
  summarise(Rule1 = first(Rule1), 
            Rule2 = first(Rule2), .groups = "drop") 

xtabs(~Rule1 + Rule2, df_rule12)

```


## 3. Check results  

### a1. Read result files  

```{r}

fns <- dir(folder_results, full.names = TRUE) %>% sort()
result_list <-lapply(fns, readRDS)
cat(length(fns), "files read \n")

# Get file size and time 
fileinfo <- file.info(fns)
# fileinfo

# Extract 'seriesno' from file name (just for checking, see nex chunk)
fileinfo_no <- substr(rownames(fileinfo), nchar(folder_results) + 8, nchar(folder_results) + 11)

```

### b. Extract metadata to dataframe for "successes"  


- jags_finished = TRUE, ok = TRUE: Model fitting done, but did not work for any k values    
- jags_finished = TRUE, ok = FALSE: Model fitting done, but did not work for any k values    
- jags_finished = FALSE, ok = FALSE: Model fitting not done (probably because JAGS was interrupted)      

```{r}

jags_finished <- map_lgl(result_list, ~!is.null(.x$k_values_ok))
ok <- map_lgl(result_list, ~!is.null(.x$DIC))

seriesno <- map_dbl(result_list, "seriesno")
PARAM <- map_chr(result_list, "PARAM")
STATION_CODE <- map_chr(result_list, "STATION_CODE")
TISSUE_NAME <- map_chr(result_list, "TISSUE_NAME")
LATIN_NAME <- map_chr(result_list, "LATIN_NAME")
Basis <- map_chr(result_list, "Basis")
k_sel <- NA
k_sel[ok] <- map_int(result_list[ok], "k_sel")
k_sel_dic <- NA
k_sel_dic[ok] <- map_int(result_list[ok], "k_sel_dic")
k_sel_forward <- NA
k_sel_forward[ok] <- map_int(result_list[ok], "k_sel_forward")

dat_success <- data.frame(seriesno, jags_finished, ok,   
                          PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis, k_sel) %>%
  # Add Rule1 and 2
  left_join(df_rule12, by = indexvars) %>%
  # Add N_years, Years_over_LOQ, k_max
  left_join(dat_series_trend %>% 
              select(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis, N_years, N_years_10yr, Years_over_LOQ, k_max),
            by = indexvars) %>%
  arrange(seriesno)

# Check that 'fileinfo' is in same order as 'dat_success' - if so, they can simply be joined
check <- seriesno == as.numeric(fileinfo_no)
if (sum(check) == 1){
  dat_success <- dat_success %>%
  bind_cols(fileinfo %>% select(mtime, size))
} else {
  warning("fileinfo not in the same order. Did not add mtime, size. (But that doesn't really matter.)")
}

```




### b2. In case of missing files    

```{r}

n_files_lacking <- nrow(dat_series_trend) - nrow(dat_success)

if (n_files_lacking > 0){
  stop("Not all files have been generated")
}

if (n_files_lacking){
  
  cat(n_files_lacking, "files lacking \n")

  # Plot series number of files
  plot(dat_success$seriesno)

  ser_orig <- 1:nrow(dat_series_trend)
  ser_lacking <- setdiff(ser_orig, dat_success$seriesno)
  plot(ser_lacking)
  tail(ser_lacking, 200)
  
  dat_series_trend_lacking <- dat_series_trend %>%
    filter(series_no %in% ser_lacking)
  
  if (FALSE){
  saveRDS(dat_series_trend_lacking, 
          paste0(folder_input, "/125_dat_series_trend_lacking.rds"))
  }
  
}

```




### c1. Check dat_success and dat_series  

- that they have the same order

```{r}

for (var in c("PARAM", "STATION_CODE", "TISSUE_NAME", "LATIN_NAME")){
  if (sum(dat_series_trend[[var]] != dat_success[[var]]) > 0){
    stop("Key columns have different order")
  }  
} 

cat("If you got this far, all key columns have the same order")  

```

### d1. Statistics  

- VDSI, VDSI/Intersex, "_exloq" parameters, and siloxanes D4-D6 at 19N contains zeros => becomes "-Inf" after log   
- Some (BDE49 at 30A) have k_max = 1 (ie no trend analysis)     
 
```{r}

xtabs(~ok, dat_success)  

cat("\n")
cat("PARAM for those with ok = FALSE (top 15) :\n")
table(subset(dat_success, !ok)$PARAM) %>% sort(decreasing = TRUE) %>% head(15)

cat("\n")
cat("STATION_CODE for those with ok = FALSE (top 15) :\n")
table(subset(dat_success, !ok)$STATION_CODE) %>% sort(decreasing = TRUE) %>% head(15)

cat("\n")
cat("Estimation success vs k_max :\n\n")
xtabs(~ok + k_max, dat_success)

if (FALSE){
  
  cat("\n")
  cat("Station for PARAM = CB_S7_exloq :\n\n")
  xtabs(~ok + STATION_CODE, subset(dat_success, PARAM == "CB_S7_exloq"))
  
  cat("\n")
  cat("Station for PARAM = BDE49 :\n\n")
  xtabs(~ok + STATION_CODE, subset(dat_success, PARAM == "BDE49"))
  
  xtabs(~addNA(k_sel) + N_years, dat_success)
  
  xtabs(~addNA(k_sel) + Years_over_LOQ, dat_success)

}

```
### d2. Statistics on selected k (model selection)
```{r}

xtabs(~k_sel + k_sel_dic, dat_success); cat("\n")
xtabs(~k_sel + k_sel_forward, dat_success)

# xtabs(~k_sel_dic + k_sel_forward + k_sel, dat_success)

```

### e. Check number of -Inf data  

- Should be "zero" only (no series that didn't work, did it because they had zeroes)   

```{r}

if (FALSE){
  
  # Can be skipped - should not be a problem anymore 
  # (after rewriting 2b in script 125)
  
  dat_lackof_success <-subset(dat_success, !ok)
  
  dat_lackof_success$n_infin <- map_int(
    dat_lackof_success$seriesno,
    ~sum(is.infinite(extract_raw_data(.x)$y))
  )
  
  table(dat_lackof_success$n_infin)  
  
}

```


### f. Plot a single series  
```{r}

# Worked:
# debugonce(tsplot_seriesno)
tsplot_seriesno(58, folder_results)

```




## 4b. For one compound  

### a. Plot time series, all stations      

```{r, fig.width=13, fig.height=10}

param <- "MCCP eksl. LOQ"
param <- "CB118"
basis <- "WW"

param <- "HG"
basis <- "WWa"

dat_success_pick <- dat_success %>%
  filter(PARAM == param,
         Basis %in% basis,
         ok)

pno <- dat_success_pick %>% pull(seriesno)

# debugonce(extract_modelfit_data)
test <- extract_modelfit_data(pno[8], folder = folder_results)    # last model is best
test <- extract_modelfit_data(pno[5], folder = folder_results)    # last model is best
test <- extract_modelfit_data(pno[13], folder = folder_results)   # k = 2 didn't work, model 3 is best

df_modelfit <- map_dfr(pno, extract_modelfit_data, folder = folder_results)
# if the previous line doesn't work:
# extract_modelfit_data_s <- safely(extract_modelfit_data)
# df_modelfit <- map(pno, extract_modelfit_data_s, folder = folder_results) %>%
#   transpose()
# df_modelfit$error %>% map_lgl(~!is.null(.x)) %>% which()    # the ones that didn't work

df_rawdata <- map_dfr(pno, extract_raw_data)

# debugonce(extract_raw_data)
# extract_raw_data(3001)

ggplot(df_modelfit, aes(x, y)) +
  geom_ribbon(aes(ymin = y_q2.5, ymax = y_q97.5), fill = "lightblue") +
  geom_point(data = df_rawdata %>% filter(!is.na(y))) +
  geom_point(data = df_rawdata %>% filter(!is.na(threshold)), aes(y = threshold), shape = 6) +
  geom_line(color = "red2") +
  facet_wrap(vars(STATION_CODE), scales = "free_y") +
  labs(title = param)

```


### b. Check trend file that goes into the big excel file  

```{r}

# Check trends in big excel file:  
# - use code in script 201 section 7 to get 'check' 
# head(check[c(1, 126:143-9)])

# Columns added:
# Trend p(long)  Detectable % change(long)  First Year(long)  Last Year(long)  No of Years(long) 
# Trend p(short) Detectable % change(short) First Year(short) Last Year(short) No of Years(short) 
# Trends.2021 ("¢/¢" format)  

# Index columns includes Basis, with values:
# DW  DWa   FB  FBa   WW  WWa

if (FALSE){
  c("PARAM", "LATIN_NAME", "TISSUE_NAME", "STATION_CODE", "Basis", 
    "Trend p(long)", "Detectable % change(long)", "First Year(long)", 
    "Last Year(long)", "No of Years(long)", "Trend p(short)", "Detectable % change(short)", 
    "First Year(short)", "Last Year(short)", "No of Years(short)", "Trends.2021")
}

# The trend symbols are made in 'set_symbol' which uses:
# Model_used ("Linear", "Nonlinear"), N_data, P_change, Annual_change (needs only Annual_change >0 or <0)

# up = é      ascii 233  Increasing concentration
# down = ê    ascii 234  Decreasing concentration
# circle = ¢  ascii 162  No significant time trend
# square = §  ascii 167  Too few years to make time trend
# star = «    ascii 171  Too few years with data over LOQ to make time trend

```



## 5. For all: Extract trends (long- and short-term)   

- Almost all code is extremely similar to code for one parameter, just removed the 'filter(PARAM == param) ' two places  

### a. All with trend: get difference data       
```{r}

pno_change1 <- dat_success %>%
  filter(ok) %>%
  # Those with a trend model selected as the best:
  filter(k_sel >= 2) %>%
  pull(seriesno)

# test
# extract_difference_data(49, folder_results)

df_with_trend_all <- map_dfr(pno_change1, extract_difference_data, folder = folder_results)

```

### b. All with trend: Extract up/down trends     
```{r}

df_trend_short <- df_with_trend_all %>%
  group_by(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis) %>%
  mutate(
    Year_pick = max((last_year-9), min(x)),
    Length_adequate = TRUE) %>%
  filter(x == Year_pick)

df_trend_long <- df_with_trend_all %>%
  group_by(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis) %>%
  mutate(Year_pick = min(x),
         Length_adequate = Year_pick < (last_year-9)) %>%
  filter(x == Year_pick) 

# Delete irrelevant info in 'df_trend_long'  
# names(df_trend_long) %>% dput()
cols_to_delete <- c("x", "y", "y_q2.5", "y_q97.5", "y_mean", "y_sd", "p", "p_category")
rows_select <- !df_trend_long$Length_adequate
df_trend_long[rows_select, cols_to_delete] <- NA

# Combine
df_trend_worked <- bind_rows(
  df_trend_long %>% ungroup() %>% mutate(Trend_type = "long"),
  df_trend_short %>% ungroup() %>% mutate(Trend_type = "short")
) %>%
  mutate(
    Trend_string = case_when(
      p < 0.05 & y_mean > 0 ~ "Decreasing",
      p < 0.05 & y_mean < 0 ~ "Increasing",
      !Length_adequate ~ paste("No data before", last_year-9),
      TRUE ~ "No change")
  ) %>%
  left_join(
    dat_success %>%
       select(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis, 
              ok, k_max, k_sel, N_years, N_years_10yr, Years_over_LOQ),
    by = indexvars) %>%
  left_join(
    dat_series %>%
       select(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis, First_year, Trend_model),
     by = indexvars)

```

### c1. k = 1 (flat): get difference data       
```{r}

pno_change2 <- dat_success %>%
  filter(ok) %>%
  # Pick those with no change. in the sense k = 1:
  filter(k_sel == 1) %>%
  pull(seriesno)

# test
# head(pno_change2)
# debugonce(extract_difference_data)
# extract_difference_data(7, folder_results)

df_flat_differences <- map_dfr(pno_change2, extract_difference_data, folder = folder_results)

```


### c2. k = 1 (flat): make   
```{r}

#
# Short
#

df_trend_flat_short <- dat_success %>%
  select(seriesno, PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis, 
              ok, k_max, k_sel, N_years, N_years_10yr, Years_over_LOQ) %>%
  filter(ok, 
         k_sel == 1) %>%
  mutate(
    Trend_string = case_when(
      N_years < 5 ~ "Too few years",
      N_years >= 5 & Years_over_LOQ < 5 ~ "Too few over-LOQ years",
      Years_over_LOQ >= 5 ~ "No trend"),
    Trend_type = "short") %>%
  left_join(
    dat_series %>%
       select(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis, First_year, Last_year, Trend_model),
     by = indexvars)

#
# Long
#

df_trend_flat_long <- df_trend_flat_short %>%
  mutate(
    Trend_string = case_when(
      First_year < (last_year-9) ~ Trend_string,
      First_year >= (last_year-9) ~ paste("No data before", last_year-9)),
    Trend_type = "long")

#
# Add differences
#

df_flat_differences_short <- df_flat_differences %>%
  group_by(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis) %>%
  mutate(
    Year_pick = max((last_year-9), min(x)),
    Length_adequate = TRUE) %>%
  filter(x == Year_pick)

df_flat_differences_long <- df_flat_differences %>%
  group_by(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis) %>%
  mutate(Year_pick = min(x),
         Length_adequate = Year_pick < (last_year-9)) %>%
  filter(x == Year_pick) 

df_trend_flat_short <- df_trend_flat_short %>%
  left_join(
    df_flat_differences_short, by = indexvars)

df_trend_flat_long <- df_trend_flat_long %>%
  left_join(
    df_flat_differences_long, by = indexvars)

```

### d. The ones that failed  
```{r}

df_trend_failed_short <- dat_success %>%
  select(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis, 
              ok, k_max, k_sel, N_years, N_years_10yr, Years_over_LOQ) %>%
  filter(!ok) %>%
  mutate(
    Trend_string = case_when(
      N_years < 5 ~ "Too few years",
      N_years >= 5 & Years_over_LOQ < 5 ~ "Too few over-LOQ years",
      Years_over_LOQ >= 5 ~ "Estimation failed"),
    Trend_type = "short") %>%
  left_join(
    dat_series %>%
       select(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis, First_year, Last_year, Trend_model),
     by = indexvars) 

df_trend_failed_long <- df_trend_failed_short %>%
    mutate(
      Trend_string = case_when(
        First_year < (last_year-9) ~ Trend_string,
        First_year >= (last_year-9) ~ paste("No data before", last_year-9)),
      Trend_type = "long")

```

### e. The ones with "no model" 
```{r}

df_trend_toofew_short <- dat_series %>%
  filter(Trend_model %in% "No model") %>%
    mutate(
    Trend_string = case_when(
      N_years < 5 ~ "Too few years",
      N_years >= 5 & Years_over_LOQ < 5 ~ "Too few over-LOQ years"),
    Trend_type = "short"
  )

df_trend_toofew_long <- df_trend_toofew_short %>%
  mutate(
    Trend_string = case_when(
      First_year < (last_year-9) ~ Trend_string,
      First_year >= (last_year-9) ~ paste("No data before", last_year-9)),
    Trend_type = "long")

```

### f. Combine   

-  Plus a bunch of checks  

```{r}

# Column 'dataset' is added only to make it easier to find errors/logical mistakes
#   (it is removed at the end of this chunk)
df_trend <- bind_rows(
  df_trend_worked %>% mutate(dataset = 1),
  df_trend_flat_long %>% mutate(dataset = 2), 
  df_trend_flat_short %>% mutate(dataset = 3),
  df_trend_failed_long %>% mutate(dataset = 4), 
  df_trend_failed_short %>% mutate(dataset = 5),
  df_trend_toofew_long %>% mutate(dataset = 6), 
  df_trend_toofew_short %>% mutate(dataset = 7)
  )

if (FALSE){
  param <- "HG"; station <- "10B"
  df_trend %>% filter(PARAM %in% param & STATION_CODE %in% station) %>% View(paste(param,station, "1"))
}

check1 <- df_trend %>%
  add_count(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis, Trend_type) %>%
  arrange(PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis, Trend_type) %>%
  filter(n > 1)

if (nrow(check1) > 0){
  stop("Some series occur more than once!")
}

check2 <- dat_series %>%
  anti_join(df_trend, by = indexvars)

if (nrow(check1) > 0){
  stop("Some series not included in df_trend!")
}

check3 <- df_trend %>%
  filter(is.na(Trend_string))

if (nrow(check3) > 0){
  stop("Some series lack 'Trend_string'!")
} 

check4 <- table(df_trend$Trend_type)
# table(df_trend$Trend_type, df_trend$dataset)
if (check4[2] - check4[1] != 0){
  stop("Some series lack 'long' (or less likely, short) trends")
} 

check5 <- table(df_trend$Trend_type, addNA(df_trend$Trend_string))
j <- grep("No data before", colnames(check5))
if (check5['short', j] > 0){
  stop("Some 'short' series have Trend_string = ", sQuote(colnames(check5)[j]))
} 


# If we got this far...
df_trend <- df_trend %>% select(-dataset)

```


### g. Add trend symbols and percent change    

- Check trends in big excel file:  
```
# - use code in script 201 section 7 to get 'check' 
head(check[c(1, 126:143-9)])

# Columns needed:
#   "PARAM", "LATIN_NAME", "Basis", "TISSUE_NAME", "STATION_CODE", "Basis", 
#   "Trend p(long)", "Detectable % change(long)", "First Year(long)", "Last Year(long)", "No of Years(long)", 
#   "Trend p(short)", "Detectable % change(short)", "First Year(short)", "Last Year(short)", "No of Years(short)", 
#   "Trends.2021"

# Index columns includes Basis, with values:
# DW  DWa   FB  FBa   WW  WWa

```

Symbols: 

- up = é      ascii 233  Increasing concentration
- down = ê    ascii 234  Decreasing concentration
- circle = ¢  ascii 162  No significant time trend
- square = §  ascii 167  Too few years to make time trend
- star = «    ascii 171  Too few years with data over LOQ to make time trend  

```{r}

df_trend <- df_trend %>%
  mutate(
    Perc_change = round((1-exp(y_mean))/exp(y_mean)*100, 1),
    D_year = last_year - First_year,
    Perc_annual = round((1-exp(y_mean/D_year))/exp(y_mean)*100, 1),
    Perc_annual_lo = round((1-exp(y_q2.5/D_year))/exp(y_mean)*100, 1),
    Perc_annual_hi = round((1-exp(y_q97.5/D_year))/exp(y_mean)*100, 1),
    Trend_symbol = case_when(
      is.na(Trend_string) ~ "x",
      Trend_string %in% "Increasing" ~ "é",
      Trend_string %in% "Decreasing" ~ "ê",
      Trend_string %in% "No change" ~ "¢",
      Trend_string %in% "No trend" ~ "¢",
      Trend_string %in% "Too few years" ~ "§",
      Trend_string %in% "Too few over-LOQ years" ~ "«")
    )

check1 <- df_trend %>%
  filter(is.na(First_year))

if (nrow(check1) > 0){
  stop("Some series lack 'First_year'")
}

```


### h. Put them on wide format (separate columns for short and long)   

```{r}
   
df_trend_wide <- df_trend %>%
  select(
    PARAM, STATION_CODE, TISSUE_NAME, LATIN_NAME, Basis, 
    Trend_type, Trend_string, Trend_symbol, Perc_change, p, First_year, Last_year, N_years, N_years_10yr) %>%
  tidyr::pivot_wider(
    names_from = Trend_type, 
    values_from = c(Trend_string, Trend_symbol, Perc_change, p, First_year, Last_year)) %>% 
  mutate(
    First_year_short = ifelse(
      First_year_short < (last_year-9), (last_year-9), First_year_short),
    Trend_symbol_long = ifelse(is.na(Trend_symbol_long), " ", Trend_symbol_long),
    Trend_symbol_short = ifelse(is.na(Trend_symbol_short), " ", Trend_symbol_short),
    Trend_symbol = paste0(Trend_symbol_long, "/", Trend_symbol_short)
  )

if (FALSE){
  # HARD_CODED
  df_trend_wide$Last_year_long <- last_year
  df_trend_wide$Last_year_short <- last_year
}

```


## 6. Save + read  

```{r}

saveRDS(dat_success, paste0(folder_output, "/126_dat_success_2021.rds"))
saveRDS(df_trend, paste0(folder_output, "/126_df_trend_2021.rds"))
saveRDS(df_trend_wide, paste0(folder_output, "/126_df_trend_wide_2021.rds"))

if (FALSE){
  # For reading back
  dat_success <- readRDS(paste0(folder_output, "/126_dat_success_2021.rds"))
  df_trend <- readRDS(paste0(folder_output, "/126_df_trend_2021.rds"))
  df_trend_wide <- readRDS(paste0(folder_output, "/126_df_trend_wide_2021.rds"))
}

```


## APPENDIX 1:  Various checking   


### Some that failed  
```{r}

# Short series (n = 4)
tsplot_param("BDE49", "30A", folder = folder_results)

# No apparent reason for failure?
tsplot_param("D5", "19B", folder = folder_results)
tsplot_param("D5", "19N", tissue = "Egg", folder = folder_results)
tsplot_param("CB101", "30B", folder = folder_results)             # Now works
# extract_raw_data(get_seriesno("CB101", "30B")) %>% View()

```

### Some that worked
```{r}

# debugonce(tsplot_seriesno)

tsplot_param("CB118", "30B", folder = folder_results)
tsplot_param("HG", "30B", folder = folder_results)
tsplot_param("BDE47", "I023", folder = folder_results)
tsplot_param("HG", "98B1", folder = folder_results, )
tsplot_param("MCCP eksl. LOQ", "28B", folder = folder_results)  
tsplot_param("CB_S7_exloq", "10A2", folder = folder_results)

```

## APPENDIX 2. Get series that didn't work  

- For re-run, after revising the models used by the trend analyses  

```{r}

dat_rerun <- dat_success %>%
  filter(!ok & k_max >= 2) %>%
  filter(!grepl("SCCP", PARAM)) %>%
  filter(!grepl("MCCP", PARAM))

xtabs(~N_years, dat_rerun)  

```


## APPENDIX 3. Old model selection stuff

### Check trend tables

```{r}

if (FALSE){
  
param <- "HG"; station <- "10B"
df_trend %>% filter(PARAM %in% param & STATION_CODE %in% station) %>% View(paste(param,station, "1"))
df_trend_wide %>% filter(PARAM %in% param & STATION_CODE %in% station) %>% View(paste(param,station, "2"))

# And plot
tsplot_param(param, station, folder = folder_results)

# And model result table  
seriesno <-  get_seriesno(param, station)
fn <- sprintf("trend_%04.0f.rda", seriesno)
resultlist <- readRDS(paste0(folder_results, "/", fn))
resultlist$dDIC

# And model log.likelihood ratio test  
resultlist$lr_table

# Model 1 is selected because model 2 is worse than model 1, and we use a strict forward selection
# But if we had skipped model 2, model 3 would be better than model 1  

}

```


### Check what would have happend if we skipped model 2, or model 2+3
```{r}

if (FALSE){

# Skip model 2

tab2 <- resultlist$lr_table[-2,]
# Recalculate ddev and p
tab2$ddev <- c(NA, -diff(tab2$dev))
for (i in 2:nrow(tab2)){
  k_diff <- as.numeric(tab2$k[i]) - as.numeric(tab2$k[i-1])
  tab2$p[i] <- 1 - pchisq(tab2$ddev[i], k_diff)
}
tab2

# Skip model 2+3

tab3 <- resultlist$lr_table[-(2:3),]
# Recalculate ddev and p
tab3$ddev <- c(NA, -diff(tab3$dev))
for (i in 2:nrow(tab3)){
  k_diff <- as.numeric(tab3$k[i]) - as.numeric(tab3$k[i-1])
  tab3$p[i] <- 1 - pchisq(tab3$ddev[i], k_diff)
}
tab3

}

```
