---
title: "03_Read_fish_individual_data"
author: "DHJ"
date: "27 8 2019"
output: 
  html_document:
    keep_md: true
---

## Libraries
```{r}
library(dplyr)
library(tidyr)
library(purrr)     # map
library(readxl)
library(ggplot2)
library(stringr)
source("150_Get_fish_individual_data_functions.R")
```

## Files
```{r}
datafolder <- "Input_data/opparbeiding_biota/Fisk_2018"
fns <- dir(datafolder, "*.xls*")
fns <- fns[substr(fns,1,1) != "~"]   # 'locked' file (remove opened excel files)
fns %>% dput()
```

## Test read of one file  
Used to make the read_fish_individual_data() function. Using search_lineno()    
```{r}
fn <- fns[1]

df_try <- read_excel(paste0(datafolder, "/", fn))

line_header <- search_lineno(df_try, 1, "Fisknr")
dat <- read_excel(paste0(datafolder, "/", fn), skip = line_header - 1)

# If file has a hole (empty row) in the data, we delete data from the first hole
first_empty <- which(is.na(dat$Fisknr))[1]
if (!is.na(first_empty))
  dat <- dat[1:(first_empty-1),]

dat
```

## Test 'read_fish_individual_data'
```{r}
# debugonce(read_fish_individual_data )
read_fish_individual_data(fns[1])
```

## Read all data into a list of data frames  
    - Reads all columns of each file  
    - Each data frame may differ from the other  
```{r}
fns_to_read <- fns[!fns %in% c("2018_MAL-FISK-stasjonskode_stasjonsnavn.xlsx", 
                               "ikke_bruk_milkys data 2018 Svalbard og Tromsø - akvaplan-niva.xlsx", 
                               "oversikt_over_fisk_2018.xlsx")]
dat_list <- fns_to_read %>% map(read_fish_individual_data) 
```

## Check column names
```{r}
# Get column name for a given column number 
get_colname <- function(df, colnumber) colnames(df)[colnumber]

# Get all column names, by column number 
for (i in 1:26){
  cat(i, "\n")
  column_names <- dat_list %>% map_chr(get_colname, colnumber = i)
  print_less_common_values(column_names)
  cat("------------------------------------\n")
  }

```

## Print all columns  
Change 'print_all' to TRUE to show result
```{r}
print_all <- FALSE
if (print_all){
  for (i in 1:26){
  cat(i, "\n")
  print(
    dat_list %>% map_chr(get_colname, colnumber = i)
  )
  cat("------------------------------------\n")
  }

}
```

## Pick specific columns from each data frame   
Levervekt not found for data set 3 (13B)  
```{r}
# Get selected columns  
# 
dat_list_selected <- dat_list %>% 
  map(get_columns, colnames = c("Station", "Species", "Fisknr", "Lengde", "Vekt", "Kjønn", 
                                "Levervekt", "Leverfarge", "Gonadevekt", "Alder"))

# Look for Levervekt:
# dat_list[[3]]
```

## Make data frame list ready for combining 1  
```{r}
#
# Check Station (NOTE: HARD-CODED FILE NUMBERS)
#
cat("Existing values of in data:\n")
dat_list_selected %>% map_chr(~unique(.$Station))

# Station is lacking for data set number 13 and 18
cat("\n\nStations - Get file names:\n")
fns_to_read[c(13,18)]
dat_list_selected[[13]]$Station <- "43B2"
dat_list_selected[[18]]$Station <- "96B"
# We will fix those "code + name" stations later

#
# Check Species (NOTE: HARD-CODED FILE NUMBERS)
#
dat_list_selected %>% map_chr(~unique(.$Species))
# Station is lacking for data set number 13 and 18
cat("\n\nSpecies - Get file names:\n")
fns_to_read[c(8,13,17,18)]
dat_list_selected[[8]]$Species <- "Torsk"
dat_list_selected[[13]]$Species <- "Torsk"
dat_list_selected[[17]]$Species <- "Torsk"
dat_list_selected[[18]]$Species <- "Torsk"

```

## Make data frame list ready for combining 2  
```{r}
#
# Check column classes
#
dat_list_selected %>% map_df(get_column_classes)

# 
# Change type of column, where needed
#
dat_list_selected <- dat_list_selected %>% map(~char2numeric(., "Fisknr"))
dat_list_selected <- dat_list_selected %>% map(~char2numeric(., "Lengde"))
dat_list_selected <- dat_list_selected %>% map(~char2numeric(., "Vekt"))
dat_list_selected <- dat_list_selected %>% map(~char2numeric(., "Gonadevekt"))
dat_list_selected <- dat_list_selected %>% map(~char2numeric(., "Alder"))

```

## Combine into a single data frame  
```{r}
dat <- bind_rows(dat_list_selected)
dat
```

## Create STATION_CODE    
    - STATION_CODE = name removed from "code + name" stations   
    - Also rename(Fish_no = Fisknr)  
```{r}
xtabs(~addNA(Station), dat)

dat <- dat %>% 
  mutate(STATION_CODE = stringr::str_extract(Station, "([^[[:blank:]]]+)") ) %>%
  mutate(STATION_CODE = sub("b", "B", STATION_CODE)) %>%
  select(-Station) %>%
  rename(Fish_no = Fisknr,
         Length = Lengde, 
         Weight = Vekt,
         Liver_weight = Levervekt,
         Liver_color = Leverfarge,
         Gonad_weight = Gonadevekt,
         Age = Alder) %>%
  select(STATION_CODE, everything())

xtabs(~addNA(STATION_CODE), dat)
```

## Check lacking data
```{r}
apply(!is.na(dat), 2, mean)

```

## Check Station/species
```{r}
dat %>%
  count(STATION_CODE, Species)
```


## Test plot 1
```{r}
ggplot(dat, aes(Length, Weight, color = Species)) +
  geom_point() +
  facet_wrap(~STATION_CODE)
```
## Test plot 2
```{r}
ggplot(dat, aes(Weight, Gonad_weight, color = Species)) +
  geom_point() +
  facet_wrap(~STATION_CODE)
```

## Test plot 3
```{r}
ggplot(dat, aes(Weight, Liver_weight, color = Species)) +
  geom_point() +
  facet_wrap(~STATION_CODE)
```

## Test plot 4
```{r}
ggplot(dat, aes(Length, Age, color = Species)) +
  geom_point() +
  facet_wrap(~STATION_CODE)

```

## Save  
```{r}
saveRDS(dat, "Data/03_Fish_individual_data.rds")
write.csv(dat, "Data/03_Fish_individual_data.csv", fileEncoding = "UTF-8") 

```

```{r}
# The "ordinary knot button" doesn't work because pandoc fails to convert the .md file to HTML
# Instead run the following line (must not be uncommented, that may create a loop):
#   knitr::knit("03_Read_fish_individual_data.Rmd", encoding = "UTF-8") 
```

