---
title: "03_Read_fish_individual_data"
author: "DHJ"
date: "31 8 2021"
output: 
  html_document:
    keep_md: true
---

## Reading fish data from excel files   

* Before starting on this, you need to manually copy the "fish files" to Jupyterhub  

    - Present (early 2022) location: 

## Libraries  

```{r}

library(dplyr)
library(tidyr)
library(purrr)     # map
library(readxl)
library(ggplot2)
library(stringr)
source("150_Get_fish_individual_data_functions.R")


```
## Year  
```{r}

selected_year <- 2022


```


## Files
```{r}

datafolder <- paste0("Input_data/opparbeiding_biota/Fisk_", selected_year)

# dir.exists(datafolder)

fns <- dir(datafolder, "*.xls*")
fns <- fns[substr(fns,1,1) != "~"]   # 'locked' file (remove opened excel files)
fns %>% dput()

```

## Test read of one file  
Used to make the read_fish_individual_data() function. Using search_lineno()    
```{r}

fn <- fns[1]

df_try <- read_excel(paste0(datafolder, "/", fn))

line_header <- search_lineno(df_try, 1, "Fisknr")
dat <- read_excel(paste0(datafolder, "/", fn), skip = line_header - 1)

# If file has a hole (empty row) in the data, we delete data from the first hole
first_empty <- which(is.na(dat$Fisknr))[1]
if (!is.na(first_empty))
  dat <- dat[1:(first_empty-1),]

dat

```

## Test 'read_fish_individual_data'
```{r}

# debugonce(read_fish_individual_data )
read_fish_individual_data(fns[1])

```

## Read all data into a list of data frames  
    - Reads all columns of each file  
    - Each data frame may differ from the other   

### Extract station code to use for list names  

* should be identical to station code    

```{r}

fns_to_read <- fns

# Start by Extract all characters to the left of the first underscore  
names(fns_to_read) <- stringr::str_extract(fns, "[^_]+")

# Check
names(fns_to_read)

# Remove first 'x'  
sel <- str_sub(names(fns_to_read), 1, 1) == "x"
names(fns_to_read)[sel] <- str_sub(names(fns_to_read)[sel], start = 2)

# Check
names(fns_to_read)

# Manual change, if needed  
# names(fns_to_read)[grepl("10B", fns)] <- "10B"
# names(fns_to_read)[grepl("96B", fns)] <- "96B"

```

    
### Read data into list   

* Takes a couple of minutes  

```{r, message=FALSE}

# Make "safe" version of 'read_fish_individual_data'
# Meaning that we continue with next file even if there is an error in one file 
read_fish_individual_data_s <- safely(read_fish_individual_data)

# Results in list with 17 items, where each item consistst of "$result" and "$error"           #ELU: 18 items/stasjoner
dat_list_safe <- fns_to_read %>% map(read_fish_individual_data_s)  

# Transpose list
# Results in list with 2 items ("$result" and "$error"), each with 17 items 
dat_list_safe <- purrr::transpose(dat_list_safe)

names(dat_list_safe$result)

```


### Check results  

```{r}

is_ok <- dat_list_safe$error %>% map_lgl(is.null)

if (sum(!is_ok) > 0){
  
  # If any files didn't work, you will get file number here
  which(!is_ok) 
  
  # If any files didn't work, you will get file name here
  fns_to_read[!is_ok] 
  
} else {
  
  message("All files were read")
  
}

# If NOT ok, uncomment the next lines ('15' should be replaced by number of file which casued error ): 
# debugonce(read_fish_individual_data)
# read_fish_individual_data(fns_to_read[15])

```

### Resulting list of results
```{r}

dat_list <- dat_list_safe$result[is_ok]

# Get names as well:
names(dat_list) <- names(dat_list_safe$result)[is_ok]

```

## Check column names   

* Goes through column numer 1, 2 etc.   
* For each column, checks the name of that column in all tables    
* If it says 'Only value', it means that all tables have the same column name (and thus are ok)  

```{r}

# Number of columns per file:
# dat_list %>% map_int(length)    # always 26
  
# Get column name for a given column number 
get_colname <- function(df, colnumber) colnames(df)[colnumber]

# Get all column names, by column number 
for (i in 1:26){
  cat("Column", i, ":  \n")
  column_names <- dat_list %>% map_chr(get_colname, colnumber = i)
  print_less_common_values(column_names)
  cat("------------------------------------\n")
  }

```

## Print all columns  
Change 'print_all' to TRUE to show result
```{r}

print_all <- FALSE                                                  
if (print_all){
  for (i in 1:26){
  cat(i, "\n")
  print(
    dat_list %>% map_chr(get_colname, colnumber = i)
  )
  cat("------------------------------------\n")
  }

}

```

## Pick specific columns from each data frame   

Levervekt not found for data set 3 (13B)                               #ELU: 13B ok for 2022
Note: Upper- and lowercase is ignored - no need to worry about this   
- e.g.: 'Vekt' will find both 'Vekt(g)' and 'vekt(g)'

```{r}
# Get selected columns                                                  
# 
dat_list_selected <- dat_list %>% 
  map(get_columns, colnames = c("Station", "Species", "Fisknr", "Lengde", "Vekt", "Kjønn", 
                                "Levervekt", "Leverfarge", "Gonadevekt", "Alder"))

# Look for Levervekt:
# dat_list[[3]]

```

## Make data frame list ready for combining, part 1  
```{r}

# Fix station codes using names of the list   
for (i in seq_along(dat_list_selected)){
  dat_list_selected[[i]]$Station <- names(dat_list_selected)[i]
}

#
# Check Station
#
cat("Existing values of in data:\n")
dat_list_selected %>% map_chr(~unique(.$Station))  


#
# Check Species (NOTE: HARD-CODED FILE NUMBERS)                          ELU: feil i 53B art rettet
#
dat_list_selected %>% map_chr(~unique(.$Species))

```

## Make data frame list ready for combining, part 2    

### Check classes of each column   

* Check up those that have unexcpected/different classes, e.g. character when it should have been numeric  

```{r}

dat_list_selected %>% map_df(get_column_classes)                  #ELU: mange har feil class

```
### Manual check of some columns    

* char2numeric handles decimal comma instead of decimal dot, but not other changes  
* So just check in case    

```{r}

if (FALSE){
  
  dat_list_selected[[3]]$Lengde                                                 
  dat_list_selected[[9]]$Lengde
  dat_list_selected[[14]]$Lengde
  dat_list_selected[[4]]$Vekt
  dat_list_selected[[2]]$Levervekt
  dat_list_selected[[1]]$Gonadevekt
  dat_list_selected[[5]]$Gonadevekt
  dat_list_selected[[11]]$Gonadevekt
  
}

# One value contained a '+', remoce it
#dat_list_selected[[2]]$Levervekt <- sub("+", "",   dat_list_selected[[2]]$Levervekt, fixed = TRUE)    #ELU:kjører ikke denne


```

### Change columns that should be numeric, to be numeric   

* If it is character and contains decimal comma, that will be converted by 'char2numeric'     
* Things like "nd" will jsut get NA value (as they should) 
* All numeric columns are changed, to be on the safe side

```{r}

dat_list_selected <- dat_list_selected %>% map(~char2numeric(., "Fisknr"))
dat_list_selected <- dat_list_selected %>% map(~char2numeric(., "Lengde"))
dat_list_selected <- dat_list_selected %>% map(~char2numeric(., "Vekt"))
dat_list_selected <- dat_list_selected %>% map(~char2numeric(., "Levervekt"))
dat_list_selected <- dat_list_selected %>% map(~char2numeric(., "Gonadevekt"))
dat_list_selected <- dat_list_selected %>% map(~char2numeric(., "Alder"))

```

### Check lacking length  
```{r}

n_lacking <- dat_list_selected %>% map_int(~sum(is.na(.$Lengde)))

if (sum(n_lacking > 0) > 0) {
  stop(sum(n_lacking > 0), " station(s) have lacking 'Length' rows. Check them up! ")
} else {
  message("No stations have lacking 'Length' rows! All OK.")
}

# Only if needed!
if (FALSE){
  
  n_lacking[n_lacking > 0]
  
  
  # Check manually:
  # dat_list_selected[["15B"]] %>% View()
  
  # Fix (HARD-CODED):
  dat_list_selected[["15B"]] <- dat_list_selected[["15B"]][1:15,]
  
  # Check again:
  n_lacking <- dat_list_selected %>% map_int(~sum(is.na(.$Lengde)))
  if (sum(n_lacking > 0) > 0) {
    stop("Still ", sum(n_lacking > 0), " station(s) have lacking 'Length' rows. Check them up! ")
  } else {
    message("Fixed - no lacking 'Length' rows") 
  }
  
}


```


## Combine into a single data frame  
```{r}

dat_1 <- bind_rows(dat_list_selected, .id = "STATION_CODE")

# dat

nrow(dat_1)

```

## Fix specific columns   

### Create STATION_CODE    
- STATION_CODE = name removed from "code + name" stations   
- NOT NEEDED anymore, did this through the file name
```{r}

# not needed, used '.id = "STATION_CODE"' in previous chunk

# xtabs(~addNA(Station), dat_1)
# 
# dat_2 <- dat_1 %>% 
#   mutate(STATION_CODE = stringr::str_extract(Station, "([^[[:blank:]]]+)") ) %>%
#   mutate(STATION_CODE = sub(",", "", STATION_CODE, fixed = TRUE)) %>%
#   mutate(STATION_CODE = sub("b", "B", STATION_CODE)) %>%
#   mutate(STATION_CODE = ifelse(STATION_CODE %in% "43B", "43B2", STATION_CODE))  # 43B2 is the present code
# 
# xtabs(~addNA(STATION_CODE), dat_2)

dat_2 <- dat_1

```

### Fix Species

```{r}

# Check   
cat("\nCheck 1 \n")
xtabs(~addNA(Species), dat_2)

dat_3 <- dat_2 %>% 
  mutate(Species = stringr::str_to_sentence(Species)) %>%   # 'torsk' => 'Torsk'
  mutate(Species = ifelse(grepl("Gadu", Species), "Torsk", Species))

# Check again  
cat("\nCheck 2 \n")
xtabs(~addNA(Species), dat_3)

# Fix the station lacking species   

# check name of station(s)  
# xtabs(~addNA(STATION_CODE), subset(dat_3, is.na(Species)))

sel <- dat_3$STATION_CODE  == "02B"
dat_3$Species[sel] <- "Torsk"

# Check again
cat("\nCheck 3 \n")
xtabs(~addNA(Species), dat_3)


```

### Renaming  
```{r}

dat_4 <- dat_3 %>% 
  select(-Station) %>%
  rename(Fish_no = Fisknr,
         Length = Lengde, 
         Weight = Vekt,
         Liver_weight = Levervekt,
         Liver_color = Leverfarge,
         Gonad_weight = Gonadevekt,
         Age = Alder) %>%
  select(STATION_CODE, everything())


```

## Check lacking data   

### Summary
```{r}

apply(!is.na(dat_4), 2, mean)

```

### Fraction OK data per station
```{r}

dat_4 %>%
  select(-Species, -Kjønn, -Liver_color) %>%
  tidyr::pivot_longer(-STATION_CODE, names_to = "Variable", values_to = "Value") %>%
  group_by(STATION_CODE, Variable) %>%
  summarise(Fraction_ok = mean(is.finite(Value))) %>%
  tidyr::pivot_wider(names_from = "Variable", values_from = "Fraction_ok")

```

### Number of OK data per station
```{r}

dat_4 %>%
  select(-Species, -Kjønn, -Liver_color) %>%
  tidyr::pivot_longer(-STATION_CODE, names_to = "Variable", values_to = "Value") %>%
  group_by(STATION_CODE, Variable) %>%
  summarise(Number_ok = sum(is.finite(Value))) %>%
  tidyr::pivot_wider(names_from = "Variable", values_from = "Number_ok")

```


## Test plot 1, length  

```{r, fig.width=9, fig.height=6}

ggplot(dat_4, aes(Length, Weight, color = Species)) +
  geom_point() +
  facet_wrap(~STATION_CODE)

```

## Test plot 2, weight  

```{r, fig.width=9, fig.height=6}

ggplot(dat_4, aes(Weight, Gonad_weight, color = Species)) +
  geom_point() +
  facet_wrap(~STATION_CODE)

```

## Test plot 3, liver weight vs weight  

```{r, fig.width=9, fig.height=6}

ggplot(dat_4, aes(Weight, Liver_weight, color = Species)) +
  geom_point() +
  facet_wrap(~STATION_CODE)

```


## Test plot 4, age vs length  

```{r, fig.width=9, fig.height=6}

ggplot(dat_4, aes(Length, Age, color = Species)) +
  geom_point() +
  facet_wrap(~STATION_CODE)

```

## Save   
- Folder `Data`  
```{r}

# Wrapped in FALSE to avoid accidental overwriting

if (TRUE){
  
  saveRDS(
    dat_4, 
    paste0("Data/150_Fish_individual_data_ELU_", selected_year, ".rds")             #ELU-testfil
  )
  
  write.csv(
    dat_4, 
    paste0("Data/150_Fish_individual_data_ELU_", selected_year, ".csv"),           #ELU-testfil
    fileEncoding = "UTF-8") 
  
}

```

```{r}
# The "ordinary knit button" doesn't work because pandoc fails to convert the .md file to HTML
# Instead run the following line (must not be uncommented, that may create a loop):
#   knitr::knit("150_Read_fish_individual_data_2021.Rmd", encoding = "UTF-8") 
```

