---
title: "161_Read_NILU_excel_data_2019"
output: html_document
---

Note: The excel file needs to be fixed beforehand   
- The sheets are either   
    - type 1 (samples in columns, substances in rows) - to be read with read_excel_nilu1  
    - type 1 (samples in rows, substances in columns) - to be read with read_excel_nilu2  
- See decriptions and instructions for each of the to types in '161_Read_excel_data_functions.R'   
- In this case, sheets "HBCD", "PBDE" and "PCB" are of type 1 and the rest are type 2   


## 1. Packages and functions
```{r}
library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)

source('161_Read_NILU_excel_data_functions.R')
source("101_Combine_with_legacy_data_functions.R")  # for sum parameters
```

## Set year
```{r}

year <- 2019
station_code_not_given <- "19N"  # used whenever STATOPN_CODE is not goiven in data

```


## Data  
### Sample data    
Used to double-check sample numbers (see part 'Checks')  
```{r}

### Sample data for 2019   
# We need to have a look at this to fix the siloksan data (which also contains cod) 
files <- dir("Input_data", "Labware_samples_") %>% rev()
filename <- files[1] 
df_samples <- readRDS(paste0("Input_data/", filename))

df_samples_sel <- df_samples %>%
  filter(grepl("2019", TEXT_ID)) %>%
  filter(AQUAMONITOR_CODE %in% c("24B","19B","43B2","30B")) %>%   # from looking in "siloksaner" in the Excel file
  select(AQUAMONITOR_CODE, TEXT_ID, SAMPLE_NUMBER, TISSUE, BIOTA_SAMPLENO, DESCRIPTION) %>%
  arrange(AQUAMONITOR_CODE, TISSUE, BIOTA_SAMPLENO)

```

### Existing main data  
ONLY for checking (variable names etc.)  
Read and reformat the most recent data (by default)  
Code identical to the code in the start of script 109   
```{r}

dat_new3 %>% names() %>% dput()

c("MYEAR", "STATION_CODE", "LATIN_NAME", "SAMPLE_NO2", "TISSUE_NAME", 
"PARAM", "BASIS", "VALUE", "FLAG1", "UNIT")


files <- dir("Data", pattern = "101_data_updated") %>% rev()

cat("Reading the last file downloaded:")
cat("\n", files[1])
cat("\n")
cat("If you want to read a different file, replace 'files[1]' with the file you want")
cat("\n")

filename <- files[1] 
data_all <- readRDS(paste0("Data/", filename))

# We save the date part of the text (e.g., '2020-04-23')
# This will be used in part 10, when we save the resulting file
file_date <- substr(filename, 18, 27)    # pick out the part of the text from character no. 17 to no. 26

```

## Read from excel  

```{r}

filename <- "Input_data/NILU_data_for_2019_korrigert_edited.xlsx"
sheets <- excel_sheets(filename)
sheets

# Empty list for data
dat <- vector("list", length(sheets))

```

### Sheet 1 - PBDE (type 1)  
```{r, message=FALSE, fig.width = 9, fig.height=6}

# Data 1-2: read_excel_nilu1
# debugonce(read_excel_nilu1)
dat[[1]] <- read_excel_nilu1(filename, "PBDE", 
                             find_upper_top = "sample number", # text to search for in first line of upper part
                             find_upper_bottom = "file",            # text to search for in last line of upper part
                             find_lower_top = "structure",          # text to search for in first line of lower part
                             name_Sample_no_NILU = "NILU-Sample number:",  # Names given in Excel sheet (must fit exactly)
                             name_NIVA_id = "NIVA-ID",  
                             name_Tissue = "type:", 
                             name_Sample_amount = "amount:",
                             name_Unit = "unit",
                             lessthans_given_as_negative_number = TRUE) %>%
  mutate(Group = "PBDE",
         STATION_CODE = "19N", 
         BIOTA_SAMPLENO = stringr::str_extract(Label, "[0-9]+") %>% as.numeric())

ggplot(dat[[1]], aes(Parameter, Value)) +
  geom_boxplot() +
  facet_wrap(vars(Tissue), nrow = 1, scales = "free_y") +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

```

### Check parameter names    
2,3,3',4,4',5',6-HepBDE occurs twice in the data, with IUPAC-no. 191 and 197   
From the Polybrominated Diphenyl Ethers EQS dossier 2011 we find that  
- 2,3,3’,4,4’,5’,6-HepBDE = 191  
- 2,2’,3,3’,4,4’,6,6’-OctBDE = 197  
```{r}

dat[[1]] %>%
  distinct(Parameter, IPUAC_no) %>%
  group_by(Parameter) %>%
  mutate(n = n()) %>%
  filter(n > 1)

sel <- dat[[1]]$IPUAC_no %in% 197
dat[[1]]$Parameter[sel] <- "2,2’,3,3’,4,4’,6,6’-OctBDE"

cat(sum(sel), "Parameter names corrected \n")

```


### Sheet 2 -PCB (type 1)    
```{r, message=FALSE, fig.width = 9, fig.height=6}
dat[[2]] <- read_excel_nilu1(filename, "PCB", 
                             find_upper_top = "sample number", # text to search for in first line of upper part
                             find_upper_bottom = "file",            # text to search for in last line of upper part
                             find_lower_top = "structure",          # text to search for in first line of lower part
                             name_Sample_no_NILU = "NILU-Sample number:",  # Names given in Excel sheet (must fit exactly)
                             name_NIVA_id = "NIVA-ID",  
                             name_Tissue = "type:", 
                             name_Sample_amount = "amount:",
                             name_Unit = "unit",
                             lessthans_given_as_negative_number = FALSE) %>%
  mutate(Group = "PCB",
         STATION_CODE = "19N", 
         BIOTA_SAMPLENO = stringr::str_extract(Label, "[0-9]+") %>% as.numeric()) %>%
  filter(!is.na(Parameter))    # filter away the bottom line in the PCB excel sheet (below Sum 7 PCB)

# head(dat[[2]])

table(addNA(dat[[2]]$Parameter))

ggplot(dat[[2]], aes(Parameter, Value)) +
  geom_boxplot() +
  facet_wrap(vars(Tissue), nrow = 1, scales = "free_y") +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

```

### Sheet 3 - HBCD (type 1)  
```{r, message=FALSE, fig.width = 9, fig.height=6}

# debugonce(read_excel_nilu1)
dat[[3]] <- read_excel_nilu1(filename, "HBCD", 
                             find_upper_top = "sample number", # text to search for in first line of upper part
                             find_upper_bottom = "file",            # text to search for in last line of upper part
                             find_lower_top = "structure",          # text to search for in first line of lower part
                             name_Sample_no_NILU = "NILU sample number:",  # Names given in Excel sheet (must fit exactly)
                             name_NIVA_id = "NIVA-ID",  
                             name_Tissue = "type:", 
                             name_Sample_amount = "amount:",
                             name_Unit = "unit",
                             lessthans_given_as_negative_number = TRUE) %>%
  mutate(Group = "HBCD",
         STATION_CODE = "19N",
         BIOTA_SAMPLENO = stringr::str_extract(Label, "[0-9]+") %>% as.numeric())

# count(dat[[3]], Unit)
# dat[[3]] %>% filter(is.na(Unit))

# head(dat[[3]])


ggplot(dat[[3]], aes(Parameter, Value)) +
  geom_boxplot() +
  facet_wrap(vars(Tissue), nrow = 1, scales = "free_y") +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))

```

### Sheet 4 - Chlorinated paraffins (type 2)     
```{r, message=FALSE, fig.width = 9, fig.height=6}

dat[[4]] <- read_excel_nilu2(filename, 
                             "CP", 
                             lessthans_given_as_negative_number = TRUE) %>%
  mutate(Group = "Chlorinated paraffins",
         STATION_CODE = "19N")
head(dat[[4]])

ggplot(dat[[4]], aes(Parameter, Value)) +
  geom_boxplot() +
  facet_wrap(vars(Tissue), nrow = 1, scales = "free_y") +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))


```

### Sheet 5 - Siloxans (type 2)    
```{r, message=FALSE, fig.width = 9, fig.height=6}

dat[[5]] <- read_excel_nilu2(filename, 
                             "siloksaner", 
                             lessthans_given_as_negative_number = TRUE) %>%
  mutate(Group = "Siloxans")
# head(dat[[5]])

```


### Sheet 5: add STATION_CODE + sample no variables for cod (siloxans) 
```{r, fig.width = 9, fig.height=6}

# STATION_CODE
# Show part of "Label" between first and second space (space = \\s)
# dat[[5]]$Label %>% stringr::str_extract("(?<=\\s)[^\\s]+(?=\\s)") %>% table()

dat[[5]] <- dat[[5]] %>%
  # Add STATION_CODE
  mutate(STATION_CODE = stringr::str_extract(Label, "(?<=\\s)[^\\s]+(?=\\s)")) %>%
  mutate(STATION_CODE = sub("-", "", STATION_CODE)) %>% # View()
  mutate(STATION_CODE = case_when(
    is.na(STATION_CODE) & grepl("Ærfugl", Label) ~ "19N", 
    !is.na(STATION_CODE) ~ STATION_CODE)
    ) %>%
  # Add sample number 
  mutate(BIOTA_SAMPLENO = stringr::str_extract(Label, "(?<=\\s)[^\\s]+(?=$)") %>% as.numeric())

# head(dat[[5]])
# mutate(STATION_CODE = sub("-", "", STATION_CODE)) %>%
  

xtabs(~addNA(BIOTA_SAMPLENO) + addNA(STATION_CODE), dat[[5]])

ggplot(dat[[5]], aes(Parameter, Value)) +
  geom_boxplot() +
  facet_wrap(vars(Tissue), nrow = 1, scales = "free_y")

```

### Sheet 6 - Fat (type 2)    
```{r, message=FALSE, fig.width = 4, fig.height=4}
dat[[6]] <- read_excel_nilu2(filename, 
                             "Fett %", 
                             lessthans_given_as_negative_number = FALSE) %>%
  filter(Parameter != "Comments") %>%
  mutate(Group = "Fat",
         STATION_CODE = "19N")

# Remove one very high value (also see comment in sheet)
sel <- dat[[6]]$Tissue == "Ærfuglblod" & dat[[6]]$Value > 0.7; sum(sel)
dat[[6]]$Value[sel] <- NA

cat(sum(sel), "value(s) deleted for being very high \n")

# head(dat[[6]])
# dat[[6]] %>% View()

ggplot(dat[[6]], aes(Tissue, Value)) +
  geom_boxplot()

```


## Combine data  

#### Standardize NIVA_ID
Change NILU ID so they conform with Nivabase:
- NILU:     NIVA_ID   "Nr. 2018-09655"  
- Labware;  TEXT_ID     "NR-2018-09676"
```{r}

xtabs(~NIVA_ID, dat[[4]])[1:5]

for (i in 1:length(dat)){
  dat[[i]] <- dat[[i]] %>%
    mutate(TEXT_ID = sub(".*Nr.+ +", "NR-", NIVA_ID)) %>%
    mutate(TEXT_ID = sub("Nr", "NR", TEXT_ID)) %>%
    mutate(TEXT_ID = sub("NR.", "NR-", TEXT_ID, fixed = TRUE)) %>%
    mutate(TEXT_ID = sub("NR ", "NR-", TEXT_ID, fixed = TRUE)) %>%
    mutate(TEXT_ID = sub(" *2019", "2019", TEXT_ID)) %>%            # note: year-specific!
    mutate(TEXT_ID = sub("2019-", "2019-0", TEXT_ID, fixed = TRUE))             # note: year-specific!
}

xtabs(~TEXT_ID, dat[[4]])[1:5]

```

### Add BIOTA_SAMPLENO to data set 4 and 6   
```{r}

# check
if (FALSE){
  dat[[1]]$TEXT_ID %>% unique() 
  dat[[4]]$TEXT_ID %>% unique() 
}

lookup_BIOTA_SAMPLENO <- dat[[1]] %>%
  distinct(TEXT_ID, BIOTA_SAMPLENO)

# Should be zero!
n <- lookup_BIOTA_SAMPLENO %>%
  count(TEXT_ID) %>%
  filter(n > 1) %>%
  nrow()

if (n > 0)
  cat("WARNING! Some NIVA_ID have more thank one BIOTA_SAMPLENO! \n")

dat[[4]] <- dat[[4]] %>%
  left_join(lookup_BIOTA_SAMPLENO, by = "TEXT_ID")

dat[[6]] <- dat[[6]] %>%
  left_join(lookup_BIOTA_SAMPLENO, by = "TEXT_ID")


```


### Combine data 
```{r}

# Check column types  
get_coltypes <- function(dataframe){
  df <- dataframe %>% purrr::map_chr(class) %>% matrix(nrow = 1) %>% as.data.frame(stringsAsFactors = FALSE)
  colnames(df) <- colnames(dataframe)
  df
}
# get_coltypes(dat[[1]])
dat %>% purrr::map_df(get_coltypes)

#
# Combine data 1-6
#
dat_nilu1 <- bind_rows(dat)

#
# Save at this stage
#
saveRDS(dat_nilu1, "Data/161_data_nilu_niluformat.rds")


```

## Checks 

### Check that we have STATION_CODE and BIOTA_SAMPLENO for all     
```{r}

xtabs(~is.na(STATION_CODE) + Group, dat_nilu1)
xtabs(~is.na(BIOTA_SAMPLENO) + Group, dat_nilu1)

```

### Check that we have one measurement per sample and parameter  
```{r}

df_check <- dat_nilu1 %>%
  count(STATION_CODE, Tissue, BIOTA_SAMPLENO, Parameter) %>%
  filter(n > 1)
n <- nrow(df_check)

if (n > 0){
  cat("We have", n , "cases with more than one measurement per sample and parameter! Check 'df_check'!")
}

```


### Check sample numbers using 'df_samples'        

```{r}

df_check <- dat_nilu1 %>%
  left_join(df_samples %>% 
              select(TEXT_ID, BIOTA_SAMPLENO) %>% 
              rename(BIOTA_SAMPLENO_sample = BIOTA_SAMPLENO),
            by = "TEXT_ID") %>%
  filter(BIOTA_SAMPLENO_sample != BIOTA_SAMPLENO) 

n <- nrow(df_check)

if (n > 0){
  cat("We have", n , "cases where BIOTA_SAMPLENO does not fit with NIVA's sample data! Check 'df_check'!")
}

```


## Make data ready for joining to rest of data  
  
Desired format:  
```
head(dat_new3, 3)
  MYEAR STATION_CODE LATIN_NAME   SAMPLE_NO2 TISSUE_NAME PARAM  BASIS VALUE FLAG1 UNIT   
  <dbl> <chr>        <chr>             <dbl> <chr>       <chr>  <chr> <dbl> <chr> <chr>  
1  2019 15B          Gadus morhua          1 Lever       BD153  W     1.9   NA    UG_P_KG
2  2019 15B          Gadus morhua          1 Lever       BDE100 W     8     NA    UG_P_KG
3  2019 15B          Gadus morhua          1 Lever       CB157  W     0.029 NA    UG_P_KG
```


###  Units and tissues 
```{r}

# xtabs(~Unit, dat_nilu)
# xtabs(~UNIT, df_2018)  # df_2018 from script 01!

# Parameter UNIT added (old "Unit" is kept)
dat_nilu2 <- nilu_fix_units(dat_nilu1)

dat_nilu2 %>%
  count(Unit, UNIT)

# FOR check, in case there are cases with unit = NA
# dat_nilu2 %>%
#   filter(is.na(UNIT))

# xtabs(~Unit + UNIT, dat_nilu2)

# Not relevant this year - no HG data:
# dat_nilu2 <- nilu_fix_hg_unit(dat_nilu2)
  
cat("Number of NA values for UNIT: \n")
sum(is.na(dat_nilu2$UNIT))  # SHOULD BE ZERO

# Parameter TISSUE_NAME added (old "Tissue" is kept)
dat_nilu2 <- nilu_fix_tissue(dat_nilu2)

cat("\n Number of NA values for TISSUE_NAME: \n")
sum(is.na(dat_nilu2$TISSUE_NAME))  # SHOULD BE ZERO

# FOR check, in case there are cases with TISSUE_NAME = NA
# dat_nilu2 %>%
#   filter(is.na(TISSUE_NAME))

cat("\n Number of rows: \n")
nrow(dat_nilu2)  # 2610


```

### Parameters (PARAM)   
Originally Parameter, we set PARAM here ('Parameter' is kept)
```{r}
dat_nilu3 <- dat_nilu2
dat_nilu3$PARAM <- ""                        # Create PARAM
dat_nilu3 <- nilu_param_pcb_pbde(dat_nilu3)   # Set PARAM for PBCs and PBDEs (using IPUAC nr)
dat_nilu3 <- nilu_param(dat_nilu3)            # Set PARAM for the most of the rest

# Finally, Sum PCB
sel <- dat_nilu3$Parameter %in% "Sum 7 PCB"; sum(sel)
dat_nilu3$PARAM[sel] <- "CB_S7"

# All parameters
# dat_nilu3 %>% count(Parameter, PARAM) %>% View()

# Check remaining missing PARAM
dat_nilu3 %>% filter(PARAM == "") %>% count(Parameter, PARAM)    # none
dat_nilu3 %>% filter(is.na(PARAM)) %>% count(Parameter, PARAM)   # the rest we can drop

```

#### Filter data for PARAM  
I.e., remove LOQ and the PCB sums  
```{r}

dat_nilu3 <- dat_nilu3 %>% filter(!is.na(PARAM))

```


### Make ready for adding data to the rest of the data   
- Will be added in script 101 after section 4 - to 'dat_new3'  
- Add MYEAR, STATION_CODE, LATIN_NAME, BASIS   
- 
```{r}

dat_nilu4 <- dat_nilu3 %>%
  mutate(MYEAR = year,
         STATION_CODE = case_when(
           !is.na(STATION_CODE) ~ STATION_CODE,
           is.na(STATION_CODE) ~ station_code_not_given            # assume that all that isn't cod, is 
         ),
         LATIN_NAME = case_when(
           STATION_CODE %in% "19N" ~ "Somateria mollisima",
           TRUE ~ "Gadus morhua"
         ),
         BASIS = "W"
  ) %>%
  rename(SAMPLE_NO2 = BIOTA_SAMPLENO,
         VALUE = Value)

```


```{r}

needed_variables <- c("MYEAR", "STATION_CODE", "LATIN_NAME", "SAMPLE_NO2", "TISSUE_NAME", 
                      "PARAM", "BASIS", "VALUE", "FLAG1", "UNIT")

dat_nilu4[needed_variables]
```

```{r}

dat_nilu1 %>%
  xtabs(~addNA(Parameter), .)
dat_nilu3 %>%
  xtabs(~addNA(Parameter), .)

dat_nilu4 %>%
  xtabs(~addNA(PARAM), .)
```


## Test adding nilu data to original data   

### Check column names  
NILU data vs original data  
```{r}

n1 <- names(dat_nilu7)
n2 <- names(data_all)
n2[!n2 %in% n1]


```

## Save  
```{r}

saveRDS(dat_nilu7, "Data/161_data_nilu_nivaformat.rds")

```

