---
title: "172_Read_cod_bioleffect_data"
output: 
  html_document:
    keep_md: true
    toc: true
---

- The data will be added in script 101 `101_Combine_with_legacy_data_2019.Rmd`, after section 4 - to 'dat_new3'   
  
Desired format:  
```
head(dat_new3, 3)
  MYEAR STATION_CODE LATIN_NAME   SAMPLE_NO2 TISSUE_NAME PARAM  BASIS VALUE FLAG1 UNIT   
  <dbl> <chr>        <chr>             <dbl> <chr>       <chr>  <chr> <dbl> <chr> <chr>  
1  2019 15B          Gadus morhua          1 Lever       BD153  W     1.9   NA    UG_P_KG
2  2019 15B          Gadus morhua          1 Lever       BDE100 W     8     NA    UG_P_KG
3  2019 15B          Gadus morhua          1 Lever       CB157  W     0.029 NA    UG_P_KG
```


## Libraries
```{r Packages, include=FALSE}
library(readxl)
library(dplyr)
library(purrr)
library(tidyr)
```

## Set year
```{r}

year <- 2019  

```


## Data  

Note that    
    - "Seqno" is station (in the old Access base - see below)
    - "Subno" is equal to SPECIMEN_NO in the NIVADATABASE, and different from SUBNO in ICES  
```{r}

# dir("Input_data")
fn <- "Input_data/BEM-JAMP_2019_ARU.xls"
dat <- read_excel(fn, sheet = "Import_format", na = "n.d.")

cat("Number of lines read: \n")
nrow(dat)

```


### Seqno lookup table   
(Could just have looked at the 'stasjon' column in the Excel file)
```{r}

# How to find the connection from Seqno to STATION_CODE
# (commented out, replaced with hard-coded version below, ie 'df_seqno2station')

# fn <- "H:/Documents/seksjon 212/COCO/Analyser/Data_extraction/Milkys_Data_2017_01_17_cemp_tables.RData"
# load(fn)
# cemp_f2 %>%
#   count(seqno, jmpst) %>%
#   filter(seqno %in% unique(dat$Seqno)) %>% 
#   select(-n) %>%
#   dput()

# Edited, as some seqno has several station codes
df_seqno2station <- data.frame(
  Seqno = c(151, 231, 301, 531), 
  STATION_CODE = c("15B", "23B", "30B", "53B"),
  stringsAsFactors = FALSE)

### Add STATION_CODE test
# dat %>%
#   left_join(df_seqno2station, by = "Seqno") %>%
#   select(Seqno, stasjon, STATION_CODE, nummer) %>%
#   View()

```



## Put in correct format   
- The "O" parameters are the "OH" values normalized by AY380:   
    - PA1O = PA1OH/AY380   
    - PYR1O = PYR1OH/AY380   
    - BAP3O = BAP3OH/AY380  
```{r}

dat2 <- dat %>%
  # Add STATION_CODE
  left_join(df_seqno2station, by = "Seqno") %>%
  rename(SAMPLE_NO2 = nummer) %>%
  # Make unnormalized data:
  mutate(PA1OH = PA1O*AY380,
         PYR1OH = PYR1O*AY380,
         BAP3OH = BAP3O*AY380) %>%
  gather("PARAM", "VALUE", PA1O, PYR1O, BAP3O, PA1OH, PYR1OH, BAP3OH, ALAD, EROD) %>%
  # reshape on tall/narrow form:
  select(STATION_CODE, SAMPLE_NO2, AY380, PARAM, VALUE) %>%
  # add extra variables:
  mutate(MYEAR = year,
         LATIN_NAME = "Gadus morhua",
         TISSUE_NAME = 
           case_when(PARAM %in% c("PA1O", "PYR1O", "BAP3O", "PA1OH", "PYR1OH", "BAP3OH") ~ "Galle",
                     PARAM %in% "ALAD" ~ "Blod",
                     PARAM %in% "EROD" ~ "Liver - microsome"),
         UNIT =
           case_when(PARAM %in% c("PA1O", "PYR1O", "BAP3O", "PA1OH", "PYR1OH", "BAP3OH") ~ "UG_P_KG",
                     PARAM %in% "ALAD" ~ "ng/min/mg protein",
                     PARAM %in% "EROD" ~ "pmol/min/mg protein"),
         BASIS = "W",
         Preferred_unit = UNIT,
         Conversion_factor = 1,
         N_par = 1) %>%
  filter(!is.na(VALUE))

```



## Uncertainties (SD in percent)  
```{r}

dat2 <- dat2 %>%
  mutate(Uncertainty = 
           case_when(
             # OH-PAH i MILKYS 2018-gallepr?ver (003).xlsx
             PARAM %in% c("PA1O", "PA1OH") ~ 40/2,
             PARAM %in% c("PYR1O", "PYR1OH") ~ 50/2,
             PARAM %in% c("BAP3O", "BAP3OH") ~ 150/2,
             # ALA-D uncertainty.txt
             PARAM %in% c("ALAD") ~ 20,
             # Conservative guess
             PARAM %in% c("EROD") ~ 30
             )
         )
```

## LOQ (kind of!)  
We simply use the minimum value recorded  
```{r}
df_valuestatistics <- dat2 %>%
  group_by(PARAM) %>%
  summarise(Val_min = min(VALUE, na.rm = TRUE) %>% round(4),
            Val_med = median(VALUE, na.rm = TRUE) %>% round(4),
            Val_max = max(VALUE, na.rm = TRUE) %>% round(4)
  )
df_valuestatistics

# Add LOQ to data
dat3 <- dat2 %>%
  left_join(df_valuestatistics %>%
              select(PARAM, Val_min) %>%
              rename(QUANTIFICATION_LIMIT = Val_min),
            by = "PARAM") %>%
  mutate(FLAG1 = case_when(
    is.na(VALUE) ~ "<",
    !is.na(VALUE) ~ as.character(NA))
    )

```

### Add UNCRT and METCU  
We here use LOQ from less-thans as LOQ, if possible  
    - i.e. LOQ_lt_med, or QUANTIFICATION_LIMIT if LOQ_lt_med is not given
```{r}
dat3 <- dat3 %>%
  mutate(s = QUANTIFICATION_LIMIT/3,
         UNCRT = sqrt(s^2 + (VALUE^2)*(Uncertainty^2)),
         METCU = "SD"
         ) %>%
  select(-s)
```



## Saving data

### Check
```{r}

needed_variables <- c("MYEAR", "STATION_CODE", "LATIN_NAME", "SAMPLE_NO2", "TISSUE_NAME", 
                      "PARAM", "BASIS", "VALUE", "FLAG1", "UNIT")

dat3[needed_variables] %>% head()

```

### Save  
These data are used in script 101 `101_Combine_with_legacy_data_2019.Rmd` 
```{r}

saveRDS(dat3, "Data/172_cod_biological_effects_2019.rds")

#
# This goes into script 34
#

# dat3 <- readRDS("Data/33_data_biological_effects.rds")
```

#### FInal check
```{r}
# xxx
dat3 %>% 
  xtabs(~PARAM + is.na(UNCRT), .)

```
