---
title: "Making overview tables"
output: html_notebook
---


Based on script '44_make_overview_trend_table' in H:/Documents/seksjon 212/Milkys 2017/Analyse  
  
* Tables are written to folder 'Tables_for_report'  
* File namse 'version' correspond to 'big excel table' version numbers  
* Part 3-7: make "trend" tables (name 'Summarytable_trend') - also includes PROREF shading   
* Part 8-11: make "EQS" tables (name 'Summarytable_EQS')  
  
## 0. Load packages and functions
### Packages
```{r}
# library(tidyverse)
library(dplyr)
library(purrr)
library(ggplot2)
library(safejoin)
library(readxl)
library(xlsx)
```

### Functions used
```{r}
source("002_Utility_functions.R")
source("210_Make_overview_tables_functions.R")
```


### Year  
```{r}
current_year <- 2019
```

### Folder to write results  
```{r}
resultfolder <- "Tables_for_report"
```


## 1. Read 'big excel' data    
Make sure to read the most recent one!   
```{r}
datafolder <- "Big_excel_table"

fns <- list_files(datafolder, "Data_xl_[:date:].+", ".rds")  

fn_full <- paste0(datafolder, "/", fns[1])
data_xl_lessthans <- readRDS(fn_full)

cat("\n")
cat("Data read from the file \n", fn_full, "\n (", nrow(data_xl_lessthans), "lines )")

# Version number
version_no <- sub(datafolder, "", fn_full) %>% substr(10, 25)
cat("\n\n")
cat("Tables made by this script will be written to folder \n  ", 
    resultfolder,
    "\nand file names will end with the following version: \n  ",
    sQuote(version_no)
)

```
```{r}

data_xl_lessthans %>%
  filter(PARAM %in% "SCCP" & STATION_CODE %in% "24B" & Basis == "WW") %>%
  select(PARAM, STATION_CODE, Basis, Trends.2019)

```


### A couple of changes to the big excel data   
```{r}

# 1. Change parameter from "VDSI/Intersex" to "VDSI"  
#    In order to keep tables compact  

sel <- data_xl_lessthans$PARAM %in% c("VDSI/Intersex", "Intersex")
data_xl_lessthans$PARAM[sel] <- "VDSI"
cat("Changed", sum(sel), "records to PARAM = 'VDSI'")

# 2. VDSI (actually intersex) value of 71G in 2019
#    Given by pers.comm (mail from Lise mandag 28. september 2020 10:51)
sel <- with(data_xl_lessthans,
            STATION_CODE %in% "71G" &
            PARAM %in% "VDSI" & 
              Basis == "WW")
data_xl_lessthans$Yr_2019[sel] <- 0
cat("\n")
cat("Changed", sum(sel), "records to Yr_2019 = 0 for VDSI")

# 3. LATIN_NAME of 71G
sel <- with(data_xl_lessthans,
            STATION_CODE %in% "71G")
data_xl_lessthans$LATIN_NAME[sel] <- "N. lapillus / L. littorea"
cat("\n")
cat("Changed", sum(sel), "records to LATIN_NAME = 'N. lapillus / L. littorea'")


```

### Checks
```{r}

if (FALSE){

  # All column names
  data_xl_lessthans %>% colnames()
  
  data_xl_lessthans %>%
    count(Substance.Group)
  
  data_xl_lessthans %>%
    filter(Substance.Group %in% "Organo-metallic compounds") %>%
    count(PARAM, Parameter.Name, IUPAC, Component.Name)
  
  data_xl_lessthans %>%
    filter(PARAM %in% "TPhT" & Basis == "WW") %>%
    select(PARAM, STATION_CODE, LATIN_NAME, TISSUE_NAME, Yr_2018, Q95, EQS)
  
  data_xl_lessthans %>%
    filter(PARAM %in% "D5" & Basis == "WW") %>%
    select(PARAM, STATION_CODE, LATIN_NAME, TISSUE_NAME, Yr_2018, Q95, EQS)
  
  data_xl_lessthans %>%
    filter(!is.na(EQS)) %>%
    count(PARAM, Parameter.Name, IUPAC, Component.Name)
  
  data_xl_lessthans %>%
    filter(grepl("71A", STATION_CODE) & Basis == "WW") %>%
    select(PARAM, STATION_CODE, LATIN_NAME, TISSUE_NAME, Yr_2018, Yr_2019)
  
  data_xl_lessthans %>%
    filter(PARAM %in% "CD" & Basis == "WW" & !is.na(Yr_2019)) %>%
    select(PARAM, STATION_CODE, LATIN_NAME, TISSUE_NAME, Yr_2018, Q95, EQS) %>%
    count(STATION_CODE, LATIN_NAME, TISSUE_NAME) %>%
    arrange(LATIN_NAME, TISSUE_NAME)   # 33F = skrubbe
  
  data_xl_lessthans %>%
  filter(PARAM %in% "SCCP" & STATION_CODE %in% "24B" & Basis == "WW") %>%
  select(PARAM, STATION_CODE, Basis, Trends.2019)
  
}

```


## 2. Read other data  

### Station names
Used in 'plot_single_series_medians_data' (see '_functions' file)  
```{r}

#
# List of station names   
#
data_stations <- readxl::read_excel("Input_data/Kartbase_edit.xlsx")
cat("\n")
cat("List of station names read \n")

# Note that 
#     Positions are the nominal positions (doesn't change from year to year)
#     Station name for report are those originally called `Til Rapport` in 
#       Norman's sheet

# Check that STATION_CODE values are unique
check <- data_stations %>%
  group_by(STATION_CODE) %>%
  mutate(n = n()) %>%
  filter(n > 1)

if (nrow(check) > 0){
  stop("The same STATION_CODE occurs more than once! Check 'check'")
}  

df_stationnames <- data_stations %>%
  filter(!is.na(STATION_CODE)) %>%
  # Add 'Report_version_name'
  mutate(
    Report_version_name = paste0(Station_name, " (st. ", STATION_CODE, ")")
  )

# Fixing station 36A and 36A1 (36A1 shall count as 36A) 
sel <- df_stationnames$STATION_CODE %in% "36A1"
df_stationnames$STATION_CODE[sel] <- "36A"
cat(sum(sel), "lines with 36A1 changed to 36A \n")

# Fixing station 36A and 36A1 (36A1 shall count as 36A) 
sel <- df_stationnames$STATION_CODE %in% "227G2"
df_stationnames$STATION_CODE[sel] <- "227G"
cat(sum(sel), "lines with 227G2 changed to 227G \n")


if (FALSE){
  # Check some stations that might be missing
  df_stationnames %>% filter(STATION_CODE %in% c("28A2","97A3","227G","19B","19N","I714", "I133"))
}

```
### Parameter order   
```{r}

fn <- "Input_data/Lookup table - preferred parameter order.xlsx"
df_param_order <- read_excel(fn, sheet = "Order") %>%
  rename(PARAM = Analysevariabel)

cat("df_param_order:", nrow(df_param_order), "lines read \n")

# Remove duplicates (but don't change the order)
duplicates <- df_param_order %>%
  count(PARAM) %>%
  filter(n > 1) %>%
  pull(PARAM)

for (param in duplicates){
  sel <- df_param_order$PARAM == param
  sum(sel)
  remove <- 2:sum(sel)
  df_param_order <- df_param_order[-which(sel)[remove],]  # delete the second one
}

cat("after removing duplicates:", nrow(df_param_order), "lines left \n")

if (FALSE){
  # Check that it workd
  duplicates <- df_param_order %>%
    count(PARAM) %>%
    filter(n > 1)
}

```


## 3. Definitions  

### Define stations
Note: order counts!
```{r}
stations_table <- c(
  "30A", "I301", "I304", "31A", "36A", "I023", "I024",     # Blue mussel, Oslofjorden
  "71A", "I714", "76A2", "I131A", "I133", "15A",           # Blue mussel, Grenland - Sørlandet
  "51A", "52A", "56A", "57A", "64A", "65A", "22A",         # Blue mussel, Vestlandet - Trøndelag
     "I241", "26A2", "28A2", "91A2",                       #  - " -
  "97A2", "97A3", "98A2", "10A2", "11X",                   # Blue mussel, Nord-Norge
  "30B", "36B", "02B", "71B","13B", "15B",                 # Cod
    "53B", "23B", "24B", "28B", "80B", "96B",              # Cod            
    "98B1", "43B2", "45B2", "10B", "19B",                  # Cod
    "33F",                                                 # European flounder
  "19N",                                                   # Eider duck - twice (blood and egg)
  "71G",                                                   # Dog whelk / periwinkle
  "36G", "76G", "131G", "15G", "227G",                     # Dog whelk
    "22G", "98G", "11G")                                   # Dog whelk

# stations_table

# Get some line numbers for table - these are used later
a <- which(stations_table == "11X")    # last blue mussel
b <- which(stations_table == "33F")    # last cod and flounder
c <- which(stations_table == "19N")    # first bird    
d <- which(stations_table == "71G")    # first snail
e <- length(stations_table)            # last
```

### Define parameters  
Again: the order counts!   
Note: PAH sums are called PAHSS (= PAH16), P_S (= PAH16 minus naphtalene) and PK_S (= KPAH)  
```{r}

par_table <- c("AG", "CD", "CO", "CR", "HG", "NI", "PB",  
               "CB_S7", 
               "DDEPP",
               "HCB", "HBCDA",
               "BDE6S", "BDE47", "BDE100", "BDE209",
               "SCCP", "MCCP", 
               "P_S", "ANT", "BAA", "BAP", "FLU", "NAP",  
               "PFOA", "PFOS", "PFOSA", 
               "TBT", "TPT", 
               "VDSI",
               "D5")

```

## 4. Make stations_df (one line per station)
### Start by adding all stations
```{r}
stations_df <- data.frame(
  Station = c(
    stations_table[1:a],
    stations_table[(a+1):b],   #  cod + flounder liver
    stations_table[(a+1):b],   #  cod + flounder muscle
    stations_table[c],         # eider duck blood
    stations_table[c],         # eider duck eggs
    stations_table[d:e]),
  stringsAsFactors = FALSE)

# Set row numbers for the finished table
i_mussel <- 1:a
i_fish_liver <- (a+1):b
i_fish_muscle <- i_fish_liver + (b-a)
i_eider_blood <- tail(i_fish_muscle,1) + 1
i_eider_egg <-   tail(i_fish_muscle,1) + 2
i_snails <- seq(i_eider_egg + 1, nrow(stations_df))

```

### Add Species and Tissue
```{r}
stations_df$Species <- NA
stations_df$Species[stations_df$Station %in% stations_table[1:a]] <- "Mytilus edulis"
stations_df$Species[stations_df$Station %in% stations_table[(a+1):b]] <- "Gadus morhua"
stations_df$Species[stations_df$Station %in% "33F"] <- "Platichthys flesus"
stations_df$Species[stations_df$Station %in% stations_table[c]] <- "Somateria mollissima"
stations_df$Species[stations_df$Station %in% stations_table[d]] <- "N. lapillus / L. littorea"
stations_df$Species[stations_df$Station %in% stations_table[(d+1):e]] <- "Nucella lapillus"

# stations_df$Tissue <- NA
# stations_df$Tissue[stations_df$Station %in% stations_table[1:a]] <- "Whole soft body"
# stations_df$Tissue[i_fish_liver] <- "Lever"
# stations_df$Tissue[i_fish_muscle] <- "Muskel"
# stations_df$Tissue[i_eider_blood] <- "Blod"
# stations_df$Tissue[i_eider_egg] <- "Egg homogenate of yolk and albumin"
# stations_df$Tissue[i_snails] <- "Whole soft body"

stations_df$Tissue <- NA
stations_df$Tissue[stations_df$Station %in% stations_table[1:a]] <- "Whole soft body"
stations_df$Tissue[i_fish_liver] <- "Liver"
stations_df$Tissue[i_fish_muscle] <- "Muscle"
stations_df$Tissue[i_eider_blood] <- "Blood"
stations_df$Tissue[i_eider_egg] <- "Egg"
stations_df$Tissue[i_snails] <- "Whole soft body"
```

### Add Report_version_name
```{r}
stations_df <- left_join(stations_df, 
                         df_stationnames %>% select(STATION_CODE, Report_version_name),
                         by = c("Station" = "STATION_CODE"))

# Example
xtabs(~Report_version_name, stations_df)[1:6]

```

## 5. Make summaries_list  

### a. Add Name_english (see code for 'write_excel_summarytable()')  
We don't care about the flatfish here  
```{r}
df_english_name <- matrix(c(
  "Gadus morhua", "Cod",
  "Platichthys flesus", "Flounder",
  "Mytilus edulis", "Blue mussel",
  "Somateria mollissima", "Eider duck",
  "Nucella lapillus", "Dog whelk",
  "N. lapillus / L. littorea", "Dog whelk / periwinkle"), ncol = 2, byrow = TRUE) %>%
  as.data.frame()
colnames(df_english_name) <- c("LATIN_NAME", "Name_english") 

data_xl_lessthans <- data_xl_lessthans %>%
  safe_left_join(df_english_name, by = "LATIN_NAME", check = "CV")

data_xl_lessthans %>%
  count(LATIN_NAME, Name_english)

# colnames(data_xl_lessthans)

```

### b. Make list  
* One list element per station/species/tissue  
* Each element is a data frame of parameters  
```{r}
N <- nrow(stations_df)
summaries_list <- vector("list", N)
for (i in 1:N)
  summaries_list[[i]] <- get_summary_onestation(
    stations_df[i,"Station"], 
    stations_df[i,"Report_version_name"],
    stations_df[i,"Species"],
    stations_df[i,"Tissue"], 
    par_table = par_table, 
    data = data_xl_lessthans,       # input data
    varname_median = paste0("Yr_", current_year), 
    varname_trends = paste0("Trends.", current_year)
  )

# Check result 1:
# summaries_list[[1]] %>% head()
# Check result 2:
# i <- which(stations_df[,"Station"] == "36A1")
# i <- which(stations_df[,"Station"] == "24B" & stations_df[,"Tissue"] == "Liver")
# summaries_list[[i]] %>% select(PARAM, TISSUE_NAME, Yr_2018, EQSclass_2018, EQS, Trends.2019) %>% View()

rows_per_station <- 1:length(summaries_list) %>% 
  map_df(~tibble(i = ., STATION_CODE = stations_df$Station[.], nrow = nrow(summaries_list[[.]])))

# rows_per_station
if (sum(rows_per_station$nrow==0) > 0){
  message("Stations not found:")
  rows_per_station %>% 
    filter(nrow == 0) %>% 
    pull(STATION_CODE) %>%
    paste(collapse = ", ")
} else (
  cat("All stations found \n")
)

if (FALSE){
  # size of list:
  length(summaries_list)
  # names: first 5 elements of list:
  names(summaries_list[1:5])
  # first 5 elements of list:
  str(summaries_list[1:5], 1)
  
  i <- which(stations_df$Station == "30A")
  summaries_list[[i]] %>% select(PARAM, STATION_CODE, Basis, Yr_2019, EQS_2019) %>% View()
  
}

```

## 6. Write excel table  
```{r}

fn <- paste0("Summarytable_trend from R version ", 
             version_no, ".xlsx")

fn_full <- paste0(resultfolder, "/", fn)

write_excel_summarytable(summaries_list, 
                         fn_full, 
                         param_names = par_table,
                         varname_trend = paste0("Trends.", current_year),
                         varname_median = paste0("Yr_", current_year))

cat("File \n  ", fn, "\nwritten to folder \n  ", resultfolder)

```

## 7. What to do with this file in Excel  
* "Save as" and change name by replacing "from R" with "for report". Make sure you save it as an "xlsm" (macro-enabled) file
* Move the HG column for cod muscle to the cod liver part. Then delete the cod muscle rows and the TISSUE column
* Open the _Macros.xlsm work book  
* Go back to our sheet, select the main part of the table (the part containing the trends symbols)
* Go to menu Developer:Macros and select "_Macros.xlsm!Sheet1.Set_character_1_and_3_to_Wingdings_RANGE"
* Save file 


## 8. EQS table - Definitions

### Parameters  
Stations are the same
```{r}

# Deleted "DD_SS", "TCDDN", "TE_DFI05", "TE_CBNOI05"
par_table_eqs <- c("BDE6S", "BDE47", "SCCP", "MCCP", "DDEPP", 
  "HBCDA", "HCB", "HCBD", "HCHG", "HG", "4-N-NP", "4-T-NP", "4-N-OP", "4-T-OP", 
  "QCB", "PFOA", "PFOS", 
  "ANT", "BAA", "BAP", "FLU", "NAP", 
  "CB_S7", 
  "TBT", "TPT", 
  "D5")

# Order 'par_table_eqs' following 'df_param_order':
par_table_eqs <- par_table_eqs %>%
  factor(levels = df_param_order$PARAM) %>%
  sort()
# par_table_eqs

# Check that all PARAM have EQS
check <- par_table_eqs %in% unique(subset(data_xl_lessthans, !is.na(EQS))$PARAM)
mean(check)  # should be 1 (confirms that all PARAM are correctly spelled)
# par_table_eqs[!check]

if (sum(!check) > 0){
  cat("Parameters without EQS: \n  ")
  cat(par_table_eqs[!check] %>% paste(collapse = ", "), "\n")
  stop("EQS not found for ", sum(!check), " parameters \n")
} else {
  cat("EQS found for all parameters \n")
}

```

## 9. EQS table - Make summaries_list  
One list element per station/species/tissue, each element is a data frame of parameters  
```{r}

N <- nrow(stations_df)
summaries_list_eqs <- vector("list", N)
for (i in 1:N)
  summaries_list_eqs[[i]] <- get_summary_onestation(
    stations_df[i,"Station"], 
    stations_df[i,"Report_version_name"],
    stations_df[i,"Species"],
    stations_df[i,"Tissue"], 
    par_table = par_table_eqs, data = data_xl_lessthans,
    varname_median = paste0("Yr_", current_year), 
    varname_trends = paste0("Trends.", current_year)
  )


# Check result
# summaries_list_eqs[[65]] %>% head()
# summaries_list_eqs[[9]] %>% select(PARAM, STATION_CODE, LATIN_NAME, TISSUE_NAME, Yr_2019, Q95, EQS)

#
# Remove members of 'summaries_list_eqs' that have no data the current year  
# (in other words, delete stations with no data in the current year)
#
# function
get_n_param_current_year <- function(X){
  value_column <- paste0("Yr_", current_year)
  sum(!is.na(X[[value_column]]))
}
# test
# get_n_param_current_year(summaries_list_eqs[[9]])

# get number of parameters
n_param_current_year <- summaries_list_eqs %>% 
  map_int(get_n_param_current_year)

# keep those with data
summaries_list_eqs <- summaries_list_eqs[n_param_current_year > 0]

#
# Are all stations in stations_df found?
#
rows_per_station <- 1:length(summaries_list_eqs) %>% 
  map_df(~tibble(i = ., STATION_CODE = stations_df$Station[.], nrow = nrow(summaries_list_eqs[[.]])))
# rows_per_station
if (sum(rows_per_station$nrow==0) > 0){
  message("Stations not found:")
  rows_per_station %>% filter(nrow == 0)
} else (
  cat("All stations found \n")
)

```


## 10. Write excel table  
```{r}
# Change 'par_table_eqs' from factor to character 
par_table_eqs <- levels(par_table_eqs)[as.numeric(par_table_eqs)]

fn <- paste0("Summarytable_EQS from R version ", 
             version_no, ".xlsx")

fn_full <- paste0(resultfolder, "/", fn)
# fn_full <- paste0(resultfolder, "/", "test_EQS.xlsx")

write_excel_summarytable_proref_eqs(
  summaries_list_eqs, 
  fn_full, 
  param_names = par_table_eqs,
  varname_median = paste0("Yr_", current_year)
  )

cat("File \n  ", fn, "\nwritten to folder \n  ", resultfolder)

```


## 11. What to do with this file in Excel
* "Save as" and change name by replacing "from R" with "for report"
* Move the HG column for cod muscle to the cod liver part. Then delete the cod muscle rows



