---
title: "252_Check_increasing_LOQ_effects"
output: 
  html_document:
    toc: true
    toc_float: true  
    keep_md: true
---   

Background: 
* LOQ has increased for PCB (in 2017), silver (2019) and napthalene (2019)   
* This leads to 'false' increasing trends due to the method of selecting random numbers between LOQ/2 and LOQ   
* Here we check those trends more carefully  

```{r}
knitr::opts_chunk$set(echo = FALSE)  
```

```{r, results='hide', message=FALSE, warning=FALSE, echo = FALSE}

## 1. Libraries and functions  

# install.packages("lubridate")

library(dplyr)
library(purrr)
library(lubridate)
library(readxl)
library(mgcv)
library(AICcmodavg)   # AICc()
library(ggplot2)
library(safejoin)     # installed from https://github.com/moodymudskipper/safejoin   
source("001_Add_trends_functions.R")  # Copied from '16_Trend_functions.R' in Milkys_2018
source("252_Check_increasing_LOQ_effects_functions.R")

#
# Define a couple of extra functions for shortening code
#
get_stats <- function(df){
  calc_models_one_station2(df, gam = TRUE, log = TRUE)$statistics_for_file
}

model_from_medians_stat <- function(...)
  model_from_medians(...)$statistics

```


## 1. Read data

### a. Main data  
Read and reformat the most recent data (by default)  
```{r, results='hold', echo = FALSE}

# If we have NOT length-adjusted the last year's data:
filepattern <- "101_data_updated_"         # entire file name except date and extension

# Normally, if we have length-adjusted the last year's data:
# filepattern <- "109_adjusted_data_"       # entire file name except date and extension

filenumber <- 1                           # filenumber = 1 means "read the newest file"

files <- dir("Data", pattern = filepattern) %>% rev()

data_list <- read_rds_file("Data",
                     files, 
                     filenumber = filenumber,   # "1" gets the newest file   
                     get_date = TRUE, time_since_modified = TRUE)

data_all <- data_list$data
file_date <- data_list$file_date   
# 'file_date' will be used in part 10, when we save the resulting file

```

### b. Medians per year   
Made in script 110  
```{r, results='hold', echo = FALSE}

filepattern <- "110_mediandata_updated_"         # entire file name except date and extension
filenumber <- 1                           # filenumber = 1 means "read the newest file"

files <- dir("Data", pattern = filepattern) %>% rev()

data_list <- read_rds_file("Data",
                     files, 
                     filenumber = filenumber,   # "1" gets the newest file   
                     get_date = TRUE, time_since_modified = TRUE)

data_med <- data_list$data%>%
  rename(Proref_median = Median,
         Median = Value) 
  
cat("File date text:", data_list$file_date, "\n")

```

### c. Time trend results   
```{r, results='hold', echo = FALSE}

last_year <- 2020
date_pattern <- "[0-9][0-9][0-9][0-9]-[0-9][0-9]-[0-9][0-9]"
version_pattern <- "run[0-9][0-9]"

pattern_10yr_last <- paste0("^120_result_10yr_", last_year, "_", 
                            date_pattern, "_", version_pattern, ".rds$")
pattern_long_last <- paste0("^120_result_long_", last_year, "_", 
                            date_pattern, "_", version_pattern, ".rds$")

# dir("Data", pattern = pattern_10yr_last)
# dir("Data", pattern = pattern_10yr_last)

result_10yr <- list_and_read_rds_file(
  folder = "Data", 
  pattern = pattern_10yr_last, 
  check_date = data_list$file_date
)
result_long <- list_and_read_rds_file(
  folder = "Data", 
  pattern = pattern_long_last, 
  check_date = data_list$file_date
)


```

## 2. PCB 7  

**Conclusion: all upward short.time (10 year) trends should be marked as dubious, except 26A2 and 30A**     

* Part c indicates that 26A2, 30A and 22A have 'robust' increase     
* However in part d-f we test each station (using the sum of CB118, CB138 and CB153), and we reject the increase of 22A, leaving us with 26A2 and 30A as having 'robust increasing trend'  

### a. Plots  

```{r, fig.width = 7, fig.height=5}

ser_plot_cb("30A")
ser_plot_cb("I023")
ser_plot_cb("22A")
ser_plot_cb("26A2")
ser_plot_cb("11X")
ser_plot_cb_raw("11X")

```


```{r, results = 'hide'}

## Short-term trends  

ser_trend("CB_S7", "30A") %>%
  select(STATION_CODE, PARAM, N, Nplus, Model_used, P_change, Dir_change)
  
ser_trend_cb10yr("30A") %>%
  select(STATION_CODE, PARAM, N, Nplus, Model_used, P_change, Dir_change)
  
ser_trend_cb10yr("I023") %>%
  select(STATION_CODE, PARAM, N, Nplus, Model_used, P_change, Dir_change)

ser_trend_cb10yr("22A") %>%
  select(STATION_CODE, PARAM, N, Nplus, Model_used, P_change, Dir_change)

```

### b. Stations given with cb7 increase
```{r}

trends_sum <- result_10yr %>%
  filter(Basis == "WW" & PARAM %in% "CB_S7" & STATION_CODE %in% stations_bluemussel) %>%
    mutate(STATION_CODE = factor(STATION_CODE, levels = stations_bluemussel)) %>%
    arrange(as.numeric(STATION_CODE))

stations_with_cb7_increase <- trends_sum %>%
  filter(Dir_change == "Up") %>%
  pull(STATION_CODE) %>%
  forcats::fct_drop()

cat("Stations with PCB7 short-term increase (original): \n")
stations_with_cb7_increase

```

### c. Stations with 'robust' cb7 increase   
* For each CB parameter, define 'robust' increase as 'increase when there are at least 2 years among 2017-2020 that has >50% over LOQ'     
* For each station, define 'robust' cb7 increase as having at least 2 parameters with robust increase   
* 26A2, 30A and 22A are selected as 'robust' increase     
* However in the next parts, we reject the increase of 22A, leaving us with 26A2 and 30A  
```{r}

df_50perc_overLOQ <- data_med %>%
  filter(Basis == "WW" & PARAM %in% cb7 & STATION_CODE %in%   stations_with_cb7_increase & MYEAR >= 2017) %>%
  mutate(Prop_over_LOQ = Over_LOQ/N_median) %>%
  group_by(STATION_CODE, PARAM) %>%
  summarise(Yrs_50perc_overLOQ = sum(Prop_over_LOQ >= 0.5), # no. years with at least 50% over LOQ
            .groups = "drop")   

# ser_trend_cb10yr(stations_with_cb7_increase) %>% count(STATION_CODE)
# df_50perc_overLOQ %>% count(STATION_CODE)

check_cb10yr <- ser_trend_cb10yr(stations_with_cb7_increase) %>%
  select(STATION_CODE, PARAM, N, Nplus, Model_used, P_change, Dir_change) %>%
  left_join(df_50perc_overLOQ, by = c("STATION_CODE", "PARAM")) %>%
  mutate(Dir_change2 = case_when(
    Yrs_50perc_overLOQ >= 2 ~ Dir_change,   # Dir_change 'counts' if >=2 years with at least 50% over LOQ 
    Yrs_50perc_overLOQ < 2 ~ "")
    )

check_cb10yr_summ <- check_cb10yr %>%
  group_by(STATION_CODE) %>%
  summarise(
    N_param_increase = sum(Dir_change2 == "Up"),
    Params_increasing = paste(PARAM[Dir_change2 == "Up"], collapse = "; ")) %>%
  mutate(`Robust increase` = case_when(
    N_param_increase >= 2 ~ "Yes",
    N_param_increase < 2 ~ "No")
    ) %>%
  arrange(desc(N_param_increase))

check_cb10yr_summ                
  
```

### d. 26A2  
```{r}

station <- "26A2"

tab <- data_med %>%
  filter(Basis == "WW" & PARAM %in% cb7 & STATION_CODE %in% station & MYEAR >= 2009) %>%
  mutate(Prop_over_LOQ = Over_LOQ/N_median) %>%
  filter(Prop_over_LOQ >= 0.5) %>%
  xtabs(~MYEAR + PARAM, .)

tab

tab2 <- apply(tab[,c("CB118", "CB138", "CB153")], 1, sum) 
pick_years <- tab2[tab2 == 3] %>% names() %>% as.numeric()

data_select <- data_med %>%
  filter(Basis == "WW" &
           PARAM %in% c("CB118", "CB138", "CB153") & 
           STATION_CODE %in% station & 
           MYEAR %in% pick_years)

data_to_test <- data_select %>%
  group_by(STATION_CODE, LATIN_NAME, TISSUE_NAME, MYEAR, 
           N_median, Over_LOQ, Basis) %>%
  summarise(Median = sum(Median)) %>%
  mutate(PARAM = "CB118_138_153")

years_short <- seq(max(pick_years)-9, max(pick_years))

# debugonce(model_from_medians)
model_from_medians("CB118_138_153", "Mytilus edulis", "Whole soft body", 
                   station, "WW", years_short, data_to_test, 
                   plotname = "window", ggplot = TRUE)$statistics %>%
  select(Nplus, p_linear:Status)

```



### e. 30A  
```{r}

station <- "30A"

tab <- data_med %>%
  filter(Basis == "WW" & PARAM %in% cb7 & STATION_CODE %in% station & MYEAR >= 2009) %>%
  mutate(Prop_over_LOQ = Over_LOQ/N_median) %>%
  filter(Prop_over_LOQ >= 0.5) %>%
  xtabs(~MYEAR + PARAM, .)

tab

tab2 <- apply(tab[,c("CB118", "CB138", "CB153")], 1, sum) 
pick_years <- tab2[tab2 == 3] %>% names() %>% as.numeric()

data_select <- data_med %>%
  filter(Basis == "WW" &
           PARAM %in% c("CB118", "CB138", "CB153") & 
           STATION_CODE %in% station & 
           MYEAR %in% pick_years)

data_to_test <- data_select %>%
  group_by(STATION_CODE, LATIN_NAME, TISSUE_NAME, MYEAR, 
           N_median, Over_LOQ, Basis) %>%
  summarise(Median = sum(Median)) %>%
  mutate(PARAM = "CB118_138_153")

years_short <- seq(max(pick_years)-9, max(pick_years))

# debugonce(model_from_medians)
model_from_medians("CB118_138_153", "Mytilus edulis", "Whole soft body", 
                   station, "WW", years_short, data_to_test, 
                   plotname = "window", ggplot = TRUE)$statistics %>%
  select(Nplus, p_linear:Status)

```



### f. 22A  
```{r}

station <- "22A"

tab <- data_med %>%
  filter(Basis == "WW" & PARAM %in% cb7 & STATION_CODE %in% station & MYEAR >= 2009) %>%
  mutate(Prop_over_LOQ = Over_LOQ/N_median) %>%
  filter(Prop_over_LOQ >= 0.5) %>%
  xtabs(~MYEAR + PARAM, .)

tab

tab2 <- apply(tab[,c("CB118", "CB138", "CB153")], 1, sum) 
pick_years <- tab2[tab2 == 3] %>% names() %>% as.numeric()

data_select <- data_med %>%
  filter(Basis == "WW" &
           PARAM %in% c("CB118", "CB138", "CB153") & 
           STATION_CODE %in% station & 
           MYEAR %in% pick_years)

data_to_test <- data_select %>%
  group_by(STATION_CODE, LATIN_NAME, TISSUE_NAME, MYEAR, 
           N_median, Over_LOQ, Basis) %>%
  summarise(Median = sum(Median)) %>%
  mutate(PARAM = "CB118_138_153")

years_short <- seq(max(pick_years)-9, max(pick_years))

# debugonce(model_from_medians)
model_from_medians("CB118_138_153", "Mytilus edulis", "Whole soft body", 
                   station, "WW", years_short, data_to_test, 
                   plotname = "window", ggplot = TRUE)$statistics %>%
  select(Nplus, p_linear:Status)

```




## 3. Silver  

**Conclusion: all upward short-time (10 year) trends (there is only one) should be marked as dubious**  

### a. Plots  

```{r, fig.width = 7, fig.height=4}

ser_plot_med("AG", "31A")
ser_plot_med("AG", "30A")
ser_plot_med("AG", "22A")
ser_plot_med("AG", "26A2")
ser_plot_med("AG", "11X")

```


### b. Stations given with increase
```{r}

param <- "AG"

trends <- result_10yr %>%
  filter(Basis == "WW" & PARAM %in% param & STATION_CODE %in% stations_bluemussel) %>%
    mutate(STATION_CODE = factor(STATION_CODE, levels = stations_bluemussel)) %>%
    arrange(as.numeric(STATION_CODE))

stations_with_increase <- trends %>%
  filter(Dir_change == "Up") %>%
  pull(STATION_CODE) %>%
  forcats::fct_drop()

cat("Stations with", param, "short-term increase (original): \n")
stations_with_increase

```

### c. Test stations with 'robust' increase   
* Regression for 30A for 2009-2018      
```{r}

years_short <- (2018-9):2018

# debugonce(model_from_medians)
model_from_medians("AG", "Mytilus edulis", "Whole soft body", 
                   "31A", "WW", years_short, data_med, plotname = "window", ggplot = TRUE)$statistics %>%
  select(Nplus, p_linear:Status)

```


## 4. Naphtalene    

**Conclusion: no upward trends given - no need to change anything**  

* Obvious from the plots that the 2019 data are not very useful and if there had been a trend, it would have been dubious  


### a. Plots  

```{r, fig.width = 7, fig.height=4}

ser_plot_med("NAP", "30A")
ser_plot_med("NAP", "98A2")
ser_plot_med("NAP", "I023")

```


### b. Stations given with short-term increase   
```{r}

param <- "NAP"

trends <- result_10yr %>%
  filter(Basis == "WW" & PARAM %in% param & STATION_CODE %in% stations_bluemussel) %>%
    mutate(STATION_CODE = factor(STATION_CODE, levels = stations_bluemussel)) %>%
    arrange(as.numeric(STATION_CODE))

stations_with_increase <- trends %>%
  filter(Dir_change == "Up") %>%
  pull(STATION_CODE) %>%
  forcats::fct_drop()

cat("Stations with", param, "short-term increase (original): \n")
stations_with_increase

```


### c. Stations given with long-term increase   
```{r}

param <- "NAP"

trends <- result_long %>%
  filter(Basis == "WW" & PARAM %in% param & STATION_CODE %in% stations_bluemussel) %>%
    mutate(STATION_CODE = factor(STATION_CODE, levels = stations_bluemussel)) %>%
    arrange(as.numeric(STATION_CODE))

stations_with_increase <- trends %>%
  filter(Dir_change == "Up") %>%
  pull(STATION_CODE) %>%
  forcats::fct_drop()

cat("Stations with", param, "long-term increase (original): \n")
stations_with_increase

```