---
title: "Trend figures TEST"
output: html_notebook
---



## 0. Lib and source
```{r}
library(dplyr)
library(purrr)
library(ggplot2)
library(lubridate)
library(openxlsx)
library(mgcv)
library(AICcmodavg)   # AICc()
# font_import(pattern = "Wingdings.ttf")    # if not already installed on your PC
```

### Functions used  
This loads the functions we need
test fra BBE
```{r}
# source("16_Trend_functions.R")
source("001_Add_trends_functions.R")
source("401 Plot time series functions.R", encoding = "UTF-8")
```


## 1. Read data
### a1. Median data produced by script 11    
```{r}

files <- dir("Data", pattern = "110_mediandata_updated_") %>% rev()

cat("Reading the last file downloaded:")
cat("\n", files[1])
cat("\n")
cat("If you want to read a different file, replace 'files[1]' with the file you want")
cat("\n")

filename <- paste0("Data/", files[1]) 

# Medians
data_med2 <- readRDS(file = filename) %>%
  rename(Proref_median = Median,
         Median = Value) 

```


### a2. Add VDSI (Actually Intersex) for 71G 2017   
Is only performed if it is needed.   
```{r}
check <- sum(with(data_med2, PARAM %in% "VDSI/Intersex" & MYEAR == 2018 & STATION_CODE %in% "71G" & Basis %in% "WW"))

check

if (check == 0){   # add only if not already added
  # Copy line to use
  data_extra <- data_med2 %>% 
    filter(PARAM %in% "VDSI" & MYEAR == 2016 & STATION_CODE %in% "71G" & Basis %in% "WW")
  # Change that line
  data_extra$MYEAR <- 2018
  data_extra$Median <- 0    # manual entry of value 
  data_med2 <- rbind(data_med2, data_extra)
  rm(data_extra)
}

```

```{r}
# library(extrafont)
# font_import(pattern = "Wingdings.ttf")

```

### a3. Read station names  
Used in 'plot_single_series_medians_data' (see '_functions' file)  
```{r}
data_stations <- readxl::read_excel("Input_data/Kartbase.xlsx")

# Note that 
#     the current station code is 'stasjonskode', not "STATION_CODE"
#     `catch LAT__1` and `catch LONG__1` are currect (planned?) positions
#     the current station name is 'stasjonsnavn', not "STATION_CODE"
#     but station name for report is `Til Rapport`
#         station name for maps are `Til Kart`

# Check the two station code columns
# data_stations %>%
#   filter(STATION_CODE != stasjonskode)    # zero rows

df_stationnames <- data_stations %>%
  select(stasjonskode, `catch LAT...40`, `catch LONG...41`, stasjonsnavn, `Til Rapport`) %>%
  rename(STATION_CODE = stasjonskode,
         Lat = `catch LAT...40`, Long = `catch LONG...41`, 
         Station_name = stasjonsnavn,
         Report_version_name = `Til Rapport`) %>%
  filter(!is.na(STATION_CODE)) %>%
  filter(!Station_name %in% "Risøy, Østerfjord")  # One duplicate (different names), we just remove 

# Check stations that might be missing
df_stationnames %>% filter(STATION_CODE %in% c("28A2","97A3","227G","19B","19N"))

# Add missing station
df_stationnames <- df_stationnames %>%
  bind_rows(
    tibble(STATION_CODE = "227G",
           Report_version_name = "Flatskjær (St. 227G)")
    )

# Check some other stations
df_stationnames %>% filter(STATION_CODE %in% c("I714", "I133"))


```

### a4. Parameter names for graphs
Used in 'plot_single_series_medians_data' (see '_functions' file)  
```{r}
df_paramnames <- readxl::read_excel("Input_data/Lookup table - parameter names for plots.xlsx")

```

### a5. Species names for graphs
Used in 'plot_single_series_medians_data' (see '_functions' file)  

```{r}
df_speciesnames <- tibble(
  LATIN_NAME = c("Gadus morhua", "Mytilus edulis", "Nucella lapillus"),
  Species_name = c("Cod", "Blue mussel", "Dog whelk")
  )

```

### a6. Big excel data - file versions   
```{r}
#
# Existing file names (dates and versions)
#
fns <- dir("Big_excel_table", pattern = "Data_xl_.+.rds") %>% rev()
fns_date <- substr(fns, 9, 18)
fns_ver <- stringr::str_extract(fns, "ver[0-9]+") %>% stringr::str_extract("[0-9]+") %>% as.numeric()
file_versions <- tibble(Filename = fns, Date = fns_date, Version = fns_ver) %>%
  arrange(desc(Date), desc(Version))
file_versions

```

### a7. Big excel data - file versions   
```{r}

# Picks the first 'Filename' in 'file_versions' - this should be the latest 'excel' file:
data_xl_name <- file_versions$Filename[1]
data_xl_lessthans <- readRDS(paste0("Big_excel_table/", data_xl_name))

cat("Read big excel table", sQuote(data_xl_name))

# for inspection:
if (FALSE){
  data_xl_lessthans %>% 
    select(PARAM, LATIN_NAME, TISSUE_NAME, STATION_CODE, Basis, Yr_2019, EQS_2019, Lt_2019, Trends.2019) %>% 
    View()
}


```



## 2. Plot in document    

```{r}
type <- "line"
```


### From '401 Plot time series functions.R'   
Used for testing parameters in function
```{r}
#
# Saving plots
# 

save_trendplot <- function(plot_single_result, folder, suffix = "", windows = FALSE){
  fn <- paste0(folder, "/", plot_single_result$fn)
  # Add suffix to file name (note: works only on png and jpg!)
  fn <- sub(".png", paste0(suffix, ".png"), fn)
  fn <- sub(".jpg", paste0(suffix, ".jpg"), fn)
  if (windows){
    ggsave(fn, plot_single_result$gg, height = 4.4, width = 5.6, dpi = 450)
  } else {
    gg <- plot_single_result$gg +
      theme(axis.text = element_text(size = 11),
            axis.title = element_text(size = 12),
            plot.title = element_text(size = 13))
    png(fn, height = 4.4*1.2, width = 5.6*1.2, units = "in", res = 450)
    print(gg)
    dev.off()
  }
  invisible(fn)   # returns file name invisibly
}

# save_trendplot(X, folder_test, suffix = "_3")  # save as file

```

### Test 1
```{r}

# dir.create("Figures_401_test")
folder_test <- "Figures_401_test"

ymax <- 0.45
X <- plot_medians_and_trends(c("HG", "Gadus morhua", "Muskel", "30B",  "WW"), 
                             eqs_type = "line", xlim = c(1980, 2020), ylim = c(0, ymax))
X                               # show below (in this Rmd file)
#save_trendplot(X, folder_test)  # save as file
save_trendplot(X, folder_test, suffix = "_3")  # save as file

ymax <- 18
X <- plot_medians_and_trends(c("CU", "Gadus morhua", "Lever", "53B",  "WW"), 
                           eqs_type = "line", xlim = c(1986, 2020), ylim = c(0, ymax))
X                               # show below (in this Rmd file)
save_trendplot(X, folder_test, suffix = "_3")  # save as file

#plot_medians_and_trends(c("QCB", "Mytilus edulis", "Whole soft body", "I133A",  "WW"), 
 #                            eqs_type = "line", xlim = c(1988, 2019), ylim = c(0, 10.5))

#X <- plot_medians_and_trends(c("OCS", "Gadus morhua", "Lever", "30B",  "WW"), 
                        #    eqs_type = type, xlim = c(1990, 2019), eqs_x = 1990, ylim = c(0, 15.0))
#save_trendplot(X, folder)
```

### Test 2
```{r}
ymax <- 0.45
X <- plot_medians_and_trends(c("HG", "Gadus morhua", "Muskel", "23B",  "WW"), 
                             eqs_type = type, xlim = c(1980, 2020), ylim = c(0, ymax))
save_trendplot(X, folder_test)  # save as file
X

```
### Test 3
```{r}
X <- plot_medians_and_trends(c("HG", "Mytilus edulis", "Whole soft body", "30A",  "WW"), 
                             eqs_type = type, xlim = c(2006, 2020), eqs_x = 2006, ylim = c(0, 0.05))
save_trendplot(X, folder_test)  # save as file
X

```

### Test 4
```{r}
# debugonce(plot_medians_and_trends)
# debugonce(get_plotdata)
X <- plot_medians_and_trends(c("CB118", "Gadus morhua", "Lever", "30B",  "WW"), 
                             eqs_type = type, xlim = c(1984, 2020), ylim = c(0, 800))
save_trendplot(X, folder_test)  # save as file
X


```
### Test 5
```{r}
# debugonce(plot_medians_and_trends)
X <- plot_medians_and_trends(c("VDSI", "Nucella lapillus", "Whole soft body", "36G",  "WW"), 
                             eqs_type = type, xlim = c(1984, 2020))
save_trendplot(X, folder_test)  # save as file
X

```

### Test 6 - square symbol  
```{r}
X <- plot_medians_and_trends(c("OCS", "Gadus morhua", "Lever", "30B",  "WW"), 
                             eqs_type = type, xlim = c(1984, 2020), ylim = c(0, 15))

save_trendplot(X, folder_test)  # save as file
X

#X <- plot_medians_and_trends(c("OCS", "Gadus morhua", "Lever", "30B",  "WW"), 
 #                            eqs_type = type, xlim = c(1984, 2019), ylim = c(0, 15))
# str(X, 1)
# X$fn <- "TSplot_OCS_Gadus morhua_Lever_30B_WW_TEST3.png"
# save_trendplot(X, "Figures_401/Til rapporten")


```
