---
title: "Report for parameter (430)"
author: "DHJ"
date: "18.08.2022"
params:
  param: "HG"
output: 
  html_document:
  keep_md: false
toc: true
toc_float: true
---
   
## 1. Settings    

```{r 01_settings}

current_year <- 2021
knitr::opts_chunk$set(echo = FALSE, results = 'hold')

```

## 2. Packages  

```{r 02_packages, results='hide', message=FALSE, warning=FALSE}

library(dplyr)
# library(tidyr)
library(ggplot2)
library(lubridate)
library(flextable)
# library(glue)
library(readxl)
library(purrr)
# library(leftcensored)   # DHJ's package (https://github.com/DagHjermann/leftcensored)
library(scico)            # colour palettes incl. "Vik" (https://github.com/thomasp85/scico)

library(ggiraph)
library(cowplot)

# library(safejoin) # https://github.com/moodymudskipper/safejoin

# source("01_Get_chemical_data_NIVAbasen_functions.R")  # for get_standard_parametername()
# source("824_Report_for_scientists_functions.R")
source("420_Report_paramgroup_functions.R", encoding = "UTF-8")

source("125_Calculate_trends_leftadjusted_functions.R")

```


## 3. Get data  

```{r 03_get_data, warning=FALSE}

chem_params <- params$param

# debugonce(get_data_tables)
dat_list <- get_data_tables(param = params$param)

str(dat_list, 1)

# Put raw dat on long format  
dat_raw <- dat_list$data %>%
  select(PARAM, STATION_CODE, LATIN_NAME, TISSUE_NAME, MYEAR, 
         VALUE_WW, VALUE_DW, VALUE_FB, VALUE_WWa, VALUE_DWa, VALUE_FBa) %>%
  tidyr::pivot_longer(
    cols = c(VALUE_WW, VALUE_DW, VALUE_FB, VALUE_WWa, VALUE_DWa, VALUE_FBa), 
    names_to = "Basis", 
    values_to = "Value") %>%
  mutate(Basis = sub("VALUE_", "", Basis, fixed = TRUE))

# Add min and max to 'dat_list$medians'  
df_min_max <- dat_raw %>%
  group_by(PARAM, STATION_CODE, LATIN_NAME, TISSUE_NAME, Basis, MYEAR) %>%
  summarise(Value_min = min(Value, na.rm = TRUE),
            Value_max = max(Value, na.rm = TRUE), .groups = "drop"
            )

n1 <- nrow(dat_list$medians)
dat_list$medians <- dat_list$medians %>%
  left_join(df_min_max, by = c("MYEAR", "STATION_CODE", "LATIN_NAME", "TISSUE_NAME", "PARAM", "Basis"))
n2 <- nrow(dat_list$medians)

if (n2 != n1){
  stop("Check uniqueness")
}

# View(dat_list$data)
# View(dat_list$medians)
# View(dat_list$medians)
# View(dat_list$lookup_eqs)

```

## 4. Medians tile plot  

### Data and test plot  
```{r}

species <- "Mytilus edulis"
tissue <- "Whole soft body"
basis <- "WW"

if (nrow(dat_list$lookup_eqs) == 1){
  dat_list$lookup_eqs$LATIN_NAME <- species
}

# Select data
dat_medians_01 <- dat_list$medians %>%
  filter(LATIN_NAME %in% species,
         TISSUE_NAME %in% tissue,
         Basis %in% "WW") %>%
  # Must have data in 'current_year'  
  group_by(STATION_CODE) %>%
  mutate(Series_last_year = max(MYEAR)) %>%
  filter(Series_last_year %in% current_year)

station_order <- dat_list$lookup_stations %>%
  filter(STATION_CODE %in% unique(dat_medians_01$STATION_CODE)) %>%
  mutate(Station = paste(STATION_CODE, Station_name)) %>%
  select(-Art) %>%
  arrange(Order)

dat_medians_02 <- dat_medians_01 %>%
  # Add EQS
  left_join(dat_list$lookup_eqs, 
            by = c("LATIN_NAME", "PARAM", "Basis"))

dat_medians_03 <- dat_medians_02 %>%
  # Add station name
  left_join(
    dat_list$lookup_stations %>% 
      filter(LATIN_NAME %in% species) %>%
      select(STATION_CODE, Station_name, Region),
    by = "STATION_CODE"
  ) %>%
  mutate(
    Station = factor(
      paste(STATION_CODE, Station_name),
      levels = rev(station_order$Station)),
    Proref_ratio = Value/Proref,
    EQS_ratio = Value/EQS,
    Above_EQS = case_when(
      Value > EQS ~ "Over",
      Value <= EQS ~ "Under",
      TRUE ~ as.character(NA)),
    Prop_underLOQ = 1 - Over_LOQ/N_median,
    FLAG1 = case_when(
      Prop_underLOQ < 0.5 ~ as.character(NA),
      Prop_underLOQ >= 0.5 ~ "<"),
    LOQ_label = ifelse(Prop_underLOQ >= 0.5, "<", "")
  ) %>%
  filter(!is.na(Station))

# Add 'Proref_ratio_cut'  
proref_max <- max(dat_medians_03$Proref_ratio, na.rm = TRUE)
dat_medians <- dat_medians_03 %>%
  mutate(
    Proref_ratio_cut = cut(Proref_ratio, breaks = c(0,0.5,0.75,0.9,1,2,3,5,10, proref_max))
  )

ggplot(dat_medians, aes(MYEAR, Station, fill = Proref_ratio_cut)) +
  geom_tile()


```

### Plot
```{r, fig.width=10, fig.height=8}

proref_cols <- c(RColorBrewer::brewer.pal(6, "Blues")[5:2],
            RColorBrewer::brewer.pal(6, "YlOrRd")[1:5])
  
col_func <- function(x){cols[x]}

startyr <- 2010

# Trick to avoid red frames also in legend:
# - put 'fill = Proref_ratio_cut' in aes of the first geom_tile (not in the aes of ggplot)
# - set 'alpha = 0' so there is no fill in the EQS tiles 

gg_tile <- ggplot(dat_medians, aes(MYEAR, Station)) +
  geom_tile(aes(fill = Proref_ratio_cut)) + 
  geom_tile(data = subset(dat_medians, Above_EQS %in% "Over"),
            color = "red", size = 1, height = 0.9, width = 0.9, alpha = 0) +
  geom_text(aes(label = round(Value, 3)), nudge_y = -0.1, size = 3) +
  geom_text(aes(label = LOQ_label), size = 3, nudge_y = 0.3) +
  scale_fill_manual(values = proref_cols) +
  scale_color_manual(values = c("red", "white")) +
  scale_alpha_manual(values = c(1, 0)) +
  scale_x_continuous(breaks = seq(startyr, 2020, 2), 
                     limits = c(startyr-0.5, current_year+0.5)) +
  theme_bw() +
  guides(colour = "none") +
  labs(
    title = paste0(params$param, " in ", species, " (", tissue, ")"),
    x = "Year", y = "")

gg_tile

```
### Plot, interactive

```{r, fig.width = 10, fig.height=11}

no_eqs <- mean(is.na(dat_medians$EQS_ratio)) == 1

if (no_eqs){
  dat_medians2 <- dat_medians %>% filter(MYEAR >= 2010) %>%
    mutate(
      Value_txt = paste0(
        "Proref ratio: ", round(Proref_ratio, 3), "<br>",
        "Median: ", LOQ_label, round(Value, 4), " ug/kg<br>",
        "(", round(Value_min, 4), "-", round(Value_max, 4), "; N =", N_median, ")")
    )
} else {
  dat_medians2 <- dat_medians %>% filter(MYEAR >= 2010) %>%
    mutate(
      Value_txt = paste0(
        "Proref ratio: ", round(Proref_ratio, 3), "<br>",
        "EQS ratio: ", round(EQS_ratio, 3), "<br>",
        "Median: ", LOQ_label, round(Value, 4), " ug/kg<br>",
        "(", round(Value_min, 4), "-", round(Value_max, 4), "; N =", N_median, ")")
    )
}

gg_tile <- ggplot(dat_plot2, aes(MYEAR, Station, label = round(Value, 3))) +
  geom_tile(aes(fill = Proref_ratio_cut)) + 
  geom_tile(data = subset(dat_medians2, Above_EQS %in% "Over"),
            color = "red", size = 1, height = 0.9, width = 0.9, alpha = 0) +
  scale_fill_manual("Proref\nratio", values = proref_cols) +
  scale_color_manual(values = c("red", "white")) +
  scale_alpha_manual(values = c(1, 0)) +
  geom_text(aes(label = LOQ_label), size = 3, nudge_y = 0.3) +
  geom_text_interactive(aes(tooltip = Value_txt), nudge_y = -0.1, size = 3) +
  scale_x_continuous(breaks = seq(startyr, 2020, 2), limits = c(startyr, current_year)) +
  theme_bw() +
  guides(colour = "none") +
  labs(
    title = paste0(params$param, " in ", species, " (", tissue, ")"),
    x = "Year", y = "")

girafe(ggobj = gg_tile, height_svg = 5.7, width_svg = 9)

```

## 5. Trend plots (long + short)   

### Get trends  

```{r}

# check <- readRDS("Data/126_df_trend_2021.rds")
# table(check$Trend_string)

dat_trends <- readRDS("Data/126_df_trend_2021.rds") %>%
  filter(PARAM %in% params$param)

xtabs(~Trend_type, dat_trends)                 # should be equally many long and short 

xtabs(~Trend_string + Trend_type, dat_trends)

# xtabs(~TISSUE_NAME + Trend_string, dat_trends)
# xtabs(~STATION_CODE + TISSUE_NAME, dat_trends %>% filter(TISSUE_NAME %in% "Whole soft body"))

if (FALSE){
  # For testing
  dat_series_trendrun <- readRDS("Data/125_dat_series.rds") %>%
    filter(PARAM %in% params$param,
           TISSUE_NAME %in% tissue)
  dat_medians %>%
    distinct(STATION_CODE) %>%
    left_join(
      dat_trends %>% filter(Trend_type %in% "long") %>%
        count(STATION_CODE)
    ) %>%
    left_join(dat_series_trendrun) %>% 
    View("test")
  # should have data in all rows for column 'n' 
}


```


### Data for trends plot  

* The difference between 'No change' and 'No trend' is very subtle (see script 126):   
   - 'No change' means that there is a time trend, but no difference between the two actual years compared  
   - 'No trend' means that there is no time trend (the best model was y = constant)   
* So we combine the two  

```{r}

shape_order <- c("Increasing", "Decreasing", "No change")
colour_order <- shape_order

dat_trends_plot <- dat_trends %>%
  filter(LATIN_NAME %in% species,
         TISSUE_NAME %in% tissue) %>%
  # Add station name
  left_join(
    dat_list$lookup_stations %>% 
      select(STATION_CODE, Station_name),
    by = "STATION_CODE"
  ) %>%
  mutate(
    Station = factor(
      paste(STATION_CODE, Station_name),
      levels = rev(station_order$Station)),
    # Shape and colour: Drop "Too few over-LOQ years" and "Too few years:
    Trend_shape = case_when(
      Trend_string %in% c("Increasing" , "Decreasing", "No change") ~ Trend_string, 
      TRUE ~ as.character(NA)),
    Trend_shape = factor(Trend_shape, levels = shape_order),
    Trend_color = Trend_shape,
    # Text: Keep only "Too few over-LOQ years", "Too few years" and "No data before 2012":
    Trend_text = case_when(
      Trend_string %in% c("Increasing" , "Decreasing", "No change") ~ as.character(NA), 
      TRUE ~ Trend_string),
    Perc_annual = case_when(
      Trend_string %in% c("No change", "No trend") ~ 0, 
      TRUE ~ Perc_annual),
    Trend_shape = factor(Trend_shape, levels = shape_order)
  )

table(dat_trends_plot$Trend_type)   # Should be equal
xtabs(~addNA(Trend_shape) + Trend_type, dat_trends_plot)   # Should be no NAs
table(addNA(dat_trends_plot$Trend_string), dat_trends_plot$Trend_type)

# First table (this WITH Trend_shape): should have no 'Trend_text'
# Second table (this WITHOUT Trend_shape): should all have a 'Trend_text'
xtabs(~addNA(Trend_text) + Trend_type + is.na(Trend_shape), dat_trends_plot)   


```

### Test plot   

```{r}

y_ran <- range(dat_trends_plot$Perc_annual, na.rm = TRUE)

ggplot(dat_trends_plot %>% filter(Trend_type %in% "long"),
       aes(x = Station, y = Perc_annual)) +
  geom_linerange(aes(ymin = Perc_annual_lo, ymax = Perc_annual_hi, colour = Trend_color)) +
  geom_point(aes(colour = Trend_color, shape = Trend_color)) +
  geom_text(aes(y = 0, label = Trend_text)) +
  coord_flip(ylim = y_ran)

```


### Test plot 2  

```{r}

# Tjukkere linjer, større punkter  
# Linjer som mangler skal ha mellomrom, så de stemmer med den andre grafen
# Aksetekst "Annual change, %"

y_ran <- range(dat_trends_plot$Perc_annual, na.rm = TRUE)

trend_shapes <- c(24, 25, 21) %>% setNames(shape_order)
trend_colours <- c("red2", "green4", "grey20") %>% setNames(colour_order)
trend_fill <- c("red2", "green4", "grey20") %>% setNames(colour_order)

ggplot(dat_trends_plot %>% filter(Trend_type %in% "long"),
       aes(x = Station, y = Perc_annual)) +
  geom_linerange(aes(ymin = Perc_annual_lo, ymax = Perc_annual_hi, colour = Trend_color)) +
  geom_point(aes(colour = Trend_color, fill = Trend_color, shape = Trend_color)) +
  geom_text(aes(y = 0, label = Trend_text)) +
  scale_colour_manual(values = trend_colours, drop = FALSE) +
  scale_fill_manual(values = trend_colours, drop = FALSE) +
  scale_shape_manual(values = trend_shapes, drop = FALSE) +
  geom_hline(yintercept = 0) +
  coord_flip(ylim = y_ran) +
  theme_bw() 


ggplot(dat_trends_plot %>% filter(Trend_type %in% "short"),
       aes(x = Station, y = Perc_annual)) +
  geom_linerange(aes(ymin = Perc_annual_lo, ymax = Perc_annual_hi, colour = Trend_color)) +
  geom_point(aes(colour = Trend_color, fill = Trend_color, shape = Trend_color)) +
  geom_text(aes(y = 0, label = Trend_text)) +
  scale_colour_manual(values = trend_colours, drop = FALSE) +
  scale_fill_manual(values = trend_colours, drop = FALSE) +
  scale_shape_manual(values = trend_shapes, drop = FALSE) +
  geom_hline(yintercept = 0) +
  coord_flip(ylim = y_ran) +
  theme_bw() 

```

### Make both trend plots   

```{r}

gg_trends <- c("long", "short") %>%
  map(
    ~ggplot(dat_trends_plot %>% filter(Trend_type %in% .x),
       aes(x = Station, y = Perc_annual)) +
  geom_linerange(aes(ymin = Perc_annual_lo, ymax = Perc_annual_hi, colour = Trend_color)) +
  geom_point(aes(colour = Trend_color, fill = Trend_color, shape = Trend_color)) +
  geom_text(aes(y = 0, label = Trend_text)) +
  scale_colour_manual(values = trend_colours, drop = FALSE) +
  scale_fill_manual(values = trend_colours, drop = FALSE) +
  scale_shape_manual(values = trend_shapes, drop = FALSE) +
  geom_hline(yintercept = 0) +
  coord_flip(ylim = y_ran) +
    labs(title = .x, y = "") +
  theme_bw() 
  )

gg_trends[[1]]
gg_trends[[2]]

```

### Test interactive plot  

```{r}

trend_shapes <- c(24, 25, 21) %>% setNames(shape_order)
trend_colours <- c("red2", "green4", "grey20") %>% setNames(colour_order)
trend_fill <- c("red2", "green4", "grey20") %>% setNames(colour_order)

dat_trends_plot <- dat_trends_plot %>%
  mutate(
    Tooltip_text = case_when(
      Trend_string %in% c("Increasing", "Decreasing") ~ 
        paste0(
          "Trend (", Trend_type, "-term) :" , round(Perc_annual,2), " % (", First_year, "-", current_year, ")\n",
          "Change since ", First_year, ": ", Perc_change, " % (", p_category, ")\n", 
          "Trend model: ", ifelse(k_sel >= 3, paste("Non-linear (k = ", k_sel, ")"), "Linear")
        ),
      TRUE ~ Trend_string
    )
  )

trendtype <- "long"

gg_range <- ggplot(dat_trends_plot %>% filter(Trend_type %in% trendtype),
       aes(x = Station, y = Perc_annual)) +
  geom_text(aes(y = 0, label = Trend_text)) +                                                # For "too few years etc.
  geom_linerange(aes(ymin = Perc_annual_lo, ymax = Perc_annual_hi, colour = Trend_color)) +
  geom_hline(yintercept = 0) +
  geom_point_interactive(aes(colour = Trend_color, fill = Trend_color, shape = Trend_color, tooltip = Tooltip_text)) +
  scale_colour_manual(values = trend_colours, drop = FALSE) +
  scale_fill_manual(values = trend_colours, drop = FALSE) +
  scale_shape_manual(values = trend_shapes, drop = FALSE) +
  coord_flip(ylim = y_ran) +
  theme_bw() +
  labs(
    title = stringr::str_to_sentence(paste0(trendtype, "-term trend"))
  )

girafe(ggobj = gg_range, height_svg = 5.7, width_svg = 9)

```


## 6. Combine tile and trend plots    

```{r, fig.width=10, fig.height=8}

ggcomb <- cowplot::plot_grid(gg_tile + theme(legend.position = "none"), 
                   gg_trends[[1]] + 
                     theme(axis.title.y = element_blank(),
                           axis.text.y = element_blank(),
                           legend.position = "none"), 
                   gg_trends[[2]] + 
                     theme(axis.title.y = element_blank(),
                           axis.text.y = element_blank()), 
                   cowplot::get_legend(gg_tile),
                   nrow = 2, ncol = 3,
                   rel_heights = c(3,1),
                   rel_widths = c(2.8, 0.6, 1)
                   )

ggsave("Figures/430_HG_combo.png", ggcomb,
       width = 14, height = 8, dpi = 200)


```


## 7. Ratio plots - data    
```{r}

region_order <- c("Viken og Oslo", "Vestfold og Telemark", "Agder", "Vestland", "Møre og Romsdal", 
                  "Trøndelag", "Nordland", "Troms og Finnmark", "Svalbard")

dat_medians_lastyear <- dat_medians %>% 
  ungroup() %>%
  filter(MYEAR == current_year) %>%
  mutate(
    Proref_ratio_min = Value_min/Proref,
    Proref_ratio_max = Value_max/Proref,
    EQS_ratio_min = Value_min/EQS,
    EQS_ratio_max = Value_max/EQS,
    Proref_tooltip = paste0(
      "Conc.: ", Value, " (Min-max: ", Value_min, "-", Value_max, ")\n",
      "Ratio: ", Proref_ratio, " (Min-max: ", Proref_ratio_min, "-", Proref_ratio_max, ")\n"),
    EQS_tooltip = paste0(
      "Conc.: ", Value, " (Min-max: ", Value_min, "-", Value_max, ")"),
    Station = factor(Station, levels = rev(levels(Station))),
    Region = factor(Region, levels = rev(region_order))
  ) 


```


## 8. Proref ratio  

### Static version 1  
```{r}

# For fisk: median som fylt punkt med interkvartilt range som strek, med enkeltprøver som mindre svart prikk   
# For blåskjell: median som fylt punkt med range som strek, med punmed punkter som mindre svart prikkkter som mindre svart prikk    
# Ingen whisker  

gg <- ratioplot(data = dat_medians_lastyear,
          x = "Station", y = "Proref_ratio", fill = "Region", 
          tooltip = "Tooltip_text", data_id = "STATION_CODE",
          ymin = "Proref_ratio_min", ymax = "Proref_ratio_max") + 
  labs(y = "Ratio concentration/Proref", x = "")

ggsave("Figures/430_HG_Prorefratio_wout_log.png", gg,
       width = 8, height = 4.5, dpi = 200)

ggsave("Figures/430_HG_Prorefratio_with_log.png", gg + scale_y_log10(),
       width = 8, height = 4.5, dpi = 200)

gg
gg + scale_y_log10()

```

### Static version 2  
```{r}

dat_raw_lastyear_01 <- dat_raw %>%
  filter(LATIN_NAME %in% species,
         TISSUE_NAME %in% tissue,
         Basis %in% "WW",
         MYEAR %in% current_year)

# Add station name
dat_raw_lastyear_02 <- dat_raw_lastyear_01 %>%
  left_join(
    dat_list$lookup_stations %>% 
      filter(LATIN_NAME %in% species) %>%
      select(STATION_CODE, LATIN_NAME, Station_name, Region),
    by = "STATION_CODE"
  ) %>%
  mutate(
    Station = factor(
      paste(STATION_CODE, Station_name),
      levels = rev(station_order$Station))
  )

# Add Proref + EQS
df_limits <- dat_medians %>%
  ungroup() %>%
  distinct(PARAM, LATIN_NAME, TISSUE_NAME, Basis, EQS, Proref)
if (nrow(df_limits) != 1)
  stop("We need one and exacly one row here")

dat_raw_lastyear <- dat_raw_lastyear_02 %>%
  left_join(df_limits) %>%
  mutate(
    Proref_ratio = Value/Proref,
    EQS_ratio = Value/EQS
  )

# Test    
# ggplot(dat_raw_lastyear, aes(Station, Proref_ratio)) +
#   geom_point(shape = 19, size = 1)

gg <- ratioplot(data = dat_medians_lastyear,
          x = "Station", y = "Proref_ratio", fill = "Region", 
          tooltip = "Tooltip_text", data_id = "STATION_CODE",
          ymin = "Proref_ratio_min", ymax = "Proref_ratio_max",
          extra_points = dat_raw_lastyear) + 
  labs(y = "Ratio concentration/Proref", x = "")

gg

```

### Dynamic  

```{r}

ratioplot_interactive(data = dat_medians_lastyear,
          x = "Station", y = "Proref_ratio", fill = "Region", 
          tooltip = "Proref_tooltip", data_id = "STATION_CODE",
          ymin = "Proref_ratio_min", ymax = "Proref_ratio_max")  

```

## 9. EQS ratio 

### Static  
```{r}

gg <- ratioplot(data = dat_medians_lastyear,
          x = "Station", y = "EQS_ratio", fill = "Region", 
          tooltip = "EQS_tooltip", data_id = "STATION_CODE",
          ymin = "EQS_ratio_min", ymax = "EQS_ratio_max") + 
  labs(y = "Ratio concentration/EQS", x = "")

ggsave("Figures/430_HG_EQSratio_wout_log.png", gg,
       width = 8, height = 4.5, dpi = 200)

ggsave("Figures/430_HG_EQSratio_with_log.png", gg + scale_y_log10(),
       width = 8, height = 4.5, dpi = 200) 

gg
gg + scale_y_log10()

```

### Interactive 
```{r}


ratioplot_interactive(data = dat_eqsplot,
                      x = "Station", y = "EQS_ratio", fill = "Region", 
                      tooltip = "Tooltip_text", data_id = "STATION_CODE",
                      ymin = "EQS_ratio_min", ymax = "EQS_ratio_max")  


```




## 10. All time series, interactive  

```{r}

library(plotly)

dat <- dat_list$data %>%
  filter(LATIN_NAME %in% species,
         TISSUE_NAME %in% tissue)

p <- ggplot(data = dat_plot, aes(x = MYEAR, y = Value, color = Station)) +
  geom_line() +
  geom_point() +
  scale_y_log10()

ggplotly(p)

```
