---
title: "433_Report_LOQ_plots"
author: "DHJ"
date: "2022-10-15"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

## Startup  
```{r 01_settings, include=FALSE}

current_year <- 2022
knitr::opts_chunk$set(echo = FALSE, results = 'hold')

```

```{r 02_packages, include=FALSE}

library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(flextable)
# library(glue)
library(readxl)
library(purrr)
library(glue)
# library(leftcensored)   # DHJ's package (https://github.com/DagHjermann/leftcensored)
library(scico)            # colour palettes incl. "Vik" (https://github.com/thomasp85/scico)

library(ggiraph)
library(cowplot)

# library(safejoin) # https://github.com/moodymudskipper/safejoin

```

## Data    

```{r 03_medians_data1}

dat_raw <- readRDS("Data/109_adjusted_data_ELU2_2023-09-12.rds") %>%
  filter(!STATION_CODE %in% c("I965","I969"))

dat_medians <- readRDS("Data/110_mediandata_updated_ELU_2023-09-12.rds") %>%
  filter(!STATION_CODE %in% c("I965","I969"))

# Stations - for corrrect ordering
lookup_stations <- read.csv("Input_data/Lookup_tables/Lookup_stationorder.csv") # %>%


```


## Median LOQ by year
```{r}

# 2021:
# params <- c("AG", "AS", "CD", "CO", "CR", "CU", "HG", "NI", "PB", "ZN", 
#   "BDE47", "BDE99", "BDE100", "BDE153", "HBCDA", "HCB", "CB118", "CB138",  "CB153", 
#   "PFOA", "PFOS", "PFOSA", "BAA", "BAP", "FLU", "PYR", "D5")

df_params <- read_excel("Input_data/Lookup_tables/Lookup_priority_for_timetrends.xlsx")
params <- df_params %>%
  filter(`Selected 2022[PARAM]` == 1) %>%
  pull(PARAM)

sel <- params %in% "SCCP eksl. LOQ"
params[sel] <- "SCCP"
sel <- params %in% "MCCP eksl. LOQ"
params[sel] <- "MCCP"

params1 <- params[!params %in% c("SCCP","MCCP")]
params2 <- params[params %in% c("SCCP","MCCP")]

```

### LOQ for all parameters except SCCP, MCCP
```{r}

# LOQ median values for all parameters except SCCP, MCCP
dat_loq1 <-  dat_raw %>%
  filter(PARAM %in% params1) %>%
  group_by(PARAM, MYEAR, UNIT) %>%
  summarise(
    medLOQ = median(VALUE_WW[!is.na(FLAG1)], na.rm = TRUE) %>% round(4),
    .groups = "drop"
  ) 

```

## LOQ for SCCP and MCCP  
```{r}

# SCCP
df2a <- dat_raw %>%
  filter(PARAM %in% c("SCCP inkl. LOQ", "SCCP eksl. LOQ")) %>%
  select(MYEAR, STATION_CODE, TISSUE_NAME, SAMPLE_NO2, PARAM, VALUE_WW, UNIT) %>%
  pivot_wider(names_from = PARAM, values_from = VALUE_WW) 
df2b <- df2a %>%
  mutate(PARAM = "SCCP") %>%
  group_by(PARAM, MYEAR, UNIT) %>%
  summarise(
    medLOQ = median(`SCCP inkl. LOQ`[`SCCP eksl. LOQ` == 0], na.rm = TRUE) %>% round(4),
    .groups = "drop"
  ) 

# MCCP
df2c <- dat_raw %>%
  filter(PARAM %in% c("MCCP inkl. LOQ", "MCCP eksl. LOQ")) %>%
  select(MYEAR, STATION_CODE, TISSUE_NAME, SAMPLE_NO2, PARAM, VALUE_WW, UNIT) %>%
  pivot_wider(names_from = PARAM, values_from = VALUE_WW) 
df2d <- df2c %>%
  filter(`MCCP eksl. LOQ` == 0) %>%
  mutate(PARAM = "MCCP") %>%
  group_by(PARAM, MYEAR, UNIT) %>%
  summarise(
    medLOQ = median(`MCCP inkl. LOQ`[`MCCP eksl. LOQ` == 0], na.rm = TRUE) %>% round(4),
    .groups = "drop"
  ) 

# check SCCP 
df2a %>%
  mutate(ratio_excluding_including = `SCCP eksl. LOQ`/`SCCP inkl. LOQ`) %>%
  ggplot() + 
  geom_histogram(aes(x = ratio_excluding_including))

dat_loq2 <- bind_rows(df2b, df2d)

```

### Combine 
```{r}

dat_loq <- bind_rows(dat_loq1, dat_loq2) %>%
  mutate(PARAM = factor(PARAM, levels = params))

```


### Tables  
```{r}

dat_loq2a <- dat_loq %>% 
  arrange(PARAM, MYEAR) %>%
  filter(MYEAR >= 2001 & MYEAR <= 2011) %>%
  tidyr::pivot_wider(names_from = MYEAR, values_from = medLOQ, names_sort = TRUE)

dat_loq2b <- dat_loq %>% 
  arrange(PARAM, MYEAR) %>%
  filter(MYEAR >= 2012) %>%
  tidyr::pivot_wider(names_from = MYEAR, values_from = medLOQ, names_sort = TRUE)

write.csv(dat_loq2a, "Data/433_LOQ_medians1.csv", row.names = FALSE, na = "")
write.csv(dat_loq2b, "Data/433_LOQ_medians2.csv", row.names = FALSE, na = "")

dat_loq2a
dat_loq2b

```


### Plot  
```{r, fig.width=12, fig.height=8}

dat_loq_ref <- dat_loq %>%
  filter(MYEAR %in% 2013:2015) %>%
  group_by(PARAM) %>%
  summarise(medLOQ_ref = median(medLOQ, na.rm = TRUE))  

dat_loq2 <- dat_loq %>% 
  filter(MYEAR >= 2002) %>%
  left_join(dat_loq_ref, by = "PARAM") %>%
  mutate(
    loq_rel = log10(medLOQ/medLOQ_ref),
    PARAM = forcats::fct_rev(PARAM)) %>%
  group_by(PARAM) %>%
  mutate(
    min_medLOQ = min(medLOQ, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    medLOQ_label = case_when(
      is.na(medLOQ) ~ "-",
      min_medLOQ < 0.001 ~ sprintf("%.4f", medLOQ),
      min_medLOQ < 0.01 ~ sprintf("%.3f", medLOQ),
      min_medLOQ < 0.1 ~ sprintf("%.2f", medLOQ),
      TRUE ~ sprintf("%.1f", medLOQ)
    )
  )

# range(subset(dat_loq2, MYEAR >= 2012)$loq_rel, na.rm = TRUE)
#  range(subset(dat_loq2, MYEAR >= 2002)$loq_rel, na.rm = TRUE)

sel <- is.na(dat_loq2$loq_rel)
dat_loq2$loq_rel[sel] <- 0

white <- c(-0.5,0.5)

gg <- ggplot(dat_loq2, aes(MYEAR, PARAM, fill = loq_rel)) +
  geom_tile() +
  scale_fill_scico("Relative LOQ", palette = "vik", direction = 1,  midpoint = 0) +
  geom_text(data = subset(dat_loq2, is.na(loq_rel)), 
            aes(label = medLOQ_label), color = "black") +
  geom_text(data = subset(dat_loq2, loq_rel >= white[1] & loq_rel <= white[2]), 
            aes(label = medLOQ_label), color = "black", size = 3.5) +
  geom_text(data = subset(dat_loq2, loq_rel < white[1] | loq_rel > white[2]), 
            aes(label = medLOQ_label), color = "white", size = 3.5) +
  scale_x_continuous(breaks = seq(2002,2022,1)) +
  theme(axis.title = element_blank())
  
gg1 <- gg + coord_cartesian(xlim = c(2002,2012))
gg2 <- gg + coord_cartesian(xlim = c(2013,2022))

ggsave("Figures/433_LOQ-median1.png", gg1, width = 10, height = 6.5, dpi = 200)
ggsave("Figures/433_LOQ-median2.png", gg2, width = 10, height = 6.5, dpi = 200)

gg1
gg2

```

### Add a 2021 version for last plot   

```{r}

gg2b <- gg + coord_cartesian(xlim = c(2013,2021))
ggsave("Figures/433_LOQ-median2_2021.png", gg2b, width = 10*9/10, height = 6.5, dpi = 200)

```


## Detection frequency


### Detection frequency 1 (by param * station, one parameter in all years)  

* Vik color scale used  
* This plot shows the median value for each station/year, it could instead show limit of quantification (median value of less-thans)  

```{r, fig.width=13, fig.height=7}

dat_detectplot1 <- dat_medians %>% 
  left_join(lookup_stations %>% select(Order, STATION_CODE, Station_name)) %>%
  arrange(Order, PARAM) %>%
  filter(Basis %in% "WW" & MYEAR %in% current_year & PARAM %in% params) %>%
  mutate(
    Parameter = factor(PARAM, levels = params),
    Parameter = forcats::fct_rev(Parameter),
    tissue_bird = ifelse(STATION_CODE == "19N", TISSUE_NAME, ""),
    Station = forcats::fct_inorder(paste(STATION_CODE, Station_name, tissue_bird)),
    Detected = round(100*(Over_LOQ/N_median), 0)
    )

# dat_detectplot1 %>% filter(is.na(Station)) %>% View()

ggplot(dat_detectplot1, aes(Station, Parameter, fill = Detected)) +
  geom_tile() +
  geom_text(data = subset(dat_detectplot1, Detected > 35 & Detected < 65), 
            aes(label = Detected), size = 3) +
  geom_text(data = subset(dat_detectplot1, Detected <= 35 | Detected >= 65), 
            aes(label = Detected), size = 3, 
            color = "white") +
  # scale_fill_viridis_b(breaks = seq(0,1,0.1)) +
  scale_fill_scico(palette = "vik", direction = -1,  midpoint = 50, limits = c(0,100)) +
  ggeasy::easy_rotate_labels("x", angle = -45) +
  theme(axis.title.x = element_blank())

```

### Other way   

```{r, fig.width=13, fig.height=11}

dat_detectplot1b <- dat_detectplot1 %>%
  mutate(Station = forcats::fct_rev(Station),
         Parameter = forcats::fct_rev(Parameter)
  )

gg <- ggplot(dat_detectplot1b, aes(Parameter, Station, fill = Detected)) +
  geom_tile() +
  geom_text(data = subset(dat_detectplot1, Detected > 35 & Detected < 65), 
            aes(label = Detected), size = 3) +
  geom_text(data = subset(dat_detectplot1, Detected <= 35 | Detected >= 65), 
            aes(label = Detected), size = 3, 
            color = "white") +
  # scale_fill_viridis_b(breaks = seq(0,1,0.1)) +
  scale_fill_scico(palette = "vik", direction = -1,  midpoint = 50, limits = c(0,100)) +
  ggeasy::easy_rotate_labels("x", angle = -45) +
  theme(axis.title = element_blank())

ggsave(paste0("Figures/433_LOQ-median-", current_year, ".png"), gg, 
       width = 10, height = 10, dpi = 200)

gg

```


### Detection frequency 2 (by year * station, one parameter in all years)  

* Vik color scale used  
* This plot shows the median value for each station/year, it could instead show limit of quantification (median value of less-thans)  

```{r, fig.width=11, fig.height=7}

param <- "CR"

dat_detectplot2 <- dat_medians %>% 
  filter(PARAM %in% param & MYEAR >= 2012) %>%
  group_by(STATION_CODE) %>%
  mutate(max_year = max(MYEAR)) %>% 
  ungroup() %>%
  filter(max_year %in% current_year) %>%
  left_join(lookup_stations %>% select(Order, STATION_CODE, Station_name)) %>%
  arrange(Order) %>%
  mutate(
    tissue_bird = ifelse(STATION_CODE == "19N", TISSUE_NAME, ""),
    Station = forcats::fct_inorder(paste(STATION_CODE, Station_name, tissue_bird)),
    Detected = round(100*(Over_LOQ/N_median), 0)
    ) %>%
  group_by(Station, MYEAR) %>%
  summarise(Detected = median(Detected))

white <- c(35, 65)

ggplot(dat_detectplot2, aes(MYEAR, Station, fill = Detected)) +
  geom_tile() +
  geom_text(data = subset(dat_detectplot2, Detected <= white[1] | Detected >= white[2]),
            aes(label = Detected), size = 3, color = "white") +
  geom_text(data = subset(dat_detectplot2, Detected > white[1] & Detected < white[2]),
            aes(label = Detected), size = 3, color = "black") +
  # scale_fill_viridis_b(breaks = seq(0,1,0.1)) +
  scale_fill_scico(palette = "vik", direction = -1,  midpoint = 50, limits = c(0,100)) +
  labs(
    title = param
  )

```
