---
title: "471_SCCP_MCCP_2021"
author: "DHJ"
date: "2022-10-30"
output: html_document
---

 we assume that uncertainty is the same for all years (sounds worse than it is?)

## 0. Settings  
```{r}

selected_year <- 2021

```


## 1. Packages  
```{r}

library(dplyr)
library(tidyr)
library(purrr)
library(readxl)
library(ggplot2)
source("002_Utility_functions.R")

```


## 2. Get data  

### a. Main data  
Read and reformat the most recent data (by default)     
```{r}

# Code from script 105

# Files 
files <- list_files("Data", pattern = "101_data_updated")

# Pick file
filename <- files[1]
cat("\nLast file:", filename, "\n")
cat("Time since this file was modified: \n")
Sys.time() - file.info(paste0("Data/", filename))$mtime 

# Info
cat("\nIf you want to read a different file, input a different name for 'filename' \n")

# Read data
dat_all <- readRDS(paste0("Data/", filename))

# We save the date part of the text (e.g., '2020-04-23')
# This will be used in part 10, when we save the resulting file
file_date <- substr(filename, 18, 27)    # pick out the part of the text from character no. 17 to no. 26

```

## 3. Tables for SCCP    

### PARAM used by year    
```{r}

dat_all %>% filter(grepl("SCCP", PARAM)) %>% xtabs(~MYEAR + PARAM, .)

```

### PARAM used by year    
```{r}

dat_all %>% filter(grepl("SCCP", PARAM)) %>% xtabs(~MYEAR + PARAM, .)

```

### Number of samples by PARAM + station, last years   
```{r}

dat_all %>% filter(grepl("SCCP", PARAM) & MYEAR == 2021) %>% xtabs(~STATION_CODE + PARAM, .)
dat_all %>% filter(grepl("SCCP", PARAM) & MYEAR == 2020) %>% xtabs(~STATION_CODE + PARAM, .)

```
### Less-than values, 2021      

```{r}

dat_all %>% filter(grepl("SCCP", PARAM) & MYEAR %in% 2021 & FLAG1 %in% "<") %>% xtabs(~STATION_CODE + MYEAR + PARAM, .)

```

### Less-than values, 2019-2020    

```{r}

dat_all %>% filter(grepl("SCCP", PARAM) & FLAG1 %in% "<") %>% xtabs(~MYEAR + PARAM, .)

dat_all %>% filter(grepl("SCCP", PARAM) & MYEAR %in% 2019:2020 & FLAG1 %in% "<") %>% xtabs(~STATION_CODE + MYEAR + PARAM, .)

```

### Check a station, 2021  

```{r}

stn <- "98B1"  
yr <- 2021  

dat_all %>% 
  filter(grepl("SCCP", PARAM) & MYEAR %in% yr & STATION_CODE == stn) %>%
  select(STATION_CODE, SAMPLE_NO2, PARAM, VALUE=VALUE_WW, FLAG1) %>%
  pivot_wider(id_cols = c(STATION_CODE, SAMPLE_NO2), names_from = c(PARAM), values_from = c(VALUE, FLAG1)) %>%
  arrange(STATION_CODE, SAMPLE_NO2)

```


### Incl. and excl. LOQ  
```{r, fig.width=8.5, fig.height=6}

# spec <- "Gadus morhua"
param <- "SCCP"
params <- paste(param, c("eksl. LOQ", "inkl. LOQ"))

for (spec in c("Gadus morhua", "Mytilus edulis")){
  
  gg <- dat_all %>% 
    filter(PARAM %in% params,
           LATIN_NAME %in% spec,
           MYEAR >= 2015) %>%
    group_by(STATION_CODE, MYEAR, PARAM) %>%
    summarise(Median = median(VALUE_WW)) %>%
    ggplot(aes(MYEAR, Median, colour = PARAM)) +
    scale_colour_manual(values = c("blue4", "red2")) +
    geom_line(shape = 1) +
    facet_wrap(vars(STATION_CODE)) +
    labs(title = paste(param, "(with/without LOQ)")) +
    theme_bw()
  
  print(gg)
  
}

```


### Incl. and excl. LOQ  
```{r, fig.width=10, fig.height=8}

spec <- "Gadus morhua"

dat_all %>% 
  filter(PARAM %in% c("SCCP eksl. LOQ", "SCCP inkl. LOQ"),
         LATIN_NAME %in% spec,
         MYEAR >= 2015) %>%
  select(STATION_CODE, SAMPLE_NO2, MYEAR, PARAM, VALUE_WW) %>%
  ggplot(aes(MYEAR, VALUE_WW, colour = PARAM)) +
  geom_point(shape = 1) +
  facet_wrap(vars(STATION_CODE))

```


## Tables for 2021, for report    

### SCCP eksl. LOQ   

```{r}

spec <- "Gadus morhua"

dat_all %>% 
  filter(PARAM %in% "SCCP eksl. LOQ") & MYEAR %in% selected_year & LATIN_NAME %in% spec) %>%
  group_by(STATION_CODE) %>%
  summarise(
    n = n(), 
    min = min(VALUE_WW),
    max = max(VALUE_WW)
    )
  pivot_wider(id_cols = c(STATION_CODE, SAMPLE_NO2), names_from = c(PARAM), values_from = c(VALUE, FLAG1)) %>%
  arrange(STATION_CODE, SAMPLE_NO2) 

```


## 

### a. Main data  
Read and reformat the most recent data (by default)     
```{r}

# Code from script 105

# Files 
files <- list_files("Data", pattern = "109_adjusted")

# Pick file
filename <- files[1]
cat("\nLast file:", filename, "\n")
cat("Time since this file was modified: \n")
Sys.time() - file.info(paste0("Data/", filename))$mtime 

# Info
cat("\nIf you want to read a different file, input a different name for 'filename' \n")

# Read data
dat_all <- readRDS(paste0("Data/", filename))

# We save the date part of the text (e.g., '2020-04-23')
# This will be used in part 10, when we save the resulting file
file_date <- substr(filename, 18, 27)    # pick out the part of the text from character no. 17 to no. 26

```


