---
title: "502_Biological_effect_data_export"
output: html_document
---


## 1. Packages
```{r}

library(dplyr)
library(purrr)
library(tidyr)
library(stringr)
library(lubridate)
library(ggplot2)

```


## 2. Data

### Main data from script 101    
Read and reformat the most recent data (by default)  
```{r}

files <- dir("Data", pattern = "101_data_updated") %>% rev()

cat("Reading the last file downloaded:")
cat("\n", files[1])
cat("\n")
cat("If you want to read a different file, replace 'files[1]' with the file you want")
cat("\n")

filename <- files[1] 
data_all <- readRDS(paste0("Data/", filename))

# We save the date part of the text (e.g., '2020-04-23')
# This will be used in part 10, when we save the resulting file
file_date <- substr(filename, 18, 27)    # pick out the part of the text from character no. 17 to no. 26

```

## 3. Get stations  

### Set 227G1 and 227G2 to '227G'   
```{r}

# Tables of data in stations 227G, 227G1, 227G2
if (FALSE){
  
  df1 <- data_all %>% filter(grepl("227G", STATION_CODE) & PARAM == "TBT")
  xtabs(~MYEAR + STATION_CODE, df1)
  
  df2 <- data_all %>% filter(grepl("227G", STATION_CODE) & PARAM == "VDSI")
  xtabs(~MYEAR + STATION_CODE, df2)
  
}

# Change 227G1 and 227G2 to '227G'
data_all2 <- data_all %>%
  mutate(STATION_CODE = case_when(
    grepl("227G", STATION_CODE) ~ "227G",
    TRUE ~ STATION_CODE)
  )
cat("227G1 and 227G2 changed to '227G' \n")

# also, remove one duplicate
sel <- with(data_all2, STATION_CODE %in% "227G" & MYEAR == 2015 & PARAM == "VDSI") %>% which()
# data_all2[sel,]
data_all3 <- data_all2[-sel[2],]
cat("One duplicate removed \n")

# Check for duplicates  
df_duplicates <- data_all3 %>%
  filter(LATIN_NAME %in% c("Littorina littorea", "Nucella lapillus")) %>%
  group_by(STATION_CODE, LATIN_NAME, TISSUE_NAME, MYEAR, SAMPLE_NO2, PARAM) %>%
  mutate(n = n()) %>%
  filter(n > 1)

if (nrow(df_duplicates) > 0){
  cat("WARNING! Duplicates in the data! Check 'df_duplicates'. \n")
  xtabs(~PARAM, df_duplicates) %>% print()
  xtabs(~MYEAR, df_duplicates) %>% print()
  xtabs(~STATION_CODE, df_duplicates) %>% print()
} else {
  cat("No more duplicates found \n")
}

```

## Extract data  
```{r}

# Get stations
biol_effect_stations <- data_all3 %>%
  filter(PARAM %in% c("PYR1O", "PYR1OH", "VDSI", "Intersex") & MYEAR >= 2010) %>%
  count(STATION_CODE) %>%
  pull(STATION_CODE)

# Get data
data_biol_effects <- data_all3 %>%
  filter(MYEAR >= 2010 & STATION_CODE %in% biol_effect_stations) %>%
  # Drop 1) TBTIN (tim atom weight) and 2) individual-based snail measurements (2010-2013 only)
  filter(!PARAM %in% c("TBTIN", "Tributyltinn (TBT)-Sn",
                       "HTMEA", "IMPS", "LNFPE", "LNMPE")
         ) %>%
  filter(TISSUE_NAME != "WO")

cat(nrow(data_biol_effects), "records extracted")

```

## Check extraction  
```{r}

data_biol_effects %>%
  xtabs(~TISSUE_NAME + LATIN_NAME, .)

```


### Check cod extraction  
- The "O" parameters are the "OH" values normalized by AY380:   
    - PA1O = PA1OH/AY380   
    - PYR1O = PYR1OH/AY380   
    - BAP3O = BAP3OH/AY380  
```{r}

# Show parameters as data
if (FALSE){
  data_biol_effects %>%
    filter(LATIN_NAME == "Gadus morhua") %>%
    distinct(PARAM, MYEAR) %>%
    group_by(PARAM) %>%
    summarise(No_of_years = n()) %>%
    View()
}

#
# Print out parameters, grouped by number of years
#

df_no_years_per_parameter <- data_biol_effects %>%
  filter(LATIN_NAME == "Gadus morhua") %>%
  distinct(PARAM, MYEAR) %>%
  count(PARAM)

for (i in sort(unique(df_no_years_per_parameter$n))){
  cat(i, "years: \n")
  pars <- df_no_years_per_parameter %>%
    filter(n == i) %>%
    pull(PARAM)
  cat(pars, sep = ", ")
  cat("\n\n")
}


```

### Check snail extraction   
```{r}

data_biol_effects %>%
  filter(LATIN_NAME %in% c("Littorina littorea", "Nucella lapillus")) %>%
  xtabs(~PARAM + MYEAR, .) 

```

### Checking out the relationship between some of the tins   
- "MBTIN" equals 1.5 times "Monobutyltinn (MBT)-Sn"
- "DBTIN" equals 2 times "Dibutyltinn-Sn (DBT-Sn)"  
```{r}

df <- data_biol_effects %>%
  filter(PARAM %in% c("Monobutyltinn (MBT)-Sn", "MBTIN") & MYEAR == 2017) %>%
  arrange(STATION_CODE, SAMPLE_NO2, PARAM) %>%
  select(STATION_CODE, SAMPLE_NO2, PARAM, VALUE_WW) %>%
  pivot_wider(names_from = PARAM, values_from = VALUE_WW)
df$Ratio <- df[[3]]/df[[4]]
df  # Ratio = 1.5

df <- data_biol_effects %>%
  filter(PARAM %in% c("Dibutyltinn-Sn (DBT-Sn)", "DBTIN") & MYEAR == 2017) %>%
  arrange(STATION_CODE, SAMPLE_NO2, PARAM) %>%
  select(STATION_CODE, SAMPLE_NO2, PARAM, VALUE_WW) %>%
  pivot_wider(names_from = PARAM, values_from = VALUE_WW)
df$Ratio <- df[[3]]/df[[4]]
df  # Ratio = 2

```


## Data in wide form  

### Cod   
```{r}

# Variables
# data_biol_effects %>% names() %>% dput() %>% paste(collapse = ", ")

df1 <- data_biol_effects %>% 
  filter(LATIN_NAME %in% c("Gadus morhua")) %>%
  select(MYEAR, STATION_CODE, LATIN_NAME, TISSUE_NAME, SAMPLE_NO2, PARAM, DRYWT, FAT_PERC, LNMEA, VALUE_WW) %>%
  tidyr::pivot_wider(names_from = PARAM, values_from = VALUE_WW) %>%
  arrange(MYEAR, STATION_CODE, LATIN_NAME, TISSUE_NAME, SAMPLE_NO2)

# flags
df2 <- data_biol_effects %>% 
  filter(LATIN_NAME %in% c("Gadus morhua")) %>%
  select(MYEAR, STATION_CODE, LATIN_NAME, TISSUE_NAME, SAMPLE_NO2, PARAM, FLAG1) %>%
  mutate(PARAM = paste0(PARAM, "_flag")) %>%
  pivot_wider(names_from = PARAM, values_from = FLAG1)

# data_biol_effects_wide_snails
# View(df1)
# View(df2)

# Sort order of columns: key data (vars1), biol. effects (vars2), all the rest (vars3 = values, vars4 = flags)
vars1 <- c("MYEAR", "STATION_CODE", "LATIN_NAME", "TISSUE_NAME", "SAMPLE_NO2", "DRYWT", "FAT_PERC", "LNMEA", "DRYWT")
vars2a <- c("ALAD", "BAP3O", "BAP3OH", "PA1O", "PA1OH", "PYR1O", "PYR1OH", "CYP1A", "AY380", "ABS380")
vars2b <- paste0(vars2a, "_flag")
vars2 <- matrix(c(vars2a,vars2b), nrow = 2, byrow = TRUE) %>% as.character()
vars3 <- names(df1)
vars3 <- vars3[!vars3 %in% c(vars1, vars2, "Fett", "DRYWT%")] 
vars4 <- names(df2)
vars4 <- vars4[!vars4 %in% c(vars1, vars2, "Fett_flag", "DRYWT%_flag")] 
vars <- c(vars1, vars2, 
          sort(c(vars3, vars4))
)
# vars

data_biol_effects_wide_cod <- left_join(df1, df2, by = c("MYEAR", "STATION_CODE", "LATIN_NAME", "TISSUE_NAME", "SAMPLE_NO2"))[vars]

d <- dim(data_biol_effects_wide_cod)
cat("Wide cod data extracted,",  d[1], "rows and", d[2], "columns \n")

```

### Snails, wide format  
```{r}

# Variables
# data_biol_effects %>% names() %>% dput() %>% paste(collapse = ", ")

df1 <- data_biol_effects %>% 
  filter(LATIN_NAME %in% c("Littorina littorea", "Nucella lapillus")) %>%
  select(MYEAR, STATION_CODE, LATIN_NAME, TISSUE_NAME, SAMPLE_NO2, PARAM, DRYWT, FAT_PERC, LNMEA, VALUE_WW) %>%
  tidyr::pivot_wider(names_from = PARAM, values_from = VALUE_WW) %>%
  arrange(MYEAR, STATION_CODE, LATIN_NAME, TISSUE_NAME, SAMPLE_NO2)

# flags
df2 <- data_biol_effects %>% 
  filter(LATIN_NAME %in% c("Littorina littorea", "Nucella lapillus")) %>%
  select(MYEAR, STATION_CODE, LATIN_NAME, TISSUE_NAME, SAMPLE_NO2, PARAM, FLAG1) %>%
  mutate(PARAM = paste0(PARAM, "_flag")) %>%
  pivot_wider(names_from = PARAM, values_from = FLAG1)

# data_biol_effects_wide_snails
# View(df1)
# View(df2)

# Sort order of columns: key data (vars1), biol. effects (vars2), all the rest (vars3 = values, vars4 = flags)
vars1 <- c("MYEAR", "STATION_CODE", "LATIN_NAME", "TISSUE_NAME", "SAMPLE_NO2", "DRYWT", "FAT_PERC", "LNMEA", "DRYWT")
vars2 <- c("VDSI", "Intersex")
vars3 <- names(df1)
vars3 <- vars3[!vars3 %in% c(vars1, vars2, "Fett", "DRYWT%")] 
vars4 <- names(df2)
vars4 <- vars4[!vars4 %in% c(vars1, vars2, "Fett_flag", "DRYWT%_flag")] 
vars <- c(vars1, vars2, 
          sort(c(vars3, vars4))
)

data_biol_effects_wide_snails <- left_join(df1, df2, 
                                           by = c("MYEAR", "STATION_CODE", "LATIN_NAME", "TISSUE_NAME", "SAMPLE_NO2"))[vars]

d <- dim(data_biol_effects_wide_snails)
cat("Wide snail data extracted,",  d[1], "rows and", d[2], "columns \n")

```

## Save to Excel  
```{r}

writexl::write_xlsx(
  list(
    `All data (long format)` = data_biol_effects,
    `Cod data (wide format)` = data_biol_effects_wide_cod,
    `Snail data (wide format)` = data_biol_effects_wide_snails
  ),
  "Data/502_Biological_effect_data_update2019.xlsx"
)

```

