---
title: "510_Industry_molluscs"
output: html_document
---

## 1. Packages
```{r, message=FALSE, warning=FALSE, results='hide'}

# 
library(readxl)
library(tidyr)
library(dplyr, warn.conflicts = FALSE)
library(purrr)
library(lubridate)
library(mgcv)
library(AICcmodavg)   # AICc()
library(ggplot2)
library(safejoin)     # installed from https://github.com/moodymudskipper/safejoin   

source("001_Add_trends_functions.R")  # Copied from '16_Trend_functions.R' in Milkys_2018
source("002_Utility_functions.R")
source("110_Medians_and_PROREF_functions.R")
source("510_Industry_molluscs_functions.R")

last_year <- 2019


#
# Define a couple of extra functions for shortening code
#
get_stats <- function(df){
  calc_models_one_station2(df, gam = TRUE, log = TRUE)$statistics_for_file
}

model_from_medians_stat <- function(...)
  model_from_medians(...)$statistics

```




## 2. Read data (individual level, with VALUE_WW etc.)
```{r}

last_file_list <- read_last_file(folder = "Data", pattern = "109_adjusted_data")
data_all <- last_file_list$data

```

## 3. Summarise data by medians and add proref  

### Summarise by median for each year/station/parameter  
```{r}

data_med_temp <- calculate_medians(data_all)

```

### Add Proref values    
```{r}

df_proref <- get_proref(folder = "Input_data")

data_med <- data_med_temp %>%
  left_join(df_proref %>% select(-N) %>% rename(Proref_median = Median), 
            by = c("LATIN_NAME", "TISSUE_NAME", "PARAM", "Basis"))

```


## 4. Time trend analysis   

### Testing  
```{r}

model_from_medians("HG", "Gadus morhua", "Muskel", 
                   "30B", "WWa", 1980:2019, data_med, plotname = "window", 
                   ggplot = TRUE)$statistics %>%
  select(Nplus, p_linear:Status)

```

## 5. Make standard plots   

### Medians 
```{r}

source("401 Plot time series functions.R", encoding = "UTF-8")

data_med2 <- data_med




```

### Station names  
```{r}

data_stations <- readxl::read_excel("Input_data/Kartbase.xlsx")

df_stationnames <- data_stations %>%
  select(stasjonskode, `catch LAT...40`, `catch LONG...41`, stasjonsnavn, `Til Rapport`) %>%
  rename(STATION_CODE = stasjonskode,
         Lat = `catch LAT...40`, Long = `catch LONG...41`, 
         Station_name = stasjonsnavn,
         Report_version_name = `Til Rapport`) %>%
  filter(!is.na(STATION_CODE)) %>%
  filter(!Station_name %in% "Risøy, Østerfjord")  # One duplicate (different names), we just remove 

# Check stations that might be missing
df_stationnames %>% filter(STATION_CODE %in% c("28A2","97A3","227G","19B","19N"))

# Add missing station
df_stationnames <- df_stationnames %>%
  bind_rows(
    tibble(STATION_CODE = "227G",
           Report_version_name = "Flatskjær (St. 227G)")
    )

# Check some other stations
df_stationnames %>% filter(STATION_CODE %in% c("I714", "I133"))


```

### Parameter and species names for graphs
Used in 'plot_single_series_medians_data' (see '_functions' file)  
```{r}

df_paramnames <- readxl::read_excel(
  "Input_data/Lookup table - parameter names for plots.xlsx")

df_speciesnames <- tibble(
  LATIN_NAME = c("Gadus morhua", "Mytilus edulis", "Nucella lapillus"),
  Species_name = c("Cod", "Blue mussel", "Dog whelk")
  )

```

### Get EQS limits  
```{r}

EQS_limits <- read_excel("Input_data/EQS_limits.xlsx", "EQS")[1:8] %>%
  filter(!is.na(PARAM)) %>%
  as.data.frame()
EQS_limits <- fact2char_df(EQS_limits)  # changes all factor variables to character variables
EQS_limits$Limit <- as.numeric(EQS_limits$`Grense brukt`)

EQS_limits <- EQS_limits %>%
  select(PARAM, Limit) %>% 
  filter(!is.na(PARAM))

```

### Big excel  
```{r}

fns <- dir("Big_excel_table", pattern = "Data_xl_.+.rds") %>% rev()
fns_date <- substr(fns, 9, 18)
fns_ver <- stringr::str_extract(fns, "ver[0-9]+") %>% stringr::str_extract("[0-9]+") %>% as.numeric()
file_versions <- tibble(Filename = fns, Date = fns_date, Version = fns_ver) %>%
  arrange(desc(Date), desc(Version))
# file_versions

# Picks the first 'Filename' in 'file_versions' - this should be the latest 'excel' file:
data_xl_name <- file_versions$Filename[1]
data_xl_lessthans <- readRDS(paste0("Big_excel_table/", data_xl_name))

cat("Read big excel table", sQuote(data_xl_name))

```

### Test 1
```{r}

source("401 Plot time series functions.R")
type <- "line"
ymax <- 0.45
# debugonce(plot_medians_and_trends)
plot_medians_and_trends(c("HG", "Gadus morhua", "Muskel", "30B",  "WW"), 
                             eqs_type = "line", xlim = c(1980, 2020), ylim = c(0, ymax))

```



