---
title: "511 - Industry data - plotting trend plots from medians, 2021 data"
output: html_document
---

## 0. -- COMMON CODE ---  
This code is used by teh parts for each area below  

### Packages
```{r, message=FALSE, warning=FALSE, results='hide'}

# 
library(readxl)
library(tidyr)
library(dplyr, warn.conflicts = FALSE)
library(purrr)
library(lubridate)
library(mgcv)
library(AICcmodavg)   # AICc()
library(ggplot2)
library(safejoin)     # installed from https://github.com/moodymudskipper/safejoin   

source("001_Add_trends_functions.R")  # Copied from '16_Trend_functions.R' in Milkys_2018
source("002_Utility_functions.R")
source("110_Medians_and_PROREF_functions.R")
source("510_Industry_data_functions.R")
source("401 Plot time series functions.R", encoding = "UTF-8")

last_year <- 2019


#
# Define a couple of extra functions for shortening code
#
get_stats <- function(df){
  calc_models_one_station2(df, gam = TRUE, log = TRUE)$statistics_for_file
}

model_from_medians_stat <- function(...)
  model_from_medians(...)$statistics

```

### Parameter and species names for graphs
Used in 'plot_single_series_medians_data' (see '_functions' file)  
```{r}

df_paramnames <- readxl::read_excel(
  "Input_data/Lookup table - parameter names for plots.xlsx")

df_speciesnames <- tibble(
  LATIN_NAME = c("Gadus morhua", "Mytilus edulis", "Nucella lapillus"),
  Species_name = c("Cod", "Blue mussel", "Dog whelk")
  )

```

### PROREF values  
```{r}
df_proref <- get_proref(folder = "Input_data")


```


### Get EQS limits  
```{r}

df_EQS <- read_excel("Input_data/EQS_limits.xlsx", "EQS")[1:8] %>%
  filter(!is.na(PARAM)) %>%
  as.data.frame()
df_EQS <- fact2char_df(df_EQS)  # changes all factor variables to character variables
df_EQS$Limit <- as.numeric(df_EQS$`Grense brukt`)

df_EQS <- df_EQS %>%
  select(PARAM, Limit) %>% 
  filter(!is.na(PARAM))

```

## .
## -- AORDAL -- 
## .

## 2. Read data (individual level, with VALUE_WW etc.)   
*Not needed here*      

## 3. Summarise data by medians and add proref  

### Summarise by median for each year/station/parameter   
*Not needed here*      

### Read median data    
From script 02 in 'NIVA-R-examples' project  
```{r}

  data_med_temp <- readRDS(
    "Input_data/510_Industry_data/Industry_data_aardal_(script_02).rds")
  
  names(data_med_temp)
  unique(data_med_temp$Basis)   # DW  DWa   FB  FBa   WW  WWa 


```


### Add Proref values    
```{r}

data_med <- data_med_temp %>%
  left_join(df_proref %>% select(-N) %>% rename(Proref_median = Median), 
            by = c("LATIN_NAME", "TISSUE_NAME", "PARAM", "Basis"))

```


## 4. Time trend analysis   

### Testing trend analysis   
Including simple plot  
```{r}


model_from_medians("PAH16", "Mytilus edulis", "Whole soft body", 
                   "G1", "WW", 2015:2020, data_med, plotname = "window", 
                   ggplot = TRUE)$statistics %>%
  select(Nplus, p_linear:Status)

```



## 5. Get ready for plotting     

### Station names  
```{r}

#
# Hand-made
#

if (TRUE){
  df_stationnames <- data.frame(
    STATION_CODE = c("G1", "G5", "G6", "G7", "G8"),
    Report_version_name = c("Hundshammar (G1)", "Kolnosi (G5)",
                            "Naddvik (G6)", "Kvitingsagi (G7)", "Ytre Offerdal (G8)"),
    stringsAsFactors = FALSE)
}

#
# Or read from excel file
#

if (FALSE){
  
  data_stations <- readxl::read_excel("Input_data/Kartbase.xlsx")
  
  df_stationnames <- data_stations %>%
    select(stasjonskode, `catch LAT...40`, `catch LONG...41`, stasjonsnavn, `Til Rapport`) %>%
    rename(STATION_CODE = stasjonskode,
           Lat = `catch LAT...40`, Long = `catch LONG...41`, 
           Station_name = stasjonsnavn,
           Report_version_name = `Til Rapport`) %>%
    filter(!is.na(STATION_CODE)) %>%
    filter(!Station_name %in% "Risøy, Østerfjord")  # One duplicate (different names), we just remove 
  
  # Check stations that might be missing
  df_stationnames %>% filter(STATION_CODE %in% c("28A2","97A3","227G","19B","19N"))
  
  # Add missing station
  df_stationnames <- df_stationnames %>%
    bind_rows(
      tibble(STATION_CODE = "227G",
             Report_version_name = "Flatskjær (St. 227G)")
    )
  
  # Check some other stations
  df_stationnames %>% filter(STATION_CODE %in% c("I714", "I133"))
  
}

```



### Test 1
```{r}
# debugonce(model_from_medians)
# debugonce(statistics_for_excel)

X <- model_from_medians("BAP", "Mytilus edulis", "Whole soft body", "G1", "WW", 
                                        2016:2020,  
                                        data_medians = data_med, 
                                        ggplot = TRUE)
X$statistics

```


### Test 2
```{r}

source("401 Plot time series functions.R")
type <- "line"
ymax <- 0.45
# debugonce(plot_medians_and_trends2)
# debugonce(plot_medians_color)

# debugonce(plot_medians_and_trends2)
X <- plot_medians_and_trends2(c("BAP", "Mytilus edulis", "Whole soft body", "G1", "WW"), 
                         eqs_type = "line", xlim = c(2004, 2020), 
                         data_medians = data_med, data_proref = df_proref, 
                         data_eqs = df_EQS)

save_trendplot(X, "Input_data/510_Industry_data/Plots_Årdal_2020")
X$gg


```


## 6. All plots   

### Select series (parameter + station) to plot  
```{r}

tab <- xtabs(~ PARAM + STATION_CODE, data_med)
# tab

df_series <- data_med %>%
  count(PARAM, STATION_CODE) %>%
  filter(n >= 2)



```

### Plot all  
```{r}

folder <- "Input_data/510_Industry_data/Plots_Årdal_2020"

for (i in seq_len(nrow(df_series))){
  
  # i <- 1
  X <- try(
    plot_medians_and_trends2(c(df_series$PARAM[i], "Mytilus edulis", "Whole soft body", 
                               df_series$STATION_CODE[i], "WW"), 
                             eqs_type = "line", xlim = c(2004, 2020), 
                             data_medians = data_med, data_proref = df_proref, 
                             data_eqs = df_EQS)
  )
  if (class(X) != "try-error"){
    save_trendplot(X, folder)
  }
  
}

```


## .
## -- 15B FARSUND ---
## .

## 1. Read data   
Median data produced by script 11    
```{r}

files <- dir("Data", pattern = "110_mediandata_updated_") %>% rev()

cat("Reading the last file downloaded:")
cat("\n", files[1])
cat("\n")
cat("If you want to read a different file, replace 'files[1]' with the file you want")
cat("\n")

filename <- paste0("Data/", files[1]) 

# Medians
data_med2 <- readRDS(file = filename) %>%
  rename(Proref_median = Median,
         Median = Value) 

```

### Station names  
```{r}

  
  data_stations <- readxl::read_excel("Input_data/Kartbase.xlsx")
  
  df_stationnames <- data_stations %>%
    select(stasjonskode, `catch LAT...40`, `catch LONG...41`, stasjonsnavn, `Til Rapport`) %>%
    rename(STATION_CODE = stasjonskode,
           Lat = `catch LAT...40`, Long = `catch LONG...41`, 
           Station_name = stasjonsnavn,
           Report_version_name = `Til Rapport`) %>%
    filter(!is.na(STATION_CODE)) %>%
    filter(!Station_name %in% "Risøy, Østerfjord")  # One duplicate (different names), we just remove 
  

```

## 2. Test plot     
```{r}

source("401 Plot time series functions.R")
type <- "line"
ymax <- 0.45
# debugonce(plot_medians_and_trends2)
# debugonce(plot_medians_color)

# debugonce(plot_medians_and_trends2)
X <- plot_medians_and_trends2(c("CD", "Gadus morhua", "Lever", "15B", "WW"), 
                         eqs_type = "line", xlim = c(2004, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         data_eqs = df_EQS)

save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019")
X$gg


```



## 3. Plot all  
### Metals
```{r}

# debugonce(plot_medians_and_trends2)

# CD, showing the PROREF level
X <- plot_medians_and_trends2(c("CD", "Gadus morhua", "Lever", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         ylim_proref = 1,
                         data_medians = data_med2, data_proref = df_proref, 
                         data_eqs = df_EQS)

save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019", suffix = "_ver1")
X$gg

# CD, not showing PROREF level
X <- plot_medians_and_trends2(c("CD", "Gadus morhua", "Lever", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         data_eqs = df_EQS)

save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019", suffix = "_ver2")
X$gg


# AG
X <- plot_medians_and_trends2(c("AG", "Gadus morhua", "Lever", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         ylim = c(0, 2),
                         data_eqs = df_EQS)
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019")
X$gg

# CR
X <- plot_medians_and_trends2(c("CR", "Gadus morhua", "Lever", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         # ylim = c(0, 2),
                         data_eqs = df_EQS)
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019")
X$gg


# CU
X <- plot_medians_and_trends2(c("CU", "Gadus morhua", "Lever", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         ylim = c(0, 18),
                         data_eqs = df_EQS)
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019")
X$gg

# NI
X <- plot_medians_and_trends2(c("NI", "Gadus morhua", "Lever", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         # ylim = c(0, 18),
                         data_eqs = df_EQS)
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019")
X$gg

# PB
X <- plot_medians_and_trends2(c("PB", "Gadus morhua", "Lever", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         ylim = c(0, 0.15),
                         data_eqs = df_EQS)
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019")
X$gg

# SN
X <- plot_medians_and_trends2(c("SN", "Gadus morhua", "Lever", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         ylim = c(0, 0.4),
                         data_eqs = df_EQS)
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019")
X$gg


# ZN
X <- plot_medians_and_trends2(c("ZN", "Gadus morhua", "Lever", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         ylim = c(0, 65),
                         data_eqs = df_EQS)
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019")
X$gg


```


### OH-pyren 1 - Bømlo as reference level  

```{r}


# PYR1OH - non-normalised OH-pyren

# Since there is no PROREF for PYR1OH, we add a reference level based on Bømlo 
# Reference level: max of the annual medians at Bømlo 
reference_level <- data_med2 %>% 
  filter(PARAM %in% "PYR1OH" & STATION_CODE %in% "23B" & MYEAR >= 2010 & Basis == "WW") %>% # View()
  pull(Median) %>%
  max()

X <- plot_medians_and_trends2(c("PYR1OH", "Gadus morhua", "Galle", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         # ylim = c(0, 65),
                         data_eqs = df_EQS)
# Modify graph to add reference level
X$gg <- X$gg +
  geom_hline(yintercept = reference_level, linetype = 2, col = "grey25") +
  annotate("text", x = 1982, y = reference_level, label = "Reference station max. (2015-2019)", 
           hjust = 0, vjust = -0.5, size = 3.5)
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019")
X$gg



#
# PYR1O - normalised OH_pyren
#

# As above, jsut ith PYR1O
reference_level <- data_med2 %>% 
  filter(PARAM %in% "PYR1O" & STATION_CODE %in% "23B" & MYEAR >= 2010 & Basis == "WW") %>% # View()
  pull(Median) %>%
  max()

# 2. PYR10 (saved as version 1)
X <- plot_medians_and_trends2(c("PYR1O", "Gadus morhua", "Galle", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         # ylim = c(0, 65),
                         data_eqs = df_EQS)

# Modify graph to add reference level
X$gg <- X$gg +
  geom_hline(yintercept = reference_level, linetype = 2, col = "grey25") +
  annotate("text", x = 1982, y = reference_level, label = "Reference station max. (2015-2019)", 
           hjust = 0, vjust = -0.5, size = 3.5)
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019", suffix = "_ver1")
X$gg

#
# 3. PYR1O, with max limit = 35 (saved as version 2)
#
X <- plot_medians_and_trends2(c("PYR1O", "Gadus morhua", "Galle", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         ylim = c(0, 35),
                         data_eqs = df_EQS)
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019", suffix = "_ver2")
X$gg


```



### OH-pyren 1 - OSPAR BAC (= 21) as reference level     
https://ocean.ices.dk/ohat/trDocuments/2020/help_ac_biota_metabolites.html   
```{r}


# For 2001-2004, we have PYR10 but lack PYR1OH (non-normalised OH-pyren) data   
# We estimate PYR1OH form PYR1O   
data_to_add <- data_med2 %>% 
  filter(PARAM %in% "PYR1O" & STATION_CODE %in% "15B" & MYEAR %in% 2001:2004 & Basis == "WW") 

# Find mean PYR1OH/PYR1O ratio 2005-2007
ratios <- data_med2 %>% 
  filter(PARAM %in% c("PYR1OH","PYR1O") & STATION_CODE %in% "15B" & MYEAR %in% 2005:2007 & Basis == "WW") %>%
  pivot_wider(names_from = "PARAM", values_from = "Median") %>%
  mutate(Ratio = PYR1OH/PYR1O)
mean(ratios$Ratio)

# Mean ra
data_to_add$PARAM <- "PYR1OH"
data_to_add$Median <- data_to_add$Median*mean(ratios$Ratio)

data_for_plot1 <- bind_rows(data_med2, data_to_add)

# debugonce(plot_medians_and_trends2)
# debugonce(get_plotdata)
X <- plot_medians_and_trends2(c("PYR1OH", "Gadus morhua", "Galle", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1997, 2020), 
                         data_medians = data_for_plot1 %>% filter(MYEAR >= 2000),
                         data_proref = df_proref, 
                         # ylim = c(0, 65),
                         data_eqs = df_EQS)

# Modify graph to add reference level
X$gg <- X$gg +
  geom_hline(yintercept = reference_level, linetype = 2, col = "red2") +
  annotate("text", x = 1997, y = 21, label = "OSPAR BAC", 
           hjust = 0, vjust = -0.8, size = 5, col = "red2")

X$gg  
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019", suffix = "_ver2")




```

```{r}



    
# PYR1OH - non-nomalised OH-pyren
   
# Since there is no PROREF for PYR1OH, we add a reference level based on Bømlo 
# Reference level: max of the annual medians at Bømlo 
reference_level <- data_med2 %>% 
  filter(PARAM %in% "PYR1OH" & STATION_CODE %in% "23B" & MYEAR >= 2010 & Basis == "WW") %>% # View()
  pull(Median) %>%
  max()

X <- plot_medians_and_trends2(c("PYR1OH", "Gadus morhua", "Galle", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         # ylim = c(0, 65),
                         data_eqs = df_EQS)
# Modify graph to add reference level
X$gg <- X$gg +
  geom_hline(yintercept = reference_level, linetype = 2, col = "grey25") +
  annotate("text", x = 1982, y = reference_level, label = "Reference station max. (2015-2019)", 
           hjust = 0, vjust = -0.5, size = 3.5)
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019")
X$gg



#
# PYR1O - normalised OH_pyren
#

# As above, jsut ith PYR1O
reference_level <- data_med2 %>% 
  filter(PARAM %in% "PYR1O" & STATION_CODE %in% "23B" & MYEAR >= 2010 & Basis == "WW") %>% # View()
  pull(Median) %>%
  max()

# 2. PYR10 (saved as version 1)
X <- plot_medians_and_trends2(c("PYR1O", "Gadus morhua", "Galle", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         # ylim = c(0, 65),
                         data_eqs = df_EQS)

# Modify graph to add reference level
X$gg <- X$gg +
  geom_hline(yintercept = reference_level, linetype = 2, col = "grey25") +
  annotate("text", x = 1982, y = reference_level, label = "Reference station max. (2015-2019)", 
           hjust = 0, vjust = -0.5, size = 3.5)
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019", suffix = "_ver1")
X$gg

#
# 3. PYR1O, with max limit = 35 (saved as version 2)
#
X <- plot_medians_and_trends2(c("PYR1O", "Gadus morhua", "Galle", "15B", "WW"), 
                         eqs_type = "line", xlim = c(1982, 2020), 
                         data_medians = data_med2, data_proref = df_proref, 
                         ylim = c(0, 35),
                         data_eqs = df_EQS)
save_trendplot(X, "Input_data/510_Industry_data/Plots_Lista_2019", suffix = "_ver2")
X$gg


```


## .
## -- KRISTIANSAND --  
## .



### 1. Read data medians      
From script 994 in 'Milkys' project  
```{r}

# [1] "STATION_CODE" "STATION_NAME" "LATIN_NAME"   "TISSUE_NAME"  "PARAM"        "UNIT"         "MYEAR"       
# [8] "N"            "Median"       "Over_LOQ"     "Min"          "Max"          "Basis" 

# dir("Input_data/510_Industry_data")

# Measurement level  
data_med_temp <- readRDS(
  "Input_data/510_Industry_data/data_chem_industry_kristiansand_med(script_994).rds")

data_med_temp %>%
  filter(PARAM %in% "BAP") %>%
  xtabs(~STATION_CODE + MYEAR, .)

```


### Add Proref values    
```{r}

data_med <- data_med_temp %>%
  left_join(df_proref %>% select(-N) %>% rename(Proref_median = Median), 
            by = c("LATIN_NAME", "TISSUE_NAME", "PARAM", "Basis"))

```


## 4. Time trend analysis   

### Testing trend analysis   
Including simple plot  
```{r}

# xtabs(~STATION_CODE, data_med)

# data_med_temp %>%
#   filter(PARAM %in% "CD") %>%
#   select(PARAM, LATIN_NAME, TISSUE_NAME, STATION_CODE, Basis, Median)

model_from_medians("CD", "Mytilus edulis", "Whole soft body", 
                   "St. 1", "WW", 2008:2020, data_med, plotname = "window", 
                   ggplot = TRUE)$statistics %>%
  select(Nplus, p_linear:Status)

model_from_medians("BAP", "Mytilus edulis", "Whole soft body", 
                   "St. 4", "WW", 2008:2020, data_med, plotname = "window", 
                   ggplot = TRUE)$statistics %>%
  select(Nplus, p_linear:Status)

```



## 5. Get ready for plotting     

### Station names  
```{r}

#
# Hand-made
#

if (TRUE){
  df_stationnames <- data.frame(
    STATION_CODE = c("St. 1", "St. 2", "St. 3", "St. 4", "St. 5"),
    Report_version_name = c("St. 1 (Lumber)", "St. 2 (Fiskå)", "St. 3 (Timlingen)",
                            "St. 4 (Svensholmen)", "St. 5 (Flekkerøygapet)"),
    stringsAsFactors = FALSE)
}

#
# Or read from excel file
#

if (FALSE){
  
  data_stations <- readxl::read_excel("Input_data/Kartbase.xlsx")
  
  df_stationnames <- data_stations %>%
    select(stasjonskode, `catch LAT...40`, `catch LONG...41`, stasjonsnavn, `Til Rapport`) %>%
    rename(STATION_CODE = stasjonskode,
           Lat = `catch LAT...40`, Long = `catch LONG...41`, 
           Station_name = stasjonsnavn,
           Report_version_name = `Til Rapport`) %>%
    filter(!is.na(STATION_CODE)) %>%
    filter(!Station_name %in% "Risøy, Østerfjord")  # One duplicate (different names), we just remove 
  
  # Check stations that might be missing
  df_stationnames %>% filter(STATION_CODE %in% c("28A2","97A3","227G","19B","19N"))
  
  # Add missing station
  df_stationnames <- df_stationnames %>%
    bind_rows(
      tibble(STATION_CODE = "227G",
             Report_version_name = "Flatskjær (St. 227G)")
    )
  
  # Check some other stations
  df_stationnames %>% filter(STATION_CODE %in% c("I714", "I133"))
  
}

```



### Test 1
```{r}
# debugonce(model_from_medians)
# debugonce(statistics_for_excel)

X <- model_from_medians("CD", "Mytilus edulis", "Whole soft body", 
                        "St. 1", "WW", 2015:2020,  
                        data_medians = data_med, 
                        ggplot = TRUE)
X$statistics

```


### Test 2
```{r}

source("401 Plot time series functions.R")
type <- "line"
ymax <- 0.45
# debugonce(plot_medians_and_trends2)
# debugonce(plot_medians_color)

# debugonce(plot_medians_and_trends2)
X <- plot_medians_and_trends2(c("BAP", "Mytilus edulis", "Whole soft body", "St. 1", "WW", 2015:2020), 
                         eqs_type = "line", xlim = c(2007, 2020), 
                         data_medians = data_med, data_proref = df_proref, 
                         data_eqs = df_EQS)

# save_trendplot(X, "Input_data/510_Industry_data/Plots_Kristiansand_2020")
X$gg


```


## 6. All plots   

### Select series (parameter + station) to plot  
```{r}

data_med$PARAM <- sub("%", "", data_med$PARAM)

tab <- xtabs(~ PARAM + STATION_CODE, data_med)
# tab

df_series <- data_med %>%
  filter(MYEAR >= 2016) %>%
  count(PARAM, STATION_CODE) %>%
  filter(n >= 2)


```

### Plot folder  
```{r}
folder <- "Input_data/510_Industry_data/Plots_Kristiansand_2020"

```


### Plot all  
```{r}

df_series[71,]
# nrow(df_series)
for (i in 1:70){
# for (i in seq(1, nrow(df_series))){
  
  # i <- 1
  X <- try(
    plot_medians_and_trends2(c(df_series$PARAM[i], "Mytilus edulis", "Whole soft body", 
                               df_series$STATION_CODE[i], "WW"), 
                             eqs_type = "line", xlim = c(2007, 2020), 
                             data_medians = data_med, data_proref = df_proref, 
                             data_eqs = df_EQS)
  )
  if (class(X) != "try-error"){
    save_trendplot(X, folder)
  }
  
}

```



#### Improve BAP (benzo[a]pyren) plots  
```{r}

param <- "BAP"
stations <- unique(df_series$STATION_CODE)

for (station in stations){
  
  if (station == "St. 1"){
    
    X <- plot_medians_and_trends2(c(param, "Mytilus edulis", "Whole soft body", station, "WW"), 
                             eqs_x = 2006.5, proref_x = 2008, show_proref = 5,
                             eqs_type = "line", xlim = c(2006, 2020), eqs_label_transparent = FALSE,
                             data_medians = data_med, data_proref = df_proref, 
                             data_eqs = df_EQS)
    
  } else if (station == "St. 4"){
    
    X <- plot_medians_and_trends2(c(param, "Mytilus edulis", "Whole soft body", station, "WW"), 
                             eqs_x = 1991.5, proref_x = 1994, 
                             eqs_type = "line", xlim = c(1992, 2020), eqs_label_transparent = FALSE,
                             data_medians = data_med, data_proref = df_proref, 
                             data_eqs = df_EQS)
    
    
  } else {
    
    X <- plot_medians_and_trends2(c(param, "Mytilus edulis", "Whole soft body", station, "WW"), 
                             eqs_x = 2006.5, proref_x = 2008, 
                             eqs_type = "line", xlim = c(2006, 2020), eqs_label_transparent = FALSE,
                             data_medians = data_med, data_proref = df_proref, 
                             data_eqs = df_EQS)
    
    
  }
  
  print(X$gg)
  
  save_trendplot(X, folder)

}

  
```

#### Improve PAH16 plots  
```{r}

param <- "PAH16"
stations <- unique(df_series$STATION_CODE)

for (station in stations){
  
  if (station == "St. 1"){
    
    X <- plot_medians_and_trends2(c(param, "Mytilus edulis", "Whole soft body", station, "WW"), 
                             eqs_x = 2006.5, proref_x = 2006.5, show_proref = 5,
                             eqs_type = "line", xlim = c(2006, 2020), eqs_label_transparent = FALSE,
                             data_medians = data_med, data_proref = df_proref, 
                             data_eqs = df_EQS)
    
  } else if (station == "St. 4"){
    
    X <- plot_medians_and_trends2(c(param, "Mytilus edulis", "Whole soft body", station, "WW"), 
                             eqs_x = 1991.5, proref_x = 1993, 
                             eqs_type = "line", xlim = c(1992, 2020), eqs_label_transparent = FALSE,
                             data_medians = data_med, data_proref = df_proref, 
                             data_eqs = df_EQS)
    
    
  } else {
    
    X <- plot_medians_and_trends2(c(param, "Mytilus edulis", "Whole soft body", station, "WW"), 
                             eqs_x = 2006.5, proref_x = 2006.5, 
                             eqs_type = "line", xlim = c(2006, 2020), eqs_label_transparent = FALSE,
                             data_medians = data_med, data_proref = df_proref, 
                             data_eqs = df_EQS)
    
    
  }
  
  
  save_trendplot(X, folder)

  print(X$gg)

  }

  
```


## .
## -- RANFJORDEN --  
## .



### 1. Read data medians      
From script 994 in 'Milkys' project  
```{r}

# [1] "STATION_CODE" "STATION_NAME" "LATIN_NAME"   "TISSUE_NAME"  "PARAM"        "UNIT"         "MYEAR"       
# [8] "N"            "Median"       "Over_LOQ"     "Min"          "Max"          "Basis" 

# dir("Input_data/510_Industry_data")

# Measurement level  
data_med_temp <- readRDS(
  "Input_data/510_Industry_data/data_chem_industry_ranfjorden_med(script_994).rds")


xtabs(~MYEAR + STATION_CODE, data_med_temp)

```


### Add Proref values    
```{r}

data_med <- data_med_temp %>%
  left_join(df_proref %>% select(-N) %>% rename(Proref_median = Median), 
            by = c("LATIN_NAME", "TISSUE_NAME", "PARAM", "Basis"))

```


## 4. Time trend analysis   

### Testing trend analysis   
Including simple plot  
```{r}

# xtabs(~STATION_CODE, data_med)

model_from_medians("CD", "Mytilus edulis", "Whole soft body", 
                   "I969", "WW", 2000:2020, data_med, plotname = "window", 
                   ggplot = TRUE)$statistics %>%
  select(Nplus, p_linear:Status)

model_from_medians("PAH16_woLOQ", "Mytilus edulis", "Whole soft body", 
                   "I965", "WW", 2000:2020, data_med, plotname = "window", 
                   ggplot = TRUE)$statistics %>%
  select(Nplus, p_linear:Status)

```



## 5. Get ready for plotting     

### Station names  
```{r}

#
# Hand-made
#

if (TRUE){
  df_stationnames <- data.frame(
    STATION_CODE = c("I964/I964b", "I965", "I969"),
    Report_version_name = c("Toraneskaien (I964/I964b)", "Moholmen (I965)", "Bjørnbærviken (I969)"),
    stringsAsFactors = FALSE)
}

#
# Or read from excel file
#

if (FALSE){
  
  data_stations <- readxl::read_excel("Input_data/Kartbase.xlsx")
  
  df_stationnames <- data_stations %>%
    select(stasjonskode, `catch LAT...40`, `catch LONG...41`, stasjonsnavn, `Til Rapport`) %>%
    rename(STATION_CODE = stasjonskode,
           Lat = `catch LAT...40`, Long = `catch LONG...41`, 
           Station_name = stasjonsnavn,
           Report_version_name = `Til Rapport`) %>%
    filter(!is.na(STATION_CODE)) %>%
    filter(!Station_name %in% "Risøy, Østerfjord")  # One duplicate (different names), we just remove 
  
  # Check stations that might be missing
  df_stationnames %>% filter(STATION_CODE %in% c("28A2","97A3","227G","19B","19N"))
  
  # Add missing station
  df_stationnames <- df_stationnames %>%
    bind_rows(
      tibble(STATION_CODE = "227G",
             Report_version_name = "Flatskjær (St. 227G)")
    )
  
  # Check some other stations
  df_stationnames %>% filter(STATION_CODE %in% c("I714", "I133"))
  
}

```



### Test 1   
```{r}
# debugonce(model_from_medians)
# debugonce(statistics_for_excel)

X <- model_from_medians("CD", "Mytilus edulis", "Whole soft body", 
                        "I965", "WW", 2010:2020,  
                        data_medians = data_med, 
                        ggplot = TRUE)
X$statistics

```


### Test 2
```{r}

source("401 Plot time series functions.R")
type <- "line"
ymax <- 0.45
# debugonce(plot_medians_and_trends2)
# debugonce(plot_medians_color)

# debugonce(plot_medians_and_trends2)
X <- plot_medians_and_trends2(c("ACNE", "Mytilus edulis", "Whole soft body", 
                                "I964/I964b", "WW", 2010:2020), 
                         eqs_type = "line", xlim = c(2007, 2020), 
                         data_medians = data_med, data_proref = df_proref, 
                         data_eqs = df_EQS)

# save_trendplot(X, "Input_data/510_Industry_data/Plots_Kristiansand_2020")
X$gg

```


## 6. All plots   

### Select series (parameter + station) to plot  
```{r}

# Dry weight  
data_med$PARAM <- sub("%", "", data_med$PARAM)

tab <- xtabs(~ PARAM + STATION_CODE, data_med)
# tab

df_series <- data_med %>%
  filter(MYEAR >= 2016) %>%
  count(PARAM, STATION_CODE) %>%
  filter(n >= 2)


```

### Plot all  
```{r}

folder <- "Input_data/510_Industry_data/Plots_Ranfjorden_2020"

# nrow(df_series)
# for (i in seq_len(nrow(df_series))){
for (i in seq(1, nrow(df_series))){
  
  # i <- 1
  X <- try(
    plot_medians_and_trends2(c(df_series$PARAM[i], "Mytilus edulis", "Whole soft body", 
                               df_series$STATION_CODE[i], "WW"), 
                             eqs_type = "line", xlim = c(2007, 2020), 
                             data_medians = data_med, data_proref = df_proref, 
                             data_eqs = df_EQS)
    
  )
  if (class(X) != "try-error"){
    # We need to avoid slash in filenames....:
    X$fn <- sub("I964/I964b", "I964", X$fn, fixed = TRUE)
    # Save
    save_trendplot(X, folder)
  }
  
}

```


