---
title: "512_Hvaler_fucus_bluemussel"
output: html_document
---



## Common code (as 511)

### Packages
```{r, message=FALSE, warning=FALSE, results='hide'}

# 
library(readxl)
library(tidyr)
library(dplyr, warn.conflicts = FALSE)
library(purrr)
library(lubridate)
library(mgcv)
library(AICcmodavg)   # AICc()
library(ggplot2)
library(safejoin)     # installed from https://github.com/moodymudskipper/safejoin   

source("001_Add_trends_functions.R")  # Copied from '16_Trend_functions.R' in Milkys_2018
source("002_Utility_functions.R")
source("110_Medians_and_PROREF_functions.R")
source("510_Industry_data_functions.R")
source("401 Plot time series functions.R", encoding = "UTF-8")

last_year <- 2021


#
# Define a couple of extra functions for shortening code
#
get_stats <- function(df){
  calc_models_one_station2(df, gam = TRUE, log = TRUE)$statistics_for_file
}

model_from_medians_stat <- function(...)
  model_from_medians(...)$statistics

```

### Parameter and species names for graphs
Used in 'plot_single_series_medians_data' (see '_functions' file)  
```{r}

df_paramnames <- readxl::read_excel(
  "Input_data/Lookup table - parameter names for plots.xlsx")

df_speciesnames <- tibble(
  LATIN_NAME = c("Gadus morhua", "Mytilus edulis", "Nucella lapillus"),
  Species_name = c("Cod", "Blue mussel", "Dog whelk")
  )

```


### PROREF values  
```{r}

df_proref <- get_proref(folder = "Input_data")

```

### Get EQS limits  
```{r}

df_EQS <- read_excel("Input_data/EQS_limits.xlsx", "EQS")[1:8] %>%
  filter(!is.na(PARAM)) %>%
  as.data.frame()
df_EQS <- fact2char_df(df_EQS)  # changes all factor variables to character variables
df_EQS$Limit <- as.numeric(df_EQS$`Grense brukt`)

df_EQS <- df_EQS %>%
  select(PARAM, Limit) %>% 
  filter(!is.na(PARAM))

```

## Fucus (blæretang)  
* bladder wrack *Fucus vesicolus*   

### Data   
* Copied from DHJs pc: Data/Seksjon 212/Glomma-estuary-metals
* Made using code from that project (script 01), also see github NIVANorge 'Glomma-estuary-metals'
```{r}

dat1 <- read.csv("Input_data/510_Industry_data/data_glomma_estuary_rockweed_ts.csv",
                 row.names = "X") %>%
  filter(!is.na(Value))

# check that encoding works  
# table(dat1$Station_name)

```

### Check data  
```{r}

xtabs(~PARAM, dat1)

```

### Fix names  
```{r}

names(dat1)

dat2 <- dat1 %>%
  rename(
    MYEAR = Year,
    Median = Value) %>%
  mutate(
    N_median = 3,
    Over_LOQ = 3
  ) 

# names(data_med)
#  [1] "STATION_CODE"  "LATIN_NAME"    "TISSUE_NAME"   "PARAM"         "UNIT"          "MYEAR"         "N_median"      "Median"        "Over_LOQ"     
# [10] "Min"           "Max"           "Basis"         "NAME"          "STATION_NAME"  "Stations"      "N_stations"    "Proref_median" "Q95"

```


### Make data_med      
```{r}

data_med <- dat2 %>%
  mutate(
    LATIN_NAME = "blæretang",
    TISSUE_NAME = "skudd",
    STATION_CODE = Station_name,
    Basis = "WW",
    UNIT = "MG_P_KG"
  ) #%>%
  #left_join(df_proref %>% select(-N) %>% rename(Proref_median = Median), 
  #          by = c("LATIN_NAME", "TISSUE_NAME", "PARAM", "Basis"))

# data_med %>% filter(PARAM == "BAP" & STATION_CODE == "B2") %>% View()

```



### Select series (parameter + station) to plot   
* Entire time series  
* Since we haven't removed NAs, min_year is alwyas 1994.... but that's OK here  
```{r}

tab <- xtabs(~ PARAM + STATION_CODE, data_med)
# tab

df_series <- data_med %>%
  # filter(!is.na(Q95)) %>%            # stations with at least 2 years of data, and existing PROREF
  group_by(PARAM, STATION_CODE, LATIN_NAME, TISSUE_NAME, Basis) %>%
  summarise(
    n = n(),
    min_year = min(MYEAR)
  ) %>%
  filter(n >= 2)     

```
### Test plot
```{r}

# debugonce(plot_medians_and_trends3)
# debugonce(calc_models_gam)
# debugonce(statistics_for_excel)

i <- 12
startyear <- df_series$min_year[i]
test <-  plot_medians_and_trends3(c(df_series$PARAM[i], 
                                    df_series$LATIN_NAME[i], df_series$TISSUE_NAME[i], 
                                    df_series$STATION_CODE[i], df_series$Basis[i]), 
                                  data_medians = data_med,
                                  trend_years = startyear:2021, trend_vjust = 1.6, trend_hjust = 2.2) 

test$gg
test$fn

# Manual ggplot
# test2 <- select_data(c(df_series$PARAM[i], 
#                                     df_series$LATIN_NAME[i], df_series$TISSUE_NAME[i], 
#                                     df_series$STATION_CODE[i], df_series$Basis[i]), 
#                                   data_medians = data_med)
# gg <- ggplot(test2, aes(MYEAR))
# gg <- gg + geom_smooth(aes(y = Median), method = "lm")
# gg 

#  annotate("text", x = Inf, y = Inf, label = "\u25CB", size = rel(13))

```

### Plot entire time series (using defaults) 

```{r}

folder <- "Input_data/510_Industry_data/Plots_Glomma_estuary/tang"
# dir.create(folder)

for (i in seq_len(nrow(df_series))){

  # For testing (make only plots 1 to 4), comment out the line above and uncomment the next line:
  # for (i in 1:10){
  # i <- 1

  startyear <- df_series$min_year[i]
  
  debugonce(plot_medians_and_trends2)
  # debug(calc_models_one_station2)
  # debug(add_model_fits)
  X <-  plot_medians_and_trends3(c(df_series$PARAM[i], 
                                    df_series$LATIN_NAME[i], df_series$TISSUE_NAME[i], 
                                    df_series$STATION_CODE[i], df_series$Basis[i]), 
                                  data_medians = data_med,
                                  trend_years = startyear:2021, trend_vjust = 1.6, trend_hjust = 2.2) 

  # Fix x and y labels
  X$gg <- X$gg + 
    labs(x = "År", y = "Konsentrasjon")
  
  # X$gg
  ggsave("Input_data/510_Industry_data/Plots_Glomma_estuary/CR_test.png", X$gg)
  
  if (class(X) != "try-error"){
    save_trendplot(X, folder, suffix = "_allyears")
  }
  
}   

```




## Blue mussel

### Data   
* Copied from DHJs pc: Data/Seksjon 212/Glomma-estuary-metals
* Made using code from that project (script 01), also see github NIVANorge 'Glomma-estuary-metals'
```{r}

dat1 <- read.csv("Input_data/510_Industry_data/data_glomma_estuary_bluemussel.csv", row.names = "X")

```

### Check data  
```{r}

xtabs(~PARAM, dat1)

```

### Fix names  
```{r}

names(dat1)

dat2 <- dat1 %>%
  rename(
    MYEAR = Year,
    Median = Value) %>%
  mutate(
    N_median = 3,
    Over_LOQ = 3
  ) 

# names(data_med)
#  [1] "STATION_CODE"  "LATIN_NAME"    "TISSUE_NAME"   "PARAM"         "UNIT"          "MYEAR"         "N_median"      "Median"        "Over_LOQ"     
# [10] "Min"           "Max"           "Basis"         "NAME"          "STATION_NAME"  "Stations"      "N_stations"    "Proref_median" "Q95"

```

### Make data_med      
```{r}

data_med <- dat2 %>%
  mutate(
    LATIN_NAME = "blæretang",
    TISSUE_NAME = "skudd",
    STATION_CODE = Station_name,
    Basis = "WW",
    UNIT = "MG_P_KG"
  ) #%>%
  #left_join(df_proref %>% select(-N) %>% rename(Proref_median = Median), 
  #          by = c("LATIN_NAME", "TISSUE_NAME", "PARAM", "Basis"))

# data_med %>% filter(PARAM == "BAP" & STATION_CODE == "B2") %>% View()

```




### Select series (parameter + station) to plot   
* Entire time series  
* Since we haven't removed NAs, min_year is alwyas 1994.... but that's OK here  
```{r}

tab <- xtabs(~ PARAM + STATION_CODE, data_med)
# tab

df_series <- data_med %>%
  # filter(!is.na(Q95)) %>%            # stations with at least 2 years of data, and existing PROREF
  group_by(PARAM, STATION_CODE, LATIN_NAME, TISSUE_NAME, Basis) %>%
  summarise(
    n = n(),
    min_year = min(MYEAR)
  ) %>%
  filter(n >= 2)     

```
### Test plot
```{r}

# debugonce(plot_medians_and_trends3)
# debugonce(calc_models_gam)
# debugonce(statistics_for_excel)

i <- 1
startyear <- df_series$min_year[i]
test <-  plot_medians_and_trends3(c(df_series$PARAM[i], 
                                    df_series$LATIN_NAME[i], df_series$TISSUE_NAME[i], 
                                    df_series$STATION_CODE[i], df_series$Basis[i]), 
                                  data_medians = data_med,
                                  trend_years = startyear:2021, trend_vjust = 1.6, trend_hjust = 2.2) 

test$gg
test$fn

# Manual ggplot
# test2 <- select_data(c(df_series$PARAM[i], 
#                                     df_series$LATIN_NAME[i], df_series$TISSUE_NAME[i], 
#                                     df_series$STATION_CODE[i], df_series$Basis[i]), 
#                                   data_medians = data_med)
# gg <- ggplot(test2, aes(MYEAR))
# gg <- gg + geom_smooth(aes(y = Median), method = "lm")
# gg 

#  annotate("text", x = Inf, y = Inf, label = "\u25CB", size = rel(13))

```

### Plot entire time series (using defaults) 

```{r}

folder <- "Input_data/510_Industry_data/Plots_Glomma_estuary/blaaskjell"
# dir.create(folder)

for (i in seq_len(nrow(df_series))){

  # For testing (make only plots 1 to 4), comment out the line above and uncomment the next line:
  # for (i in 1:10){
  # i <- 1

  startyear <- df_series$min_year[i]
  
  # debug(calc_models_one_station2)
  # debug(add_model_fits)
  X <-  plot_medians_and_trends3(c(df_series$PARAM[i], 
                                    df_series$LATIN_NAME[i], df_series$TISSUE_NAME[i], 
                                    df_series$STATION_CODE[i], df_series$Basis[i]), 
                                  data_medians = data_med,
                                  trend_years = startyear:2021, trend_vjust = 1.6, trend_hjust = 2.2) 
  
  # Fix x and y labels
  X$gg <- X$gg + 
    labs(x = "År", y = "Konsentrasjon")

  # X$gg
  ggsave("Input_data/510_Industry_data/Plots_Glomma_estuary/CR_test.png", X$gg)
  
  if (class(X) != "try-error"){
    save_trendplot(X, folder, suffix = "_allyears")
  }
  
}   

```


```{r}

folder
dir(folder, full.names = TRUE)
setwd(folder)
setwd("home/jovyan/shared/DHJ")
system("rm -r 00LOCK-ranger")
To go back you have to setwd to home/jovyan/shared/DHJ/icpw-R

system("tar -czvf ranger.tar.gz ../../../../../opt/R/ranger")
system("tar -czvf Input_data/510_Industry_data/Plots_Glomma_estuary/tang/* Input_data/510_Industry_data/Plots_Glomma_estuary")

```



## Blue mussel using 'plot_medians_and_trends2'  

* Kept for purely sentimental reasons  
    - Contains Proref and EQS (dropped in plot_medians_and_trends3 as the data are on dry-weight)  
    - English
    - Two trend symbols when only one is needed 

### Data   
* Copied from DHJs pc: Data/Seksjon 212/Glomma-estuary-metals
* Made using code from that project (script 01), also see github NIVANorge 'Glomma-estuary-metals'
```{r}

dat1 <- read.csv("Input_data/510_Industry_data/data_glomma_estuary_bluemussel.csv", row.names = "X")

```

### Check data  
```{r}

xtabs(~PARAM, dat1)

```

### Fix names  
```{r}

names(dat1)

dat2 <- dat1 %>%
  rename(
    MYEAR = Year,
    Median = Value) %>%
  mutate(
    N_median = 3,
    Over_LOQ = 3
  ) 

# names(data_med)
#  [1] "STATION_CODE"  "LATIN_NAME"    "TISSUE_NAME"   "PARAM"         "UNIT"          "MYEAR"         "N_median"      "Median"        "Over_LOQ"     
# [10] "Min"           "Max"           "Basis"         "NAME"          "STATION_NAME"  "Stations"      "N_stations"    "Proref_median" "Q95"

```


### Add Proref values    
```{r}

data_med <- dat2 %>%
  mutate(
    LATIN_NAME = "Mytilus edulis",
    TISSUE_NAME = "Whole soft body",
    STATION_CODE = "Kirkøy",
    Basis = "WW",
    UNIT = "MG_P_KG"
  ) #%>%
  # left_join(df_proref %>% select(-N) %>% rename(Proref_median = Median), 
  #          by = c("LATIN_NAME", "TISSUE_NAME", "PARAM", "Basis"))

# data_med %>% filter(PARAM == "BAP" & STATION_CODE == "B2") %>% View()

```

### Station name file  
* For printing in title  
```{r}

df_stationnames <- data.frame(
  STATION_CODE = c("Kirkøy"),
  Report_version_name = c("Kirkøy"),
  stringsAsFactors = FALSE)

```

### Select series (parameter + station) to plot   
* Entire time series  
* Since we haven't removed NAs, min_year is alwyas 1994.... but that's OK here  
```{r}

tab <- xtabs(~ PARAM + STATION_CODE, data_med)
# tab

df_series <- data_med %>%
  # filter(!is.na(Q95)) %>%            # stations with at least 2 years of data, and existing PROREF
  group_by(PARAM, STATION_CODE) %>%
  summarise(
    n = n(),
    min_year = min(MYEAR)
  ) %>%
  filter(n >= 2)     

```
### Plot entire time series (using defaults) 

```{r}

folder <- "Input_data/510_Industry_data/Plots_Glomma_estuary"
# dir.create(folder)

for (i in seq_len(nrow(df_series))){

  # For testing (make only plots 1 to 4), comment out the line above and uncomment the next line:
  # for (i in 1:10){
  # i <- 4

  startyear <- df_series$min_year[i]
  
  # debugonce(plot_medians_and_trends2)
  X <- try(
    plot_medians_and_trends2(c(df_series$PARAM[i], "Mytilus edulis", "Whole soft body", 
                               df_series$STATION_CODE[i], "WW"), 
                             eqs_type = "line", 
                             trend_years = startyear:2021, xlim = c(startyear - 7,2022),   
                             data_medians = data_med, data_proref = df_proref, 
                             data_eqs = df_EQS, eqs_x = 2017,
                             y_rel = 0.95, ymax_factor = 1.25,
                             show_proref = NA)                     # no proref, since this is dry-weight values  
  )
  # X$gg
  ggsave("Input_data/510_Industry_data/Plots_Glomma_estuary/CR_test.png", X$gg)
  
  if (class(X) != "try-error"){
    save_trendplot(X, folder, suffix = "_allyears")
  }
  
}

```


### Select series longer than 2004 for shorter plots   
* Entire time series  
* Since we haven't removed NAs, min_year is alwyas 1994.... but that's OK here  
```{r}

df_series2 <- data_med %>%
  filter(!is.na(Median)) %>%            # CHANGED
  group_by(PARAM, STATION_CODE) %>%
  summarise(
    n = n(),
    min_year = min(MYEAR)
  ) %>%
  filter(n >= 2 & min_year < 2004)                 # CHANGED

```


### Plot time series 2004-2021   


```{r}

folder <- "Input_data/510_Industry_data/Plots_Glomma_estuary"
# dir.create(folder)

for (i in seq_len(nrow(df_series2))){

  # For testing (make only plots 1 to 4), comment out the line above and uncomment the next line:
  # for (i in 1:10){
  # i <- 4

  startyear <- df_series2$min_year[i]
  
  # debugonce(plot_medians_and_trends2)
  X <- try(
    plot_medians_and_trends2(c(df_series2$PARAM[i], "Mytilus edulis", "Whole soft body", 
                               df_series2$STATION_CODE[i], "WW"), 
                             eqs_type = "line", 
                             trend_years = 2004:2021, xlim = c(1999, 2022),   
                             data_medians = data_med %>% filter(MYEAR >= 2004), 
                             data_proref = df_proref, 
                             data_eqs = df_EQS, eqs_x = 2017,
                             y_rel = 0.95, ymax_factor = 1.25,
                             show_proref = NA)                     # no proref, since this is dry-weight values
  )

  if (class(X) != "try-error"){
    save_trendplot(X, folder, suffix = "_2004-2021")
  }
  
}

```



