---
title: "915_Check_PFAS"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

## 1. Packages  
```{r}

library(dplyr)
library(purrr)
library(ggplot2)
library(forcats)
library(plotly)

```

## 2. Data 

### Main data  
Read and reformat the most recent data (by default)  
```{r, collapse=TRUE}

files <- dir("Data", pattern = "109_adjusted_data") %>% rev()

cat("Reading the last file downloaded:")
cat("\n", files[1])
cat("\n")
cat("If you want to read a different file, replace 'files[1]' with the file you want")
cat("\n\n")

filename <- files[1]
cat("Time since this file was modified: \n")
Sys.time() - file.info(paste0("Data/", filename))$mtime

data_all <- readRDS(paste0("Data/", filename))

# We save the date part of the text (e.g., '2020-04-23')
# This will be used in part 10, when we save the resulting file
pos1 <- regexpr("20", filename)[1]   # search for the first place "20" occurs inside filename
pos2 <- pos1 + 9
file_date <- substr(filename, pos1, pos2)    # pick out the part of the text from character no. 17 to no. 26

```

## 3. Percentage over LOQ, all stations    

### Table    
```{r}

df1 <- data_all %>% 
  filter(substr(PARAM, 1, 2) == "PF") %>%
  group_by(PARAM, MYEAR) %>%
  summarise(
    Over_LOQ = round(100*mean(is.na(FLAG1)),0)
    )# regrouping output by 'PARAM'

# For correct oredring of PARAM
# The same ordering used for Indre Oslofjord also  
df_param <- df1 %>%    
  # the 'mutate' and 'arrange' below is only for setting factors ('fct_inorder')    
  summarise(
    Over_LOQ_mean = mean(Over_LOQ, na.rm = TRUE)) %>%   
  arrange(desc(Over_LOQ_mean))

df1 <- df1 %>%
  mutate(
    PARAM = factor(PARAM, levels = df_param$PARAM)
  )
# levels(df1$PARAM)

df2 <- df1 %>%
  arrange(PARAM, MYEAR) %>%
  tidyr::pivot_wider(names_from = PARAM, values_from = Over_LOQ) %>%
  arrange(MYEAR)

df2

```

### Plot 

The most common ones  
```{r}

gg <- df1 %>%
  filter(PARAM %in% c("PFAS", "PFOS", "PFOSA")) %>%
  ggplot(aes(MYEAR, Over_LOQ, color = PARAM)) +
  geom_line() +
  geom_point()

gg

plotly::ggplotly(gg)  

```
The less common ones  
```{r}

gg <- df1 %>%
  filter(!PARAM %in% c("PFAS", "PFOS", "PFOSA")) %>%
  ggplot(aes(MYEAR, Over_LOQ, color = PARAM)) +
  geom_line() +
  geom_point()

gg

plotly::ggplotly(gg)  

```


### Table    
```{r}

df1 <- data_all %>% 
  filter(substr(PARAM, 1, 2) == "PF") %>%
  group_by(PARAM, MYEAR) %>%
  summarise(
    Over_LOQ = round(100*mean(is.na(FLAG1)),0)
    )# regrouping output by 'PARAM'

# For correct oredring of PARAM
df_param <- df1 %>%    
  # the 'mutate' and 'arrange' below is only for setting factors ('fct_inorder')    
  summarise(
    Over_LOQ_mean = mean(Over_LOQ, na.rm = TRUE)) %>%   
  arrange(desc(Over_LOQ_mean))

df1 <- df1 %>%
  mutate(
    PARAM = factor(PARAM, levels = df_param$PARAM)
  )
# levels(df1$PARAM)

df2 <- df1 %>%
  arrange(PARAM, MYEAR) %>%
  tidyr::pivot_wider(names_from = PARAM, values_from = Over_LOQ) %>%
  arrange(MYEAR)

df2

```


## 4. Percentage over LOQ, Oslo  

### Table    
```{r}

# Same ordering as for 'all stations' ('df_param' from part 3)

df1 <- data_all %>% 
  filter(substr(PARAM, 1, 2) == "PF" & STATION_CODE == "30B") %>%
  group_by(PARAM, MYEAR) %>%
  summarise(
    Over_LOQ = round(100*mean(is.na(FLAG1)),0)
    )  %>% # regrouping output by 'PARAM'
  mutate(
    PARAM = factor(PARAM, levels = df_param$PARAM)
  )
# levels(df1$PARAM)

df2 <- df1 %>%
  arrange(PARAM, MYEAR) %>%
  tidyr::pivot_wider(names_from = PARAM, values_from = Over_LOQ) %>%
  arrange(MYEAR)

df2

```

### Plot 

The most common ones  
```{r}

gg <- df1 %>%
  filter(PARAM %in% c("PFAS", "PFOS", "PFOSA")) %>%
  ggplot(aes(MYEAR, Over_LOQ, color = PARAM)) +
  geom_line() +
  geom_point()

gg

plotly::ggplotly(gg)  

```
The less common ones  
```{r}

gg <- df1 %>%
  filter(!PARAM %in% c("PFAS", "PFOS", "PFOSA")) %>%
  ggplot(aes(MYEAR, Over_LOQ, color = PARAM)) +
  geom_line() +
  geom_point()

gg

plotly::ggplotly(gg)  

```


## 5. Concentrations, all stations    

### Data      
```{r}

# Same ordering as for 'all stations' ('df_param' from part 3)

df1 <- data_all %>% 
  filter(substr(PARAM, 1, 2) == "PF") %>%
  mutate(Conc = case_when(
    is.na(FLAG1) ~ VALUE_WW,
    !is.na(FLAG1) ~ VALUE_WW/2)) %>%
  group_by(PARAM, STATION_CODE, MYEAR) %>%
  summarise(
    Mean_concentration = mean(Conc, na.rm = TRUE),
    Median_concentration = median(Conc, na.rm = TRUE),
    Max_concentration = max(Conc, na.rm = TRUE),
    LOQ = median(VALUE_WW[!is.na(FLAG1)], na.rm = TRUE)
    )  %>% # regrouping output by 'PARAM'
  mutate(
    PARAM = factor(PARAM, levels = df_param$PARAM),
    N_years = length(unique(MYEAR))                # per PARAM, STATION_CODE
  )
# levels(df1$PARAM)


```


### Plot the most common ones, mean    
```{r}

gg <- df1 %>%
  filter(PARAM %in% c("PFAS", "PFOS", "PFOSA") & N_years >= 5) %>%
  ggplot(aes(MYEAR, Mean_concentration, color = PARAM)) +
  geom_line() +
  geom_point() +
  facet_wrap(vars(STATION_CODE))

gg

# plotly::ggplotly(gg)  

```

The most common ones, mean since 2009    
```{r}

gg <- df1 %>%
  filter(PARAM %in% c("PFAS", "PFOS", "PFOSA") & N_years >= 5 & MYEAR >= 2009) %>%
  ggplot(aes(MYEAR, Mean_concentration, color = PARAM)) +
  geom_line() +
  geom_point() +
  facet_wrap(vars(STATION_CODE))

gg

# plotly::ggplotly(gg)  

```


### Plot the less common ones, mean    

* Strongly affected by high LOQ for PFOA and PFBS 2008, so these are deleted  
```{r}

gg1 <- df1 %>%
  filter(!PARAM %in% c("PFAS", "PFOS", "PFOSA") & N_years >= 5) %>%  
  ggplot(aes(MYEAR, Mean_concentration, color = PARAM)) +
  geom_line() +
  geom_point() +
  facet_wrap(vars(STATION_CODE))

plotly::ggplotly(gg1)  



```
### Plot less commons since 2009  
```{r}
gg2 <- df1 %>%
  filter(!PARAM %in% c("PFAS", "PFOS", "PFOSA") & N_years >= 5) %>%  
  filter(MYEAR >= 2009) %>%  
  ggplot(aes(MYEAR, Mean_concentration, color = PARAM)) +
  geom_line() +
  geom_point() +
  facet_wrap(vars(STATION_CODE))

gg2

```


## 6. Concentrations, Oslo  

### Data      
```{r}

# Same ordering as for 'all stations' ('df_param' from part 3)

df1 <- data_all %>% 
  filter(substr(PARAM, 1, 2) == "PF" & STATION_CODE == "30B") %>%
  mutate(Conc = case_when(
    is.na(FLAG1) ~ VALUE_WW,
    !is.na(FLAG1) ~ VALUE_WW/2)) %>%
  group_by(PARAM, MYEAR) %>%
  summarise(
    Mean_concentration = mean(Conc, na.rm = TRUE),
    Median_concentration = median(Conc, na.rm = TRUE),
    Max_concentration = max(Conc, na.rm = TRUE),
    LOQ = median(VALUE_WW[!is.na(FLAG1)], na.rm = TRUE)
    )  %>% # regrouping output by 'PARAM'
  mutate(
    PARAM = factor(PARAM, levels = df_param$PARAM)
  )
# levels(df1$PARAM)

df2 <- df1 %>%
  select(PARAM, MYEAR, Mean_concentration) %>%
  arrange(PARAM, MYEAR) %>%
  tidyr::pivot_wider(names_from = PARAM, values_from = Mean_concentration) %>%
  arrange(MYEAR)

df2

```

### LOQ   
2008 has very high LOQ sepcecially for PFOA and PFBS, but also for PFOS  
```{r}
df1 %>%
  select(PARAM, MYEAR, Mean_concentration) %>%
  arrange(PARAM, MYEAR) %>%
  tidyr::pivot_wider(names_from = PARAM, values_from = Mean_concentration) %>%
  arrange(MYEAR)
```


### Plot the most common ones, mean    
```{r}

gg <- df1 %>%
  filter(PARAM %in% c("PFAS", "PFOS", "PFOSA")) %>%
  ggplot(aes(MYEAR, Mean_concentration, color = PARAM)) +
  geom_line() +
  geom_point()

gg

plotly::ggplotly(gg)  

```

The most common ones, max    
```{r}

gg <- df1 %>%
  filter(PARAM %in% c("PFAS", "PFOS", "PFOSA")) %>%
  ggplot(aes(MYEAR, Max_concentration, color = PARAM)) +
  geom_line() +
  geom_point()

gg

plotly::ggplotly(gg)  

```

### Plot the less common ones, mean    

* Strongly affected by high LOQ for PFOA and PFBS 2008, so these are deleted  
```{r}

gg1 <- df1 %>%
  filter(!PARAM %in% c("PFAS", "PFOS", "PFOSA")) %>%  
  ggplot(aes(MYEAR, Mean_concentration, color = PARAM)) +
  geom_line() +
  geom_point()

plotly::ggplotly(gg1)  

gg2 <- df1 %>%
  filter(!PARAM %in% c("PFAS", "PFOS", "PFOSA")) %>%  
  filter(MYEAR >= 2009) %>%  
  ggplot(aes(MYEAR, Mean_concentration, color = PARAM)) +
  geom_line() +
  geom_point()

gg2

```

